## Stylized fact A: Capital Distribution Curve (CDC) and stability

The empirical CDC is the mean (across weeks) of `share_topK` at each rank within top-K.

```{r cdc-setup, include=FALSE}
if (!exists("endpoint_bundle")) {
  endpoint_bundle <- load_endpoint_weekly(cfg = CFG)
}
endpoint_weekly <- endpoint_bundle$endpoint_weekly
K_cut <- endpoint_bundle$K_cut

horizons_growth <- CFG$horizons_growth

if (!exists("emp_cdc")) {
  emp_cdc <- load_emp_cdc(endpoint_weekly, K_cut, cfg = CFG)
}
```

```{r emp-cdc}
emp_cdc

ggplot(emp_cdc, aes(rank, w_bar)) +
  geom_line() +
  scale_x_log10() +
  scale_y_log10(labels = percent_format(accuracy = 0.01)) +
  labs(
    title = "Empirical CDC (mean share within top-K)",
    subtitle = paste0("K_cut = ", K_cut),
    x = "Rank (log scale)",
    y = "Mean share within top-K (log scale)"
  ) +
  theme_bw()
```

### CDC stability over time windows

A simple stability check: compute period-level CDCs and compare to the full-sample CDC.

```{r cdc-stability}
cdc_by_period <- compute_cdc_by_period(endpoint_weekly, K_cut, period_unit = "quarter")

ggplot(cdc_by_period, aes(rank, w_bar, group = period)) +
  geom_line(alpha = 0.25) +
  scale_x_log10() +
  scale_y_log10(labels = percent_format(accuracy = 0.01)) +
  labs(
    title = "CDC curves by quarter (visual stability check)",
    x = "Rank (log scale)",
    y = "Mean share within top-K (log scale)"
  ) +
  theme_bw()

cdc_rmse_vs_full <- cdc_by_period %>%
  inner_join(emp_cdc, by = "rank", suffix = c("_period", "_full")) %>%
  group_by(period) %>%
  summarise(
    rmse_log = sqrt(mean((log(w_bar_period) - log(w_bar_full))^2, na.rm = TRUE)),
    .groups = "drop"
  )

ggplot(cdc_rmse_vs_full, aes(period, rmse_log)) +
  geom_line() +
  geom_point(size = 1.5) +
  labs(
    title = "CDC stability: RMSE (log shares) vs full-sample curve",
    x = NULL,
    y = "RMSE in log space"
  ) +
  theme_bw()
```

## Endpoint-level forward growth distributions (v2/v3)

Compute forward-looking log growth in *global share* for individual endpoints over multiple horizons,
and compare distributions across rank buckets.

```{r endpoint-growth}
endpoint_base <- endpoint_weekly %>%
  transmute(
    week,
    endpoint_id,
    rank,
    bucket = rank_bucket_simple(rank),
    log_share = log(pmax(share_global, 1e-12))
  )

growth_long <- purrr::map_dfr(horizons_growth, function(h) {
  endpoint_base %>%
    mutate(
      horizon_weeks = as.integer(h),
      week_tp = week + lubridate::weeks(as.integer(h))
    ) %>%
    left_join(
      endpoint_base %>%
        transmute(endpoint_id, week_tp = week, log_share_tp = log_share),
      by = c("endpoint_id", "week_tp")
    ) %>%
    mutate(dlog_share = log_share_tp - log_share) %>%
    filter(is.finite(dlog_share))
})

growth_summ <- growth_long %>%
  group_by(bucket, horizon_weeks) %>%
  summarise(
    n = n(),
    med = median(dlog_share, na.rm = TRUE),
    p10 = quantile(dlog_share, 0.10, na.rm = TRUE),
    p90 = quantile(dlog_share, 0.90, na.rm = TRUE),
    .groups = "drop"
  )

growth_summ
```

```{r endpoint-growth-plot}
ggplot(growth_long, aes(x = dlog_share, fill = bucket)) +
  geom_density(alpha = 0.35) +
  facet_wrap(~ horizon_weeks, scales = "free_y") +
  labs(
    title = "Forward log-growth in global share by rank bucket",
    subtitle = "Endpoint-level growth distributions (identity-based), as in v2/v3",
    x = "Delta log(share_global) over horizon",
    y = "Density",
    fill = "Bucket"
  ) +
  theme_bw()
```
