```{r model-zoo-setup, include=FALSE}
if (!exists("endpoint_bundle")) {
  endpoint_bundle <- load_endpoint_weekly(cfg = CFG)
}

fit_end_week <- CFG$model_zoo_fit_end_week %||% as.Date("2022-07-01")
endpoint_weekly <- endpoint_bundle$endpoint_weekly %>%
  filter(as.Date(week) < fit_end_week)
if (nrow(endpoint_weekly) == 0L) {
  stop("No model-zoo data before fit_end_week=", as.character(fit_end_week))
}

max_rank_seen <- endpoint_weekly %>%
  summarise(max_rank = max(rank, na.rm = TRUE)) %>%
  pull(max_rank) %>%
  as.integer()
K_cut <- min(endpoint_bundle$K_cut, max_rank_seen)
K_max <- min(endpoint_bundle$K_max, max_rank_seen)
sim_K_xi <- min(endpoint_bundle$sim_K_xi, K_cut)

bucket_def <- CFG$bucket_def
Smoothing_h <- CFG$smoothing_h
pca_K <- CFG$pca_K

rank_panel <- build_rank_panel(endpoint_weekly)
gauss_params_raw <- compute_gauss_params_raw(rank_panel, K_max)
sm_params <- build_gaussian_params_smoothed(gauss_params_raw, K_max, h = Smoothing_h)

message(
  "[model-zoo-window] week < ", as.character(fit_end_week),
  " | endpoint rows=", nrow(endpoint_weekly),
  " | rank_panel rows=", nrow(rank_panel),
  " | max endpoint week=", as.character(max(endpoint_weekly$week, na.rm = TRUE)),
  " | max rank_panel week=", as.character(max(rank_panel$week, na.rm = TRUE))
)
```

## Stylized fact E: Common shocks vs idiosyncratic movement (new)

A pragmatic diagnostic: PCA on rank-slot increments for the top `pca_K` ranks.

```{r common-shocks-pca}
K_pca <- min(pca_K, K_cut)

pca_cache <- list(
  beta_k = rep(0, K_max),
  pc1_var_explained = NA_real_,
  pc1 = NULL
)

if (is.na(K_pca) || K_pca < 2) {
  message("Skipping PCA: K_pca < 2 (K_pca=", K_pca, ").")
} else {
  rank_slot_wide <- rank_panel %>%
    filter(rank <= K_pca) %>%
    select(week, rank, dlogw) %>%
    pivot_wider(names_from = rank, values_from = dlogw)

  rank_slot_mat <- rank_slot_wide %>%
    drop_na() %>%
    arrange(week)

  X <- rank_slot_mat %>%
    select(-week) %>%
    as.matrix()

  if (nrow(X) < 2 || ncol(X) < 2) {
    message(
      "Skipping PCA: need at least 2 complete weeks and 2 rank columns; got ",
      nrow(X), "x", ncol(X), "."
    )
  } else {
    pca_fit <- prcomp(X, center = TRUE, scale. = TRUE)

    var_explained <- (pca_fit$sdev^2) / sum(pca_fit$sdev^2)

    tibble(
      component = seq_along(var_explained),
      var_explained = var_explained
    ) %>%
      slice(1:10)

    ggplot(
      tibble(component = seq_along(var_explained), var_explained = var_explained) %>% slice(1:20),
      aes(component, var_explained)
    ) +
      geom_line() +
      geom_point(size = 1.5) +
      scale_y_continuous(labels = percent_format(accuracy = 0.1)) +
      labs(
        title = "PCA on rank-slot increments: variance explained",
        subtitle = paste0("Top ", K_pca, " ranks; complete weeks only"),
        x = "Principal component",
        y = "Share of variance explained"
      ) +
      theme_bw()

    pc1 <- pca_fit$x[, 1]
    pc1_df <- tibble(week = rank_slot_mat$week, pc1 = pc1)

    ggplot(pc1_df, aes(week, pc1)) +
      geom_line() +
      labs(
        title = "First principal component over time (common shock proxy)",
        x = "Week",
        y = "PC1 score"
      ) +
      theme_bw()

    # cache loadings for jump-factor model (smoothed, padded to K_max)
    beta_raw <- pca_fit$rotation[, 1]
    beta_vec <- rep(0, K_max)
    beta_vec[seq_len(K_pca)] <- beta_raw
    beta_vec <- moving_average_rank(beta_vec, h = as.integer(Smoothing_h))
    beta_scale <- sqrt(mean(beta_vec[seq_len(K_pca)]^2, na.rm = TRUE))
    if (is.finite(beta_scale) && beta_scale > 0) {
      beta_vec <- beta_vec / beta_scale
    }

    pca_cache$beta_k <- beta_vec
    pca_cache$pc1_var_explained <- var_explained[1]
    pca_cache$pc1 <- pc1_df
  }
}
```

## Tail & skew diagnostics for increments (empirical)

This section checks whether rank-slot and identity-based increments have heavy tails and directional skew, including robust skew/kurtosis and a Hill tail-index estimate.

```{r tail-skew-setup}
rank_inc <- build_rank_slot_increments(rank_panel %>% filter(rank <= K_cut), K_cut)
rank_inc_std <- standardize_increments(rank_inc, sm_params$mean_dlogw_s, sm_params$sd_dlogw_s, bucket_def = bucket_def) %>%
  mutate(z = dlogw_std)

identity_inc <- build_identity_increments(endpoint_weekly, horizon = 1L)
identity_inc_std <- standardize_increments(identity_inc, sm_params$mean_dlogw_s, sm_params$sd_dlogw_s, bucket_def = bucket_def) %>%
  mutate(z = dlogw_std)

inc_long <- bind_rows(
  rank_inc_std %>% mutate(type = "rank_slot") %>% select(week, rank, bucket, dlogw, dlogw_std, type),
  identity_inc_std %>% mutate(type = "identity") %>% select(week, endpoint_id, rank, bucket, dlogw, dlogw_std, type)
)

tail_rank <- tail_skew_diagnostics(rank_inc_std, thresholds = CFG$model_zoo_tail_thresholds, probs = CFG$model_zoo_tail_probs)
tail_identity <- tail_skew_diagnostics(identity_inc_std, thresholds = CFG$model_zoo_tail_thresholds, probs = CFG$model_zoo_tail_probs)

tail_rank
tail_identity
```

```{r jump-clustering}
cluster_rank <- jump_clustering_diagnostics(rank_inc_std, thresholds = c(3, 4, 5))
cluster_rank
```

If clustering is material (positive autocorrelation in extreme indicators), set `CFG$enable_jump_clustering_models = TRUE` to include self-exciting / regime-switching variants in the model zoo.

```{r qq-plot-normal, echo=FALSE, out.width="85%"}
plot_dir <- here::here("output", "plots")
dir.create(plot_dir, showWarnings = FALSE, recursive = TRUE)

qq_normal_plot <- ggplot(rank_inc_std, aes(sample = z)) +
  stat_qq() +
  stat_qq_line() +
  facet_wrap(~ bucket, scales = "free") +
  labs(
    title = "QQ plot of standardized increments by bucket",
    subtitle = "Normal reference",
    x = "Theoretical quantiles",
    y = "Sample quantiles"
  ) +
  theme_bw()

qq_normal_path <- file.path(plot_dir, "qq_normal_rank.png")
ggsave(qq_normal_path, qq_normal_plot, dpi = 300, width = 8, height = 5, units = "in")
knitr::include_graphics(qq_normal_path)
```

```{r qq-plot-t, echo=FALSE, out.width="85%"}
fit_t_rank <- fit_student_t(rank_inc_std, mode = "bucket", df_min = CFG$model_zoo_df_min)
qq_t_rank <- rank_inc_std %>%
  group_by(bucket) %>%
  group_modify(~{
    b <- .y$bucket
    df_hat <- fit_t_rank$params %>% filter(bucket == b) %>% pull(df)
    df_hat <- df_hat[1]
    if (!is.finite(df_hat) || length(df_hat) == 0) df_hat <- max(6, CFG$model_zoo_df_min + 1)
    qq_df <- qqplot_data(.x$z, dist = "t", params = list(df = df_hat))
    qq_df
  }) %>%
  ungroup()

qq_t_plot <- ggplot(qq_t_rank, aes(x = theo_q, y = sample_q)) +
  geom_point(size = 0.7, alpha = 0.6) +
  geom_abline(linetype = "dashed", color = "gray40") +
  facet_wrap(~ bucket, scales = "free") +
  labs(
    title = "QQ plot vs Student-t (fitted df)",
    x = "Theoretical quantiles",
    y = "Sample quantiles"
  ) +
  theme_bw()

qq_t_path <- file.path(plot_dir, "qq_t_rank.png")
ggsave(qq_t_path, qq_t_plot, dpi = 300, width = 8, height = 5, units = "in")
knitr::include_graphics(qq_t_path)
```

## Simulation framework (v8–v11) + unified fit metrics

### 12.1 Empirical entrant pool for entry/exit

```{r entrant-pool}
entrant_pool <- build_entrant_pool(endpoint_weekly, K_cut, CFG$K_tail_buffer, max_rank_seen)

stopifnot(nrow(entrant_pool) > 100)

entrant_sampler <- make_entrant_sampler(entrant_pool)

w0_ext <- build_w0_ext(endpoint_weekly, K_max, entrant_sampler)
```

### 12.2 Simulator (Gaussian increments with smoothed moments; optional Xi output)

This merges v9’s simulator with v11’s Xi storage.

### 12.3 Fit metrics (CDC, durable change, Xi)

```{r fit-metrics, include=FALSE}
emp_cdc <- compute_emp_cdc(endpoint_weekly, K_cut)
emp_targets <- compute_emp_endpoint_change(
  df = endpoint_weekly %>% filter(rank <= K_cut),
  horizons = CFG$horizons_durable,
  bucket_def = bucket_def
)
```
