---
title: 'Facebook Weekly Rank Diffusion: Master EDA + Model Fit'
output:
  pdf_document:
    toc: true
    toc_depth: '3'
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
params:
  fb_weekly_parquet: !r here::here("Facebook", "fb_ranked_weekly.parquet")
  seed: 1823
  K_cut_target: 12000L
  K_tail_buffer: 2000L
  horizons_growth: !r c(1L, 7L, 28L)
  horizons_durable: !r c(4L, 8L)
  horizons_micro: !r c(1L, 4L, 12L)
  bucket_breaks: !r c(0, 25, 250, Inf)
  bucket_labels: !r c("large", "midsize", "small")
  smoothing_h: 5L
  sim_T_weeks: 52L
  sim_n_paths: 200L
  sim_mu: 0
  sim_sigma: 0.16
  sim_entry_frac: 0.1
  sim_K_xi: 500L
  turnover_K_list: !r c(25L, 250L, 1000L)
  turnover_horizons: !r c(1L, 4L, 12L)
  pca_K: 500L
  run_grid_search: false
  run_bootstrap_model: false
  run_jump_model: false
  jump_model_type: "jump_factor"
  jump_df_F: 5
  jump_phi_F: 0
  jump_df_eps_large: 6
  jump_df_eps_midsize: 8
  jump_df_eps_small: 10
  jump_p_sys: 0.02
  jump_p_idio_large: 0.01
  jump_p_idio_midsize: 0.005
  jump_p_idio_small: 0.002
  jump_dist: "laplace"
  jump_scale_sys: 0.35
  jump_scale_idio: 0.20
  jump_scale: 0.10
  jump_df: 5
  jump_sys_loading: "beta"
  jump_mean_reversion: true
  jump_kappa_large: 0.04
  jump_kappa_midsize: 0.03
  jump_kappa_small: 0.02
  bootstrap_B: 200L
  bootstrap_k_min: 10L
  bootstrap_period_unit: quarter
  bootstrap_method: "local_rank"
  bootstrap_rank_bandwidth: 25L
  bootstrap_kernel: "triangular"
  bootstrap_block_bootstrap: false
  bootstrap_block_length: 4L
  bootstrap_n_factors: 1L
  bootstrap_K_pca: 500L
  bootstrap_tailsplice: false
  bootstrap_tail_q: 0.99
  bootstrap_tail_fit_min_n: 30L
  bootstrap_diag_paths: 3L
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 9,
  fig.height = 5
)
```

## Setup

```{r libraries}
library(tidyverse)
library(arrow)
library(lubridate)
library(scales)
library(here)
library(broom)
```

```{r resolve-data-paths}
fb_weekly_parquet_path <- params$fb_weekly_parquet
if (!file.exists(fb_weekly_parquet_path)) {
  candidates <- c(
    here::here("fb_ranked_weekly.parquet"),
    here::here("Facebook", "fb_ranked_weekly.parquet")
  )
  fb_weekly_parquet_path <- candidates[file.exists(candidates)][1]
}

if (is.na(fb_weekly_parquet_path) || !file.exists(fb_weekly_parquet_path)) {
  stop(
    "Weekly parquet not found. Tried: ",
    paste(c(params$fb_weekly_parquet,
            here::here("fb_ranked_weekly.parquet"),
            here::here("Facebook", "fb_ranked_weekly.parquet")),
          collapse = ", ")
  )
}
```

```{r seed}
set.seed(params$seed)
```

```{r normalize-params}
coerce_numeric <- function(x) {
  if (is.null(x)) return(NA_real_)
  if (is.numeric(x)) return(as.numeric(x))
  if (is.character(x)) {
    return(suppressWarnings(readr::parse_number(x)))
  }
  suppressWarnings(as.numeric(x))
}

coerce_numeric_vec <- function(x) {
  if (is.null(x)) return(numeric(0))
  if (is.list(x)) x <- unlist(x, use.names = FALSE)
  if (is.numeric(x)) return(as.numeric(x))
  if (is.character(x)) return(readr::parse_number(x))
  suppressWarnings(as.numeric(x))
}

K_cut_target <- coerce_numeric(params$K_cut_target)
K_tail_buffer <- coerce_numeric(params$K_tail_buffer)
Smoothing_h <- coerce_numeric(params$smoothing_h)
bootstrap_B <- coerce_numeric(params$bootstrap_B)
bootstrap_k_min <- coerce_numeric(params$bootstrap_k_min)
bootstrap_method <- params$bootstrap_method
bootstrap_rank_bandwidth <- params$bootstrap_rank_bandwidth
bootstrap_kernel <- params$bootstrap_kernel
bootstrap_block_bootstrap <- isTRUE(params$bootstrap_block_bootstrap)
bootstrap_block_length <- as.integer(coerce_numeric(params$bootstrap_block_length))
bootstrap_n_factors <- as.integer(coerce_numeric(params$bootstrap_n_factors))
bootstrap_K_pca <- as.integer(coerce_numeric(params$bootstrap_K_pca))
bootstrap_tailsplice <- isTRUE(params$bootstrap_tailsplice)
bootstrap_tail_q <- coerce_numeric(params$bootstrap_tail_q)
bootstrap_tail_fit_min_n <- as.integer(coerce_numeric(params$bootstrap_tail_fit_min_n))
bootstrap_diag_paths <- as.integer(coerce_numeric(params$bootstrap_diag_paths))
sim_T_weeks <- as.integer(coerce_numeric(params$sim_T_weeks))
sim_n_paths <- as.integer(coerce_numeric(params$sim_n_paths))
sim_mu <- coerce_numeric(params$sim_mu)
sim_sigma <- coerce_numeric(params$sim_sigma)
sim_entry_frac <- coerce_numeric(params$sim_entry_frac)
sim_K_xi <- as.integer(coerce_numeric(params$sim_K_xi))
pca_K <- as.integer(coerce_numeric(params$pca_K))


jump_model_type <- if (is.null(params$jump_model_type) || is.na(params$jump_model_type)) {
  "jump_factor"
} else {
  as.character(params$jump_model_type)
}

jump_dist <- if (is.null(params$jump_dist) || is.na(params$jump_dist)) {
  "laplace"
} else {
  as.character(params$jump_dist)
}

jump_sys_loading <- if (is.null(params$jump_sys_loading) || is.na(params$jump_sys_loading)) {
  "beta"
} else {
  as.character(params$jump_sys_loading)
}

jump_df_F <- coerce_numeric(params$jump_df_F)
jump_phi_F <- coerce_numeric(params$jump_phi_F)

jump_df_eps_large <- coerce_numeric(params$jump_df_eps_large)
jump_df_eps_midsize <- coerce_numeric(params$jump_df_eps_midsize)
jump_df_eps_small <- coerce_numeric(params$jump_df_eps_small)

jump_p_sys <- coerce_numeric(params$jump_p_sys)
jump_p_idio_large <- coerce_numeric(params$jump_p_idio_large)
jump_p_idio_midsize <- coerce_numeric(params$jump_p_idio_midsize)
jump_p_idio_small <- coerce_numeric(params$jump_p_idio_small)

jump_scale_sys <- coerce_numeric(params$jump_scale_sys)
jump_scale_idio <- coerce_numeric(params$jump_scale_idio)
jump_scale <- coerce_numeric(params$jump_scale)
jump_df <- coerce_numeric(params$jump_df)

jump_mean_reversion <- isTRUE(params$jump_mean_reversion)

jump_kappa_large <- coerce_numeric(params$jump_kappa_large)
jump_kappa_midsize <- coerce_numeric(params$jump_kappa_midsize)
jump_kappa_small <- coerce_numeric(params$jump_kappa_small)

bootstrap_method <- if (is.null(bootstrap_method)) "iid_rank" else as.character(bootstrap_method)
bootstrap_kernel <- if (is.null(bootstrap_kernel)) "uniform" else as.character(bootstrap_kernel)
if (!is.function(bootstrap_rank_bandwidth)) {
  bootstrap_rank_bandwidth <- coerce_numeric(bootstrap_rank_bandwidth)
}
if (is.na(bootstrap_rank_bandwidth) && !is.function(bootstrap_rank_bandwidth)) {
  bootstrap_rank_bandwidth <- 25L
}
if (is.na(bootstrap_block_length)) bootstrap_block_length <- 4L
if (is.na(bootstrap_n_factors) || bootstrap_n_factors < 1L) bootstrap_n_factors <- 1L
if (is.na(bootstrap_K_pca) || bootstrap_K_pca < 2L) bootstrap_K_pca <- pca_K
if (is.na(bootstrap_tail_q)) bootstrap_tail_q <- 0.99
if (is.na(bootstrap_tail_fit_min_n)) bootstrap_tail_fit_min_n <- 30L
if (is.na(bootstrap_diag_paths) || bootstrap_diag_paths < 0L) bootstrap_diag_paths <- 0L
bootstrap_tailsplice <- isTRUE(bootstrap_tailsplice) || identical(bootstrap_method, "factor_local_tailsplice")

horizons_growth <- as.integer(coerce_numeric_vec(params$horizons_growth))
horizons_durable <- as.integer(coerce_numeric_vec(params$horizons_durable))
horizons_micro <- as.integer(coerce_numeric_vec(params$horizons_micro))
turnover_K_list <- as.integer(coerce_numeric_vec(params$turnover_K_list))
turnover_horizons <- as.integer(coerce_numeric_vec(params$turnover_horizons))

if (is.na(K_cut_target) || is.na(K_tail_buffer) || is.na(Smoothing_h)) {
  stop(
    "K_cut_target, K_tail_buffer, and smoothing_h must be numeric. Got: ",
    "K_cut_target=", params$K_cut_target, ", ",
    "K_tail_buffer=", params$K_tail_buffer, ", ",
    "smoothing_h=", params$smoothing_h
  )
}

if (Smoothing_h < 0) {
  stop("smoothing_h must be >= 0. Got: ", params$smoothing_h)
}

if (is.na(bootstrap_B) || bootstrap_B < 1) {
  stop("bootstrap_B must be a positive integer. Got: ", params$bootstrap_B)
}

if (is.na(bootstrap_k_min) || bootstrap_k_min < 1) {
  stop("bootstrap_k_min must be a positive integer. Got: ", params$bootstrap_k_min)
}

valid_bootstrap_methods <- c(
  "iid_rank", "local_rank", "factor_local", "week_vector", "factor_local_tailsplice"
)
if (!bootstrap_method %in% valid_bootstrap_methods) {
  stop(
    "bootstrap_method must be one of: ", paste(valid_bootstrap_methods, collapse = ", "),
    ". Got: ", bootstrap_method
  )
}

valid_bootstrap_kernels <- c("uniform", "triangular", "exp")
if (!bootstrap_kernel %in% valid_bootstrap_kernels) {
  stop(
    "bootstrap_kernel must be one of: ", paste(valid_bootstrap_kernels, collapse = ", "),
    ". Got: ", bootstrap_kernel
  )
}

if (!is.function(bootstrap_rank_bandwidth)) {
  if (is.na(bootstrap_rank_bandwidth) || bootstrap_rank_bandwidth < 0) {
    stop("bootstrap_rank_bandwidth must be >= 0. Got: ", params$bootstrap_rank_bandwidth)
  }
}

if (isTRUE(bootstrap_block_bootstrap) && (is.na(bootstrap_block_length) || bootstrap_block_length < 1)) {
  stop("bootstrap_block_length must be a positive integer when block bootstrap is on. Got: ",
       params$bootstrap_block_length)
}

if (is.na(bootstrap_n_factors) || bootstrap_n_factors < 1) {
  stop("bootstrap_n_factors must be >= 1. Got: ", params$bootstrap_n_factors)
}

if (is.na(bootstrap_K_pca) || bootstrap_K_pca < 2) {
  stop("bootstrap_K_pca must be >= 2. Got: ", params$bootstrap_K_pca)
}

if (is.na(bootstrap_tail_q) || bootstrap_tail_q <= 0.5 || bootstrap_tail_q >= 1) {
  stop("bootstrap_tail_q must be in (0.5, 1). Got: ", params$bootstrap_tail_q)
}

if (is.na(bootstrap_tail_fit_min_n) || bootstrap_tail_fit_min_n < 10) {
  stop("bootstrap_tail_fit_min_n must be >= 10. Got: ", params$bootstrap_tail_fit_min_n)
}

if (is.na(bootstrap_diag_paths) || bootstrap_diag_paths < 0) {
  stop("bootstrap_diag_paths must be >= 0. Got: ", params$bootstrap_diag_paths)
}

if (is.na(sim_T_weeks) || sim_T_weeks < 1) {
  stop("sim_T_weeks must be a positive integer. Got: ", params$sim_T_weeks)
}

if (is.na(sim_n_paths) || sim_n_paths < 1) {
  stop("sim_n_paths must be a positive integer. Got: ", params$sim_n_paths)
}

if (is.na(sim_mu) || is.na(sim_sigma) || is.na(sim_entry_frac)) {
  stop(
    "sim_mu, sim_sigma, and sim_entry_frac must be numeric. Got: ",
    "sim_mu=", params$sim_mu, ", ",
    "sim_sigma=", params$sim_sigma, ", ",
    "sim_entry_frac=", params$sim_entry_frac
  )
}

if (sim_entry_frac < 0 || sim_entry_frac > 1) {
  stop("sim_entry_frac must be in [0, 1]. Got: ", params$sim_entry_frac)
}

if (is.na(sim_K_xi) || sim_K_xi < 2) {
  stop("sim_K_xi must be an integer >= 2. Got: ", params$sim_K_xi)
}

horizons_growth <- horizons_growth[!is.na(horizons_growth) & horizons_growth >= 1L]
horizons_durable <- horizons_durable[!is.na(horizons_durable) & horizons_durable >= 1L]
horizons_micro <- horizons_micro[!is.na(horizons_micro) & horizons_micro >= 1L]
turnover_K_list <- turnover_K_list[!is.na(turnover_K_list) & turnover_K_list >= 1L]
turnover_horizons <- turnover_horizons[!is.na(turnover_horizons) & turnover_horizons >= 1L]

if (length(horizons_growth) == 0 || length(horizons_durable) == 0 || length(horizons_micro) == 0) {
  stop(
    "horizons_growth, horizons_durable, and horizons_micro must include at least one integer >= 1."
  )
}

if (length(turnover_K_list) == 0 || length(turnover_horizons) == 0) {
  stop(
    "turnover_K_list and turnover_horizons must include at least one integer >= 1."
  )
}

max_h_durable <- max(horizons_durable)
if (sim_T_weeks < max_h_durable) {
  message(
    "sim_T_weeks (", sim_T_weeks, ") < max(horizons_durable) (", max_h_durable,
    "); using sim_T_weeks=", max_h_durable, "."
  )
  sim_T_weeks <- max_h_durable
}
```

## Load weekly Facebook data and construct core panel

The exploration PDFs treat the weekly Facebook file as weekly snapshots with (at least):
`date`, `endpoint_id`, `metric_value`, `rank` (rank is recomputed defensively below).

We construct:

- `share_global`: each endpoint's share of total weekly attention (sums to 1 within week).
- `share_topK`: same share, renormalized within the top-K (`K_cut`) each week.

```{r load-data}
raw <- read_parquet(fb_weekly_parquet_path)

endpoint_weekly <- raw %>%
  transmute(
    week = as.Date(date),
    endpoint_id,
    metric_value,
    rank = rank
  ) %>%
  group_by(week) %>%
  arrange(desc(metric_value), .by_group = TRUE) %>%
  mutate(
    rank = row_number(),
    share_global = metric_value / sum(metric_value, na.rm = TRUE)
  ) %>%
  ungroup()

max_rank_by_week <- endpoint_weekly %>%
  group_by(week) %>%
  summarise(max_rank = max(rank), .groups = "drop")

max_rank_seen <- max(max_rank_by_week$max_rank, na.rm = TRUE)

K_cut <- min(as.integer(K_cut_target), as.integer(max_rank_seen))
K_max <- min(as.integer(K_cut + K_tail_buffer), as.integer(max_rank_seen))

if (sim_K_xi > K_cut) {
  message("sim_K_xi (", sim_K_xi, ") > K_cut (", K_cut, "); using sim_K_xi=", K_cut, ".")
  sim_K_xi <- K_cut
}

endpoint_weekly <- endpoint_weekly %>%
  group_by(week) %>%
  mutate(
    share_topK = if_else(
      rank <= K_cut,
      share_global / sum(share_global[rank <= K_cut], na.rm = TRUE),
      NA_real_
    )
  ) %>%
  ungroup()

glimpse(endpoint_weekly)
```

## Data QA and coverage

This reproduces the “weekly coverage & scale checks” used in v6.

```{r qa-summary}
endpoint_weekly %>%
  summarise(
    min_week = min(week),
    max_week = max(week),
    n_weeks = n_distinct(week),
    n_endpoints = n_distinct(endpoint_id),
    max_rank_seen = max(rank)
  )
```

```{r qa-weekly-row-counts}
weekly_totals <- endpoint_weekly %>%
  count(week, name = "n_rows") %>%
  arrange(week)

summary(weekly_totals$n_rows)

ggplot(weekly_totals, aes(week, n_rows)) +
  geom_line() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Weekly row counts (weekly snapshots)",
    x = "Week",
    y = "Rows"
  ) +
  theme_bw()
```

```{r qa-max-rank}
ggplot(max_rank_by_week, aes(week, max_rank)) +
  geom_line() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Maximum observed rank by week",
    x = "Week",
    y = "Max rank"
  ) +
  theme_bw()
```

## Stylized fact A: Capital Distribution Curve (CDC) and stability

The empirical CDC is the mean (across weeks) of `share_topK` at each rank within top-K.

```{r emp-cdc}
emp_cdc <- endpoint_weekly %>%
  filter(rank <= K_cut) %>%
  group_by(rank) %>%
  summarise(w_bar = mean(share_topK, na.rm = TRUE), .groups = "drop") %>%
  arrange(rank)

ggplot(emp_cdc, aes(rank, w_bar)) +
  geom_line() +
  scale_x_log10() +
  scale_y_log10(labels = percent_format(accuracy = 0.01)) +
  labs(
    title = "Empirical CDC (mean share within top-K)",
    subtitle = paste0("K_cut = ", K_cut),
    x = "Rank (log scale)",
    y = "Mean share within top-K (log scale)"
  ) +
  theme_bw()
```

### CDC stability over time windows

A simple stability check: compute period-level CDCs and compare to the full-sample CDC.

```{r cdc-stability}
cdc_by_period <- endpoint_weekly %>%
  filter(rank <= K_cut) %>%
  mutate(period = floor_date(week, "quarter")) %>%
  group_by(period, rank) %>%
  summarise(w_bar = mean(share_topK, na.rm = TRUE), .groups = "drop")

ggplot(cdc_by_period, aes(rank, w_bar, group = period)) +
  geom_line(alpha = 0.25) +
  scale_x_log10() +
  scale_y_log10(labels = percent_format(accuracy = 0.01)) +
  labs(
    title = "CDC curves by quarter (visual stability check)",
    x = "Rank (log scale)",
    y = "Mean share within top-K (log scale)"
  ) +
  theme_bw()

cdc_rmse_vs_full <- cdc_by_period %>%
  inner_join(emp_cdc, by = "rank", suffix = c("_period", "_full")) %>%
  group_by(period) %>%
  summarise(
    rmse_log = sqrt(mean((log(w_bar_period) - log(w_bar_full))^2, na.rm = TRUE)),
    .groups = "drop"
  )

ggplot(cdc_rmse_vs_full, aes(period, rmse_log)) +
  geom_line() +
  geom_point(size = 1.5) +
  labs(
    title = "CDC stability: RMSE (log shares) vs full-sample curve",
    x = NULL,
    y = "RMSE in log space"
  ) +
  theme_bw()
```

## Endpoint-level forward growth distributions (v2/v3)

This reproduces the early v2/v3 exploratory step: compute forward-looking log growth in *global share*
for individual endpoints over multiple horizons, and compare distributions across rank buckets.

```{r endpoint-growth}
rank_bucket_simple <- function(r) {
  case_when(
    r <= 25 ~ "large",
    r <= 250 ~ "midsize",
    TRUE ~ "small"
  )
}

endpoint_base <- endpoint_weekly %>%
  transmute(
    week,
    endpoint_id,
    rank,
    bucket = rank_bucket_simple(rank),
    log_share = log(pmax(share_global, 1e-12))
  )

growth_long <- purrr::map_dfr(horizons_growth, function(h) {
  endpoint_base %>%
    mutate(
      horizon_weeks = as.integer(h),
      week_tp = week + lubridate::weeks(as.integer(h))
    ) %>%
    left_join(
      endpoint_base %>%
        transmute(endpoint_id, week_tp = week, log_share_tp = log_share),
      by = c("endpoint_id", "week_tp")
    ) %>%
    mutate(dlog_share = log_share_tp - log_share) %>%
    filter(is.finite(dlog_share))
})

growth_summ <- growth_long %>%
  group_by(bucket, horizon_weeks) %>%
  summarise(
    n = n(),
    med = median(dlog_share, na.rm = TRUE),
    p10 = quantile(dlog_share, 0.10, na.rm = TRUE),
    p90 = quantile(dlog_share, 0.90, na.rm = TRUE),
    .groups = "drop"
  )

growth_summ
```

```{r endpoint-growth-plot}
ggplot(growth_long, aes(x = dlog_share, fill = bucket)) +
  geom_density(alpha = 0.35) +
  facet_wrap(~ horizon_weeks, scales = "free_y") +
  labs(
    title = "Forward log-growth in global share by rank bucket",
    subtitle = "Endpoint-level growth distributions (identity-based), as in v2/v3",
    x = "Delta log(share_global) over horizon",
    y = "Density",
    fill = "Bucket"
  ) +
  theme_bw()
```

## Stylized fact B: Rank-dependent drift and volatility (rank-slot increments)

We work with rank-slot series: for each rank `k`, the sequence of shares of the endpoint
occupying that rank each week. This is the object that rank-based diffusion models target.

```{r rank-panel}
rank_panel <- endpoint_weekly %>%
  group_by(rank) %>%
  arrange(week, .by_group = TRUE) %>%
  mutate(
    share_global = pmax(share_global, 1e-12),
    log_w = log(share_global),
    dlogw = log_w - lag(log_w)
  ) %>%
  ungroup() %>%
  filter(is.finite(dlogw))

gauss_params_raw <- rank_panel %>%
  filter(rank <= K_max) %>%
  group_by(rank) %>%
  summarise(
    mean_dlogw = mean(dlogw, na.rm = TRUE),
    sd_dlogw = sd(dlogw, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  right_join(tibble(rank = 1:K_max), by = "rank") %>%
  arrange(rank) %>%
  mutate(
    mean_dlogw = replace_na(mean_dlogw, 0),
    sd_dlogw = replace_na(sd_dlogw, 0),
    sd_dlogw = if_else(sd_dlogw <= 0 | !is.finite(sd_dlogw), 1e-6, sd_dlogw)
  )

ggplot(gauss_params_raw, aes(rank, sd_dlogw)) +
  geom_line() +
  scale_x_continuous(trans = "log10") +
  labs(
    title = "Rank-dependent volatility proxy",
    subtitle = "sd(d log w) by rank (rank-slot series, global shares)",
    x = "Rank (log scale)",
    y = "sd(d log w)"
  ) +
  theme_bw()
```

### Smoothed per-rank moments (v9+)

```{r smoothing-helpers}
moving_average_rank <- function(x, h) {
  stopifnot(h >= 0L)
  n <- length(x)
  if (h == 0L) return(x)
  cs <- c(0, cumsum(replace_na(x, 0)))
  out <- numeric(n)
  for (r in seq_len(n)) {
    lo <- max(1L, r - h)
    hi <- min(n, r + h)
    out[r] <- (cs[hi + 1L] - cs[lo]) / (hi - lo + 1L)
  }
  out
}

build_gaussian_params_smoothed <- function(h) {
  m <- gauss_params_raw$mean_dlogw
  s <- gauss_params_raw$sd_dlogw
  tibble(
    rank = 1:K_max,
    mean_dlogw_s = moving_average_rank(m, h = h),
    sd_dlogw_s = pmax(moving_average_rank(s, h = h), 1e-6)
  )
}

sm_params <- build_gaussian_params_smoothed(h = as.integer(Smoothing_h))

sm_plot <- sm_params %>%
  select(rank, mean_dlogw_s, sd_dlogw_s) %>%
  left_join(gauss_params_raw %>% select(rank, mean_dlogw, sd_dlogw), by = "rank") %>%
  pivot_longer(
    cols = c(mean_dlogw, mean_dlogw_s, sd_dlogw, sd_dlogw_s),
    names_to = "series",
    values_to = "value"
  ) %>%
  mutate(kind = case_when(
    str_detect(series, "^mean") ~ "mean",
    TRUE ~ "sd"
  ))

ggplot(filter(sm_plot, kind == "mean"), aes(rank, value, color = series)) +
  geom_line() +
  scale_x_continuous(trans = "log10") +
  labs(
    title = "Per-rank drift proxy: raw vs smoothed",
    x = "Rank (log scale)",
    y = "mean(d log w)"
  ) +
  theme_bw() +
  guides(color = "none")

ggplot(filter(sm_plot, kind == "sd"), aes(rank, value, color = series)) +
  geom_line() +
  scale_x_continuous(trans = "log10") +
  labs(
    title = "Per-rank volatility proxy: raw vs smoothed",
    x = "Rank (log scale)",
    y = "sd(d log w)"
  ) +
  theme_bw() +
  guides(color = "none")
```

## Stylized fact C: Durable change targets (v7–v10)

These targets summarize the typical magnitude of change over horizons (default 4 and 8 weeks),
separately by rank buckets.

```{r durable-targets}
bucket_def <- list(
  breaks = params$bucket_breaks,
  labels = params$bucket_labels
)

assign_bucket <- function(rank, bucket_def) {
  cut(
    rank,
    breaks = bucket_def$breaks,
    labels = bucket_def$labels,
    include.lowest = TRUE,
    right = TRUE
  )
}

compute_emp_endpoint_change <- function(df, horizons, bucket_def) {
  df2 <- df %>%
    group_by(rank) %>%
    arrange(week, .by_group = TRUE) %>%
    mutate(log_w = log(pmax(share_topK, 1e-12)))
  purrr::map_dfr(horizons, function(h) {
    df2 %>%
      mutate(d = dplyr::lead(log_w, h) - log_w) %>%
      ungroup() %>%
      mutate(
        horizon = as.integer(h),
        bucket = assign_bucket(rank, bucket_def)
      ) %>%
      filter(!is.na(d), !is.na(bucket)) %>%
      group_by(bucket, horizon) %>%
      summarise(
        med_abs_emp = median(abs(d), na.rm = TRUE),
        .groups = "drop"
      )
  })
}

emp_targets <- compute_emp_endpoint_change(
  df = endpoint_weekly %>% filter(rank <= K_cut),
  horizons = horizons_durable,
  bucket_def = bucket_def
)

ggplot(emp_targets, aes(x = factor(horizon), y = med_abs_emp, fill = bucket)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  labs(
    title = "Empirical durable change targets",
    x = "Horizon (weeks ahead)",
    y = "median(|Delta log share_topK|)",
    fill = "Bucket"
  ) +
  theme_bw()
```

## Martingale-style diagnostic (v7)

A simple one-step regression: `dlogw(t+1)` on `dlogw(t)` for rank-slot series, summarized by bucket.

```{r martingale-diagnostic}
diag_df <- rank_panel %>%
  group_by(rank) %>%
  arrange(week, .by_group = TRUE) %>%
  mutate(
    dlogw_t = dlogw,
    dlogw_t1 = lead(dlogw_t, 1)
  ) %>%
  ungroup() %>%
  mutate(bucket = assign_bucket(rank, bucket_def)) %>%
  filter(is.finite(dlogw_t), is.finite(dlogw_t1), !is.na(bucket))

martingale_results <- diag_df %>%
  group_by(bucket) %>%
  group_modify(~ {
    fit <- lm(dlogw_t1 ~ dlogw_t, data = .x)
    broom::tidy(fit)
  }) %>%
  ungroup()

ggplot(martingale_results %>% filter(term == "dlogw_t"), aes(bucket, estimate)) +
  geom_point() +
  geom_errorbar(
    aes(ymin = estimate - 1.96 * std.error, ymax = estimate + 1.96 * std.error),
    width = 0.1
  ) +
  geom_hline(yintercept = 0, linetype = 3) +
  labs(
    title = "Martingale diagnostic: slope of dlogw(t+1) on dlogw(t)",
    x = "Rank bucket",
    y = "Beta estimate"
  ) +
  theme_bw()
```

## Stylized fact D: Gap structure via Xi_k(t) = log(w_k / w_{k+1}) (v11)

We compute Xi using *global* shares, with ranks recomputed within each week.

```{r xi-weekly}
xi_weekly <- endpoint_weekly %>%
  group_by(week) %>%
  arrange(rank, .by_group = TRUE) %>%
  mutate(
    w = pmax(share_global, 1e-18),
    w_next = lead(w),
    xi = log(w / w_next),
    k = rank
  ) %>%
  ungroup() %>%
  filter(is.finite(xi), is.finite(k))

K_xi_est <- min(2000L, K_cut)

# scatter (restricted for clarity)
ggplot(filter(xi_weekly, k <= 500L), aes(x = k, y = xi)) +
  geom_point(alpha = 0.35, size = 0.4) +
  scale_x_continuous(trans = "log10") +
  labs(
    title = "Log gaps (Xi) by rank: Xi_k(t) = log(w_k / w_{k+1})",
    subtitle = "Restricted to top 500 ranks for visibility",
    x = "Rank k (log scale)",
    y = "Xi"
  ) +
  theme_bw()

rho_by_k_med <- xi_weekly %>%
  filter(k <= (K_xi_est - 1L)) %>%
  group_by(k) %>%
  summarise(rho_k = median(xi, na.rm = TRUE), .groups = "drop")

ggplot(rho_by_k_med, aes(k, rho_k)) +
  geom_line() +
  theme_bw() +
  labs(
    title = "Median Xi by rank (rho_k)",
    x = "Rank k",
    y = "rho_k"
  )
```

### v11-style sigma_hat(k) construction

Following v11:

- Let `g_k` be smoothed per-rank drift proxy (mean of `d log w`) for ranks.
- Define `S_k = cumsum(g_k)`.
- Combine with `rho_k` to form `sigma_hat(k) = sqrt( -2 S_k rho_k )` when positive.

```{r sigma-hat}
g_by_k <- sm_params %>%
  transmute(k = rank, g_k = mean_dlogw_s) %>%
  filter(k <= (K_xi_est + 1L)) %>%
  arrange(k)

S_by_k <- g_by_k %>%
  mutate(S_k = cumsum(g_k))

sigma_hat_by_k <- rho_by_k_med %>%
  inner_join(S_by_k %>% select(k, S_k), by = "k") %>%
  mutate(
    sigma2_hat = -2 * S_k * rho_k,
    sigma_hat = if_else(sigma2_hat > 0, sqrt(sigma2_hat), NA_real_)
  )

sigma_hat_by_k %>%
  summarise(med_sigma = median(sigma_hat, na.rm = TRUE))

ggplot(
  filter(sigma_hat_by_k, k >= 10L, !is.na(sigma_hat)),
  aes(x = k, y = sigma_hat)
) +
  geom_line() +
  scale_x_continuous(trans = "log10") +
  labs(
    title = "sigma_hat(k) vs rank (v11-style)",
    x = "Rank k (log scale)",
    y = "sigma_hat(k)"
  ) +
  theme_bw()
```

### Stability across time (quarterly summary)

```{r sigma-hat-over-time}
xi_by_period <- xi_weekly %>%
  mutate(period = floor_date(week, params$bootstrap_period_unit)) %>%
  filter(k <= (K_xi_est - 1L))

rho_period_by_k <- xi_by_period %>%
  group_by(period, k) %>%
  summarise(rho_k = median(xi, na.rm = TRUE), .groups = "drop")

sigma_hat_period <- rho_period_by_k %>%
  inner_join(S_by_k %>% select(k, S_k), by = "k") %>%
  mutate(
    sigma2_hat = -2 * S_k * rho_k,
    sigma_hat = if_else(sigma2_hat > 0, sqrt(sigma2_hat), NA_real_)
  )

sigma_hat_period_summ <- sigma_hat_period %>%
  group_by(period) %>%
  summarise(
    sigma_med = median(sigma_hat, na.rm = TRUE),
    sigma_p25 = quantile(sigma_hat, 0.25, na.rm = TRUE),
    sigma_p75 = quantile(sigma_hat, 0.75, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(sigma_hat_period_summ, aes(x = period, y = sigma_med)) +
  geom_line() +
  geom_ribbon(aes(ymin = sigma_p25, ymax = sigma_p75), alpha = 0.2) +
  labs(
    title = "sigma_hat over time (median across ranks) with IQR band",
    x = NULL,
    y = "sigma_hat (median)"
  ) +
  theme_bw()
```

### Bootstrap CI over time (v11-style)

```{r bootstrap-ci}
bootstrap_sigma_period <- function(
  xi_df_period,
  S_by_k,
  B = 200,
  k_min = 10L,
  K_xi_est = 2000L
) {
  weeks <- unique(xi_df_period$week)
  boot_stats <- numeric(B)

  for (b in seq_len(B)) {
    w_samp <- sample(weeks, size = length(weeks), replace = TRUE)

    rho_b <- xi_df_period %>%
      filter(week %in% w_samp, k <= (K_xi_est - 1L)) %>%
      group_by(k) %>%
      summarise(rho_k = median(xi, na.rm = TRUE), .groups = "drop")

    sig_b <- rho_b %>%
      inner_join(S_by_k %>% select(k, S_k), by = "k") %>%
      mutate(
        sigma2_hat = -2 * S_k * rho_k,
        sigma_hat = if_else(sigma2_hat > 0, sqrt(sigma2_hat), NA_real_)
      ) %>%
      filter(k >= k_min, k <= (K_xi_est - 1L))

    boot_stats[b] <- median(sig_b$sigma_hat, na.rm = TRUE)
  }

  tibble(
    sigma_hat_med = median(boot_stats, na.rm = TRUE),
    ci_lo = quantile(boot_stats, 0.025, na.rm = TRUE),
    ci_hi = quantile(boot_stats, 0.975, na.rm = TRUE)
  )
}

bootstrap_ci_deps <- list(
  cache_version = params$cache_version %||% "v1",
  raw_mtime = deps_file_mtime(fb_weekly_parquet_path),
  bootstrap_period_unit = params$bootstrap_period_unit,
  bootstrap_B = as.integer(bootstrap_B),
  bootstrap_k_min = as.integer(bootstrap_k_min),
  K_xi_est = K_xi_est,
  S_by_k_fp = digest::digest(S_by_k$S_k, algo = "xxhash64"),
  code = deps_code_mtime(here::here("FB_weekly_master_analysis.Rmd"))
)

sigma_ci_by_period <- cache_or_compute(
  name = "sigma_ci_by_period_bootstrap_master",
  compute_fn = function() {
    withr::with_seed(params$seed, {
      xi_by_period %>%
        group_by(period) %>%
        group_modify(~ bootstrap_sigma_period(
          xi_df_period = .x,
          S_by_k = S_by_k,
          B = as.integer(bootstrap_B),
          k_min = as.integer(bootstrap_k_min),
          K_xi_est = K_xi_est
        )) %>%
        ungroup()
    })
  },
  deps = bootstrap_ci_deps,
  cfg = CFG
)

ggplot(sigma_ci_by_period, aes(x = period, y = sigma_hat_med)) +
  geom_line() +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), alpha = 0.2) +
  labs(
    title = "Bootstrap CI for sigma_hat over time",
    x = NULL,
    y = "sigma_hat (median across k)"
  ) +
  theme_bw()
```

## Stylized fact C (expanded): Turnover and rank mobility (new)

This section adds rank-mobility summaries that were not fully developed in the PDFs.

### 9.1 Top-K membership turnover (set overlap)

```{r topk-turnover}
topK_overlap <- function(df, K, horizons = c(1L, 4L, 12L)) {
  K <- as.integer(K)

  df_topK <- df %>%
    filter(rank <= K) %>%
    select(week, endpoint_id) %>%
    distinct()

  week_vec <- sort(unique(df_topK$week))

  purrr::map_dfr(as.integer(horizons), function(h) {
    base_weeks <- week_vec[week_vec + lubridate::weeks(h) <= max(week_vec)]

    purrr::map_dfr(base_weeks, function(w) {
      w_tp <- w + lubridate::weeks(h)

      a <- df_topK %>% filter(week == w) %>% pull(endpoint_id)
      b <- df_topK %>% filter(week == w_tp) %>% pull(endpoint_id)

      # guard against weeks where the top-K list is shorter than K
      if (length(a) < K || length(b) < K) {
        return(tibble(K = K, horizon = h, week = w, overlap_frac = NA_real_))
      }

      tibble(
        K = K,
        horizon = h,
        week = w,
        overlap_frac = length(intersect(a, b)) / K
      )
    })
  })
}

turnover_tbl <- purrr::map_dfr(
  turnover_K_list,
  ~ topK_overlap(endpoint_weekly, K = .x, horizons = turnover_horizons)
)

ggplot(turnover_tbl, aes(x = week, y = overlap_frac, color = factor(horizon))) +
  geom_line(alpha = 0.6) +
  facet_wrap(~ K, scales = "free_y") +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(
    title = "Top-K overlap over time",
    subtitle = "Fraction of endpoints in top-K at week t that also appear in top-K at t+h",
    x = "Week",
    y = "Overlap fraction",
    color = "Horizon (weeks)"
  ) +
  theme_bw()
```

### 9.2 Rank-conditioned mobility: where does the endpoint at rank k go?

This extends the v12 micro-object by recording future rank as well as future share.

```{r micro-emp-extended}
emp_micro_for_k <- function(endpoint_weekly, k, horizons = c(1L, 4L, 12L)) {
  base <- endpoint_weekly %>%
    filter(rank == k) %>%
    transmute(
      week_t = week,
      endpoint_id,
      share_t = share_topK
    )

  fut <- endpoint_weekly %>%
    transmute(
      week_tp = week,
      endpoint_id,
      rank_tp = rank,
      share_tp = share_topK
    )

  purrr::map_dfr(horizons, function(h) {
    base %>%
      mutate(week_tp = week_t + weeks(h), horizon = as.integer(h)) %>%
      left_join(fut, by = c("week_tp", "endpoint_id")) %>%
      mutate(
        present_tp = !is.na(share_tp),
        rank_change = if_else(present_tp, rank_tp - k, NA_integer_)
      )
  })
}

k_grid_mobility <- c(1L, 5L, 12L, 25L, 100L, 500L, 1000L)

emp_micro_mobility <- purrr::map_dfr(
  k_grid_mobility,
  ~ emp_micro_for_k(endpoint_weekly, k = .x, horizons = horizons_micro) %>%
    mutate(k = .x)
)

mobility_summ <- emp_micro_mobility %>%
  group_by(k, horizon) %>%
  summarise(
    retention = mean(present_tp),
    p_within_1 = mean(abs(rank_change) <= 1, na.rm = TRUE),
    p_within_5 = mean(abs(rank_change) <= 5, na.rm = TRUE),
    med_abs_rank_change = median(abs(rank_change), na.rm = TRUE),
    .groups = "drop"
)

mobility_summ
```

```{r mobility-plots}
ggplot(mobility_summ, aes(x = horizon, y = retention, group = factor(k))) +
  geom_line() +
  geom_point(size = 1.5) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_x_continuous(breaks = horizons_micro) +
  labs(
    title = "Retention by horizon for rank-conditioned cohorts",
    x = "Horizon (weeks ahead)",
    y = "Fraction still observed",
    caption = "Cohort: endpoint that sits at rank k at week t"
  ) +
  theme_bw()

ggplot(mobility_summ, aes(x = horizon, y = med_abs_rank_change, group = factor(k))) +
  geom_line() +
  geom_point(size = 1.5) +
  scale_x_continuous(breaks = horizons_micro) +
  labs(
    title = "Median |rank change| for rank-conditioned cohorts",
    x = "Horizon (weeks ahead)",
    y = "Median absolute rank change"
  ) +
  theme_bw()
```

### 9.3 Rank transition matrix (deciles + out-of-topK)

```{r transition-matrix}
make_rank_bin <- function(rank, K, n_bins = 10L) {
  case_when(
    is.na(rank) ~ "out",
    rank > K ~ "out",
    TRUE ~ paste0("D", pmin(n_bins, pmax(1L, ceiling(n_bins * rank / K))))
  )
}

rank_bins <- endpoint_weekly %>%
  select(week, endpoint_id, rank) %>%
  mutate(bin = make_rank_bin(rank, K = 1000L, n_bins = 10L))

# 1-week transition (calendar weeks; missing at t+1 treated as out)
trans_1w <- rank_bins %>%
  transmute(
    endpoint_id,
    week_t = week,
    bin_t = bin,
    week_tp = week + lubridate::weeks(1L)
  ) %>%
  left_join(
    rank_bins %>% transmute(endpoint_id, week_tp = week, bin_tp = bin),
    by = c("endpoint_id", "week_tp")
  ) %>%
  mutate(bin_tp = replace_na(bin_tp, "out")) %>%
  count(bin_t, bin_tp) %>%
  group_by(bin_t) %>%
  mutate(p = n / sum(n)) %>%
  ungroup()

ggplot(trans_1w, aes(x = bin_tp, y = bin_t, fill = p)) +
  geom_tile() +
  scale_fill_continuous(labels = percent_format(accuracy = 0.1)) +
  labs(
    title = "Rank-bin transition matrix (1 week)",
    subtitle = "Bins are deciles within top 1000 ranks; 'out' means rank>1000 or missing",
    x = "Bin at t+1",
    y = "Bin at t"
  ) +
  theme_bw()
```

## Stylized fact E: Common shocks vs idiosyncratic movement (new)

A pragmatic diagnostic: PCA on rank-slot increments for the top `pca_K` ranks.

```{r common-shocks-pca}

K_pca <- min(pca_K, K_cut)

pca_cache <- list(
  beta_k = rep(0, K_max),
  pc1_var_explained = NA_real_,
  pc1 = NULL
)

if (is.na(K_pca) || K_pca < 2) {
  message("Skipping PCA: K_pca < 2 (K_pca=", K_pca, ").")
} else {
  rank_slot_wide <- rank_panel %>%
    filter(rank <= K_pca) %>%
    select(week, rank, dlogw) %>%
    pivot_wider(names_from = rank, values_from = dlogw)

  # keep complete weeks for PCA (top ranks usually complete)
  rank_slot_mat <- rank_slot_wide %>%
    drop_na() %>%
    arrange(week)

  X <- rank_slot_mat %>%
    select(-week) %>%
    as.matrix()

  if (nrow(X) < 2 || ncol(X) < 2) {
    message(
      "Skipping PCA: need at least 2 complete weeks and 2 rank columns; got ",
      nrow(X), "x", ncol(X), "."
    )
  } else {
    pca_fit <- prcomp(X, center = TRUE, scale. = TRUE)

    var_explained <- (pca_fit$sdev^2) / sum(pca_fit$sdev^2)

    tibble(
      component = seq_along(var_explained),
      var_explained = var_explained
    ) %>%
      slice(1:10)

    ggplot(
      tibble(component = seq_along(var_explained), var_explained = var_explained) %>% slice(1:20),
      aes(component, var_explained)
    ) +
      geom_line() +
      geom_point(size = 1.5) +
      scale_y_continuous(labels = percent_format(accuracy = 0.1)) +
      labs(
        title = "PCA on rank-slot increments: variance explained",
        subtitle = paste0("Top ", K_pca, " ranks; complete weeks only"),
        x = "Principal component",
        y = "Share of variance explained"
      ) +
      theme_bw()

    # first PC time series (common shock proxy)
    pc1 <- pca_fit$x[, 1]
    pc1_df <- tibble(week = rank_slot_mat$week, pc1 = pc1)

    ggplot(pc1_df, aes(week, pc1)) +
      geom_line() +
      labs(
        title = "First principal component over time (common shock proxy)",
        x = "Week",
        y = "PC1 score"
      ) +
      theme_bw()

    # cache loadings for jump-factor model (smoothed, padded to K_max)
    beta_raw <- pca_fit$rotation[, 1]
    beta_vec <- rep(0, K_max)
    beta_vec[seq_len(K_pca)] <- beta_raw
    beta_vec <- moving_average_rank(beta_vec, h = as.integer(Smoothing_h))
    beta_scale <- sqrt(mean(beta_vec^2, na.rm = TRUE))
    if (is.finite(beta_scale) && beta_scale > 0) {
      beta_vec <- beta_vec / beta_scale
    }

    pca_cache$beta_k <- beta_vec
    pca_cache$pc1_var_explained <- var_explained[1]
    pca_cache$pc1 <- pc1_df
  }
}
```

## Heavy-tail / jump diagnostics for increments (new)

This section checks whether rank-slot increments have tails that are substantially heavier than Gaussian.

```{r tail-diagnostics}
build_bucket_z <- function(increments_df, bucket_def) {
  increments_df %>%
    mutate(bucket = assign_bucket(rank, bucket_def)) %>%
    filter(!is.na(bucket)) %>%
    group_by(bucket) %>%
    mutate(
      z = (dlogw - mean(dlogw, na.rm = TRUE)) / sd(dlogw, na.rm = TRUE)
    ) %>%
    ungroup()
}

tail_diagnostics <- function(increments_df, bucket_def) {
  inc_by_bucket <- build_bucket_z(increments_df, bucket_def)
  inc_by_bucket %>%
    group_by(bucket) %>%
    summarise(
      n = n(),
      p_gt_5sd = mean(abs(z) >= 5, na.rm = TRUE),
      p_gt_3sd = mean(abs(z) >= 3, na.rm = TRUE),
      q_0.001 = quantile(z, 0.001, na.rm = TRUE),
      q_0.999 = quantile(z, 0.999, na.rm = TRUE),
      .groups = "drop"
    )
}

# Build standardized increments by bucket
inc_by_bucket <- build_bucket_z(
  rank_panel %>% filter(rank <= K_cut) %>% select(rank, dlogw),
  bucket_def
)

tail_summ <- tail_diagnostics(
  rank_panel %>% filter(rank <= K_cut) %>% select(rank, dlogw),
  bucket_def
)

tail_summ
```

```{r qq-plot}
ggplot(inc_by_bucket, aes(sample = z)) +
  stat_qq() +
  stat_qq_line() +
  facet_wrap(~ bucket, scales = "free") +
  labs(
    title = "QQ plot of standardized increments by bucket",
    subtitle = "Visual check for heavy tails vs Gaussian"
  ) +
  theme_bw()
```

## Simulation framework (v8–v11) + unified fit metrics

### 12.1 Empirical entrant pool for entry/exit

```{r entrant-pool}
entrant_pool <- endpoint_weekly %>%
  filter(rank > K_cut, rank <= min(K_cut + K_tail_buffer, max_rank_seen)) %>%
  transmute(share = share_global) %>%
  filter(is.finite(share), share > 0)

stopifnot(nrow(entrant_pool) > 100)

entrant_sampler <- function(n) {
  sample(entrant_pool$share, size = n, replace = TRUE)
}

# baseline seed curve w0 up to K_max
mean_share <- endpoint_weekly %>%
  group_by(rank) %>%
  summarise(w_bar = mean(share_global, na.rm = TRUE), .groups = "drop") %>%
  arrange(rank)

w0_raw <- mean_share$w_bar[seq_len(min(K_max, nrow(mean_share)))]
w0_raw <- w0_raw / sum(w0_raw)

# extend once to full universe length K_max
if (length(w0_raw) < K_max) {
  w0_ext <- c(w0_raw, entrant_sampler(K_max - length(w0_raw)))
  w0_ext <- w0_ext / sum(w0_ext)
} else {
  w0_ext <- w0_raw[1:K_max]
  w0_ext <- w0_ext / sum(w0_ext)
}
```

### 12.2 Simulator (Gaussian + heavy-tail/jump innovations; optional Xi output)

This merges v9’s simulator with v11’s Xi storage and abstracts the innovation draw so we can plug in heavy-tail and jump models.

```{r simulator}

`%||%` <- function(x, y) if (!is.null(x)) x else y

clamp01 <- function(x) pmin(pmax(x, 0), 1)

rstd_t <- function(n, df) {
  if (df <= 2) stop("df must be > 2 for unit-variance t")
  rt(n, df) * sqrt((df - 2) / df)
}

rstd_t_vec <- function(df_vec) {
  out <- numeric(length(df_vec))
  for (d in unique(df_vec)) {
    idx <- which(df_vec == d)
    out[idx] <- rstd_t(length(idx), d)
  }
  out
}

rlaplace <- function(n, scale) {
  scale <- as.numeric(scale)
  if (!is.finite(scale) || scale <= 0) stop("scale must be > 0")
  sign <- sample(c(-1, 1), size = n, replace = TRUE)
  sign * rexp(n, rate = 1 / scale)
}

resolve_bucket_param <- function(param, K_max, bucket_def = NULL, default = NULL) {
  if (is.null(param)) return(rep(default, K_max))
  if (length(param) == 1L) return(rep(param, K_max))
  if (length(param) == K_max) return(param)
  if (is.null(bucket_def) || is.null(bucket_def$breaks)) return(rep(default, K_max))

  if (!is.null(names(param))) {
    buckets <- assign_bucket(seq_len(K_max), bucket_def)
    out <- rep(default, K_max)
    for (b in levels(buckets)) {
      if (b %in% names(param)) out[buckets == b] <- param[[b]]
    }
    return(out)
  }
  if (!is.null(bucket_def$labels) && length(param) == length(bucket_def$labels)) {
    buckets <- assign_bucket(seq_len(K_max), bucket_def)
    out <- rep(default, K_max)
    for (i in seq_along(bucket_def$labels)) {
      out[buckets == bucket_def$labels[i]] <- param[i]
    }
    return(out)
  }
  rep(default, K_max)
}

resolve_beta_k <- function(model, cache, K_max) {
  if (!is.null(model$beta_k)) {
    if (length(model$beta_k) >= K_max) return(model$beta_k[seq_len(K_max)])
    return(c(model$beta_k, rep(0, K_max - length(model$beta_k))))
  }
  if (!is.null(cache$beta_k) && length(cache$beta_k) == K_max) return(cache$beta_k)
  rep(0, K_max)
}

resolve_sys_loading <- function(model, beta_k, sd_vec) {
  if (is.character(model$sys_loading)) {
    if (model$sys_loading == "beta") return(beta_k)
    if (model$sys_loading == "sigma") {
      s <- sd_vec / mean(sd_vec)
      return(s)
    }
  }
  if (is.numeric(model$sys_loading) && length(model$sys_loading) == length(sd_vec)) {
    return(model$sys_loading)
  }
  sd_vec / mean(sd_vec)
}

draw_jump_sizes <- function(n, spec) {
  if (n <= 0) return(numeric(0))
  dist <- spec$dist %||% "laplace"
  if (dist == "laplace") {
    scale <- spec$scale %||% 0.1
    return(rlaplace(n, scale = scale))
  }
  if (dist == "t") {
    df <- spec$df %||% 5
    scale <- spec$scale %||% 0.1
    return(rstd_t(n, df = df) * scale)
  }
  stop("Unknown jump dist: ", dist)
}

draw_increments <- function(t, w_t, moment_curves, model, cache) {
  type <- model$type %||% "gaussian"
  mean_vec <- moment_curves$mean_vec
  sd_vec <- moment_curves$sd_vec
  K_max <- length(mean_vec)

  mu <- model$mu %||% 0
  sigma <- model$sigma %||% 1

  if (type == "gaussian") {
    return(rnorm(K_max, mean = mu + mean_vec, sd = sigma * sd_vec))
  }

  if (type %in% c("t", "student_t")) {
    df <- model$df %||% 6
    z <- rstd_t(K_max, df = df)
    return(mu + mean_vec + sigma * sd_vec * z)
  }

  if (type == "jump_factor") {
    df_F <- model$df_F %||% 6
    phi_F <- model$phi_F %||% 0

    state <- cache$state
    if (is.null(state$F_prev)) state$F_prev <- 0

    if (abs(phi_F) >= 1) phi_F <- sign(phi_F) * 0.99
    eta <- rstd_t(1, df = df_F)
    F_t <- phi_F * state$F_prev + sqrt(1 - phi_F^2) * eta
    state$F_prev <- F_t

    beta_k <- resolve_beta_k(model, cache, K_max)

    df_eps_vec <- resolve_bucket_param(
      param = model$df_eps %||% 8,
      K_max = K_max,
      bucket_def = moment_curves$bucket_def,
      default = model$df_eps %||% 8
    )
    df_eps_vec <- pmax(df_eps_vec, 2.01)
    eps <- rstd_t_vec(df_eps_vec)

    p_sys <- clamp01(model$p_sys %||% 0)
    I_sys <- rbinom(1, 1, p_sys)
    J_sys <- rep(0, K_max)
    if (I_sys == 1) {
      S_sys <- draw_jump_sizes(1, spec = model$jump_sys %||% list(
        dist = model$jump_dist %||% "laplace",
        scale = model$jump_scale_sys %||% model$jump_scale %||% 0.1,
        df = model$jump_df_sys %||% model$jump_df
      ))
      a_k <- resolve_sys_loading(model, beta_k, sd_vec)
      J_sys <- a_k * S_sys
    }

    p_idio_vec <- resolve_bucket_param(
      param = model$p_idio %||% 0,
      K_max = K_max,
      bucket_def = moment_curves$bucket_def,
      default = 0
    )
    p_idio_vec <- clamp01(p_idio_vec)
    I_idio <- rbinom(K_max, 1, p_idio_vec)
    J_idio <- rep(0, K_max)
    if (any(I_idio == 1)) {
      S_idio <- draw_jump_sizes(sum(I_idio == 1), spec = model$jump_idio %||% list(
        dist = model$jump_dist %||% "laplace",
        scale = model$jump_scale_idio %||% model$jump_scale %||% 0.05,
        df = model$jump_df_idio %||% model$jump_df
      ))
      J_idio[I_idio == 1] <- S_idio
    }

    return(mu + mean_vec + beta_k * F_t + sigma * sd_vec * eps + J_sys + J_idio)
  }

  stop("Unknown model type: ", type)
}



sample_increments_for_tail <- function(
  model, moment_curves, cache,
  K_cut,
  n_steps = 200L,
  n_paths = 30L,
  ranks_per_bucket = 200L
) {
  K_max <- length(moment_curves$mean_vec)
  ranks <- seq_len(min(K_cut, K_max))

  bucket_def <- moment_curves$bucket_def %||%
    (if (exists("bucket_def")) bucket_def else NULL)
  if (is.null(bucket_def)) stop("bucket_def is required for tail sampling")

  buckets <- assign_bucket(ranks, bucket_def)
  keep <- !is.na(buckets)
  ranks <- ranks[keep]
  buckets <- buckets[keep]

  if (is.null(cache)) cache <- list()
  if (is.null(cache$state)) cache$state <- new.env(parent = emptyenv())

  use_mean_reversion <- isTRUE(model$mean_reversion)
  if (use_mean_reversion) {
    kappa_vec <- moment_curves$kappa_vec
    if (is.null(kappa_vec)) {
      kappa_vec <- resolve_bucket_param(
        param = model$kappa %||% 0,
        K_max = K_max,
        bucket_def = bucket_def,
        default = 0
      )
    }
  }

  res <- vector("list", n_paths * n_steps)
  idx <- 1

  for (p in seq_len(n_paths)) {
    cache$state$F_prev <- 0
    if (use_mean_reversion) y <- rep(0, K_max)

    for (t in seq_len(n_steps)) {
      dlogw_innov <- draw_increments(
        t = t,
        w_t = rep(1 / K_max, K_max),
        moment_curves = moment_curves,
        model = model,
        cache = cache
      )

      if (use_mean_reversion) {
        dlogw <- dlogw_innov - kappa_vec * y
        y <- (1 - kappa_vec) * y + dlogw_innov
      } else {
        dlogw <- dlogw_innov
      }

      dlogw <- dlogw[ranks]

      df <- tibble(rank = ranks, dlogw = dlogw, bucket = buckets)
      if (!is.null(ranks_per_bucket) && is.finite(ranks_per_bucket)) {
        df <- df %>%
          group_by(bucket) %>%
          slice_sample(n = min(as.integer(ranks_per_bucket), n())) %>%
          ungroup()
      }
      res[[idx]] <- df
      idx <- idx + 1
    }
  }

  bind_rows(res)
}
simulate_rank_paths <- function(
  w0, K_cut, K_max, T = 52, n_paths = 200,
  mu = 0, sigma = 1, entry_frac = 0.10,
  mean_vec, sd_vec,
  horizons = c(4L, 8L),
  K_xi = NULL,
  model = list(type = "gaussian"),
  moment_curves = NULL,
  cache = NULL
) {
  stopifnot(length(mean_vec) == K_max, length(sd_vec) == K_max)

  top_n_sorted <- function(w, n) w[order(w, decreasing = TRUE)][seq_len(n)]

  horizons <- as.integer(horizons)
  stopifnot(all(horizons >= 1L), all(horizons <= T))

  if (!is.null(K_xi)) {
    K_xi <- as.integer(K_xi)
    stopifnot(K_xi >= 2L, K_xi <= K_cut)
  }

  if (is.null(moment_curves)) {
    moment_curves <- list(
      mean_vec = mean_vec,
      sd_vec = sd_vec,
      bucket_def = if (exists("bucket_def")) bucket_def else NULL
    )
  } else {
    moment_curves$mean_vec <- mean_vec
    moment_curves$sd_vec <- sd_vec
    if (is.null(moment_curves$bucket_def) && exists("bucket_def")) {
      moment_curves$bucket_def <- bucket_def
    }
  }

  if (is.null(cache)) cache <- list()
  if (is.null(cache$state)) cache$state <- new.env(parent = emptyenv())

  if (is.null(model$mu)) model$mu <- mu
  if (is.null(model$sigma)) model$sigma <- sigma

  # baseline for mean reversion (if enabled)
  if (isTRUE(model$mean_reversion)) {
    if (is.null(moment_curves$log_wbar)) {
      w_base <- if (length(w0) < K_max) {
        w0_ext <- c(w0, entrant_sampler(K_max - length(w0)))
        w0_ext / sum(w0_ext)
      } else {
        w0[1:K_max] / sum(w0[1:K_max])
      }
      moment_curves$log_wbar <- log(pmax(w_base, 1e-18))
    }
    if (is.null(moment_curves$kappa_vec)) {
      moment_curves$kappa_vec <- resolve_bucket_param(
        param = model$kappa %||% 0,
        K_max = K_max,
        bucket_def = moment_curves$bucket_def,
        default = 0
      )
    }
  }

  snapshots <- vector("list", n_paths)
  growth_collect <- setNames(vector("list", length(horizons)), paste0("h", horizons))
  xi_collect <- vector("list", n_paths)

  for (p in seq_len(n_paths)) {

    # reset common-factor state per path
    cache$state$F_prev <- 0

    # initialize to full universe length K_max
    if (length(w0) < K_max) {
      w <- c(w0, entrant_sampler(K_max - length(w0)))
      w <- w / sum(w)
    } else {
      w <- w0[1:K_max]
      w <- w / sum(w)
    }

    # baseline top-K for durable change (within-top-K normalization)
    w_baseK <- top_n_sorted(w, K_cut)
    w_baseK <- w_baseK / sum(w_baseK)

    snaps_p <- vector("list", T)
    xi_p <- vector("list", T)

    for (t in seq_len(T)) {

      dlogw <- draw_increments(t = t, w_t = w, moment_curves = moment_curves,
                               model = model, cache = cache)

      if (isTRUE(model$mean_reversion)) {
        log_wbar <- moment_curves$log_wbar
        kappa_vec <- moment_curves$kappa_vec %||% rep(0, K_max)
        y <- log(pmax(w, 1e-18)) - log_wbar
        y_new <- (1 - kappa_vec) * y + dlogw
        w <- exp(log_wbar + y_new)
      } else {
        w <- w * exp(dlogw)
      }

      w <- pmax(w, 1e-18)
      w <- w / sum(w)

      # entry/exit: keep top-K_cut, refill tail to K_max
      w_top <- top_n_sorted(w, K_cut)          # global shares of current top-K
      n_tail <- K_max - K_cut

      if (n_tail > 0) {
        n_add <- max(1L, floor(n_tail * entry_frac))
        w <- c(w_top, entrant_sampler(n_add), entrant_sampler(n_tail - n_add))
      } else {
        w <- w_top
      }

      w <- w / sum(w)

      # snapshot (top-K, within-top-K normalization)
      wK <- top_n_sorted(w, K_cut)
      wK <- wK / sum(wK)

      snaps_p[[t]] <- tibble(
        path = p,
        t = t,
        rank = seq_len(K_cut),
        share = wK
      )

      # Xi (global shares; scale-invariant, but use current top-K global vector)
      if (!is.null(K_xi)) {
        w_top_now <- top_n_sorted(w, K_cut) %>% pmax(1e-18)
        xi_vals <- log(w_top_now[seq_len(K_xi - 1L)] / w_top_now[seq(2L, K_xi)])
        xi_p[[t]] <- tibble(
          path = p,
          t = t,
          k = seq_len(K_xi - 1L),
          xi = xi_vals
        )
      }

      # durable change at selected horizons relative to baseline (top-K, within-top-K)
      if (t %in% horizons) {
        lg <- log(wK / w_baseK)
        lbl <- paste0("h", t)
        growth_collect[[lbl]] <- rbind(
          growth_collect[[lbl]],
          cbind(
            path = p,
            rank = seq_len(K_cut),
            horizon = t,
            log_growth = lg
          )
        )
      }
    }

    snapshots[[p]] <- bind_rows(snaps_p)
    if (!is.null(K_xi)) xi_collect[[p]] <- bind_rows(xi_p)
  }

  growth_df <- bind_rows(lapply(growth_collect, as_tibble))

  out <- list(
    snapshots = bind_rows(snapshots),
    growth = growth_df
  )

  if (!is.null(K_xi)) out$xi <- bind_rows(xi_collect)
  out
}
```

### 12.3 Fit metrics (CDC, durable change, Xi)

```{r fit-metrics}
cdc_rmse <- function(emp, sim) {
  df <- emp %>%
    inner_join(sim, by = "rank", suffix = c("_emp", "_sim"))
  sqrt(mean((log(df$w_bar_emp) - log(df$w_bar_sim))^2, na.rm = TRUE))
}

durable_rmse <- function(emp_targets, sim_growth, bucket_def) {
  sim_targets <- sim_growth %>%
    mutate(bucket = assign_bucket(rank, bucket_def)) %>%
    filter(!is.na(bucket)) %>%
    group_by(bucket, horizon) %>%
    summarise(med_abs_sim = median(abs(log_growth), na.rm = TRUE), .groups = "drop")

  emp_targets %>%
    inner_join(sim_targets, by = c("bucket", "horizon")) %>%
    summarise(rmse = sqrt(mean((med_abs_emp - med_abs_sim)^2, na.rm = TRUE))) %>%
    pull(rmse)
}

xi_rmse <- function(emp_rho_by_k, sim_xi_df) {
  sim_rho <- sim_xi_df %>%
    group_by(k) %>%
    summarise(rho_k = median(xi, na.rm = TRUE), .groups = "drop")
  df <- emp_rho_by_k %>%
    inner_join(sim_rho, by = "k", suffix = c("_emp", "_sim"))
  sqrt(mean((df$rho_k_emp - df$rho_k_sim)^2, na.rm = TRUE))
}
```

## Baseline simulation run + comparison (CDC, durable, Xi)

```{r run-sim}
sm_fixed <- build_gaussian_params_smoothed(h = as.integer(Smoothing_h))
mean_vec_s <- sm_fixed$mean_dlogw_s
sd_vec_s <- sm_fixed$sd_dlogw_s

sim_out <- simulate_rank_paths(
  w0 = w0_ext,
  K_cut = K_cut,
  K_max = K_max,
  T = sim_T_weeks,
  n_paths = sim_n_paths,
  mu = sim_mu,
  sigma = sim_sigma,
  entry_frac = sim_entry_frac,
  mean_vec = mean_vec_s,
  sd_vec = sd_vec_s,
  horizons = horizons_durable,
  K_xi = sim_K_xi
)

# CDC comparison
sim_cdc <- sim_out$snapshots %>%
  group_by(rank) %>%
  summarise(w_bar = mean(share, na.rm = TRUE), .groups = "drop") %>%
  arrange(rank)

bind_rows(
  emp_cdc %>% mutate(source = "Empirical"),
  sim_cdc %>% mutate(source = "Simulated")
) %>%
  ggplot(aes(rank, w_bar, linetype = source)) +
  geom_line() +
  scale_x_log10() +
  scale_y_log10(labels = percent_format(accuracy = 0.01)) +
  labs(
    title = "CDC: Empirical vs Simulated",
    subtitle = paste0("mu=", sim_mu, ", sigma=", sim_sigma,
                      ", h=", Smoothing_h),
    x = "Rank (log scale)",
    y = "Mean share within top-K (log scale)",
    linetype = NULL
  ) +
  theme_bw()

# Scoreboard
score_tbl <- tibble(
  rmse_cdc = cdc_rmse(emp_cdc, sim_cdc),
  rmse_durable = durable_rmse(emp_targets, sim_out$growth, bucket_def),
  rmse_xi = xi_rmse(rho_by_k_med %>% filter(k <= sim_K_xi - 1L),
                    sim_out$xi %>% filter(k <= sim_K_xi - 1L))
)

score_tbl
```

## Jump model simulation (heavy tails; optional)

```{r run-jump-model}
if (isTRUE(params$run_jump_model)) {
  jump_out <- tryCatch({
    moment_curves_jump <- list(
      mean_vec = mean_vec_s,
      sd_vec = sd_vec_s,
      log_wbar = log(pmax(w0_ext, 1e-18)),
      bucket_def = bucket_def
    )

    jump_model_type_i <- ifelse(is.na(jump_model_type) || jump_model_type == "", "jump_factor", jump_model_type)
    jump_dist_i <- ifelse(is.na(jump_dist) || jump_dist == "", "laplace", jump_dist)
    jump_sys_loading_i <- ifelse(is.na(jump_sys_loading) || jump_sys_loading == "", "beta", jump_sys_loading)

    jump_df_F_i <- ifelse(is.na(jump_df_F), 5, jump_df_F)
    jump_phi_F_i <- ifelse(is.na(jump_phi_F), 0, jump_phi_F)

    jump_df_eps_large_i <- ifelse(is.na(jump_df_eps_large), 6, jump_df_eps_large)
    jump_df_eps_midsize_i <- ifelse(is.na(jump_df_eps_midsize), 8, jump_df_eps_midsize)
    jump_df_eps_small_i <- ifelse(is.na(jump_df_eps_small), 10, jump_df_eps_small)

    jump_p_sys_i <- ifelse(is.na(jump_p_sys), 0.02, jump_p_sys)
    jump_p_idio_large_i <- ifelse(is.na(jump_p_idio_large), 0.01, jump_p_idio_large)
    jump_p_idio_midsize_i <- ifelse(is.na(jump_p_idio_midsize), 0.005, jump_p_idio_midsize)
    jump_p_idio_small_i <- ifelse(is.na(jump_p_idio_small), 0.002, jump_p_idio_small)

    jump_scale_sys_i <- ifelse(is.na(jump_scale_sys), 0.35, jump_scale_sys)
    jump_scale_idio_i <- ifelse(is.na(jump_scale_idio), 0.20, jump_scale_idio)
    jump_scale_i <- ifelse(is.na(jump_scale), 0.10, jump_scale)
    jump_df_i <- ifelse(is.na(jump_df), 5, jump_df)

    jump_kappa_large_i <- ifelse(is.na(jump_kappa_large), 0.04, jump_kappa_large)
    jump_kappa_midsize_i <- ifelse(is.na(jump_kappa_midsize), 0.03, jump_kappa_midsize)
    jump_kappa_small_i <- ifelse(is.na(jump_kappa_small), 0.02, jump_kappa_small)

    jump_model <- list(
      type = jump_model_type_i,
      mu = sim_mu,
      sigma = sim_sigma,
      df_F = jump_df_F_i,
      phi_F = jump_phi_F_i,
      df_eps = c(large = jump_df_eps_large_i, midsize = jump_df_eps_midsize_i, small = jump_df_eps_small_i),
      p_sys = jump_p_sys_i,
      p_idio = c(large = jump_p_idio_large_i, midsize = jump_p_idio_midsize_i, small = jump_p_idio_small_i),
      jump_dist = jump_dist_i,
      jump_scale_sys = jump_scale_sys_i,
      jump_scale_idio = jump_scale_idio_i,
      jump_scale = jump_scale_i,
      jump_df = jump_df_i,
      sys_loading = jump_sys_loading_i,
      mean_reversion = jump_mean_reversion,
      kappa = c(large = jump_kappa_large_i, midsize = jump_kappa_midsize_i, small = jump_kappa_small_i)
    )


    jump_cache <- list(
      beta_k = pca_cache$beta_k,
      state = new.env(parent = emptyenv())
    )

    sim_jump <- simulate_rank_paths(
      w0 = w0_ext,
      K_cut = K_cut,
      K_max = K_max,
      T = sim_T_weeks,
      n_paths = sim_n_paths,
      mu = sim_mu,
      sigma = sim_sigma,
      entry_frac = sim_entry_frac,
      mean_vec = mean_vec_s,
      sd_vec = sd_vec_s,
      horizons = horizons_durable,
      K_xi = sim_K_xi,
      model = jump_model,
      moment_curves = moment_curves_jump,
      cache = jump_cache,
      entrant_sampler = entrant_sampler
    )

    sim_jump_cdc <- sim_jump$snapshots %>%
      group_by(rank) %>%
      summarise(w_bar = mean(share, na.rm = TRUE), .groups = "drop") %>%
      arrange(rank)

    score_jump <- tibble(
      model = "jump_factor",
      rmse_cdc = cdc_rmse(emp_cdc, sim_jump_cdc),
      rmse_durable = durable_rmse(emp_targets, sim_jump$growth, bucket_def),
      rmse_xi = xi_rmse(rho_by_k_med %>% filter(k <= sim_K_xi - 1L),
                        sim_jump$xi %>% filter(k <= sim_K_xi - 1L))
    )

    # Tail-metric comparison (empirical vs jump simulation)
    tail_sample_steps <- 200L
    tail_sample_paths <- 30L
    tail_sample_ranks_per_bucket <- 200L

    tail_cache <- list(
      beta_k = pca_cache$beta_k,
      state = new.env(parent = emptyenv())
    )

    sim_inc_tail <- sample_increments_for_tail(
      model = jump_model,
      moment_curves = moment_curves_jump,
      cache = tail_cache,
      K_cut = K_cut,
      n_steps = tail_sample_steps,
      n_paths = tail_sample_paths,
      ranks_per_bucket = tail_sample_ranks_per_bucket
    )

    tail_summ_jump <- tail_diagnostics(sim_inc_tail, bucket_def)

    tail_comp_jump <- tail_summ %>%
      select(bucket, p_gt_5sd, p_gt_3sd, q_0.001, q_0.999) %>%
      pivot_longer(-bucket, names_to = "metric", values_to = "emp") %>%
      left_join(
        tail_summ_jump %>%
          select(bucket, p_gt_5sd, p_gt_3sd, q_0.001, q_0.999) %>%
          pivot_longer(-bucket, names_to = "metric", values_to = "sim"),
        by = c("bucket", "metric")
      ) %>%
      arrange(bucket, metric)

    list(score_jump = score_jump, tail_comp_jump = tail_comp_jump)
  }, error = function(e) {
    cat("Jump model failed: ", conditionMessage(e), "\n")
    NULL
  })

  if (!is.null(jump_out)) {
    print(jump_out$score_jump)
    print(jump_out$tail_comp_jump)
  }
}
```


## Micro-level movement validation (v12) + rank-slot comparator

### 14.1 Empirical micro object (identity-based, v12)

```{r micro-emp}
emp_micro_k12 <- emp_micro_for_k(
  endpoint_weekly = endpoint_weekly %>% filter(rank <= K_cut),
  k = 12L,
  horizons = horizons_micro
)

emp_summary_k12 <- emp_micro_k12 %>%
  group_by(horizon) %>%
  summarise(
    n_total = n(),
    retention = mean(present_tp),
    n_present = sum(present_tp),
    med = median(share_tp[present_tp], na.rm = TRUE),
    p10 = quantile(share_tp[present_tp], 0.10, na.rm = TRUE),
    p90 = quantile(share_tp[present_tp], 0.90, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(source = "empirical")
```

### 14.2 Simulation micro object (rank-slot, v12 style)

```{r micro-sim}
sim_micro_for_k <- function(sim_snapshots, k, horizons = c(1L, 4L, 12L)) {

  base <- sim_snapshots %>%
    filter(rank == k) %>%
    transmute(path, t, share_t = share)

  fut <- sim_snapshots %>%
    filter(rank == k) %>%
    transmute(path, t_tp = t, share_tp = share)

  purrr::map_dfr(horizons, function(h) {
    base %>%
      mutate(t_tp = t + as.integer(h), horizon = as.integer(h)) %>%
      left_join(fut, by = c("path", "t_tp")) %>%
      filter(!is.na(share_tp))
  })
}

sim_micro_k12 <- sim_micro_for_k(
  sim_snapshots = sim_out$snapshots,
  k = 12L,
  horizons = horizons_micro
)

sim_summary_k12 <- sim_micro_k12 %>%
  group_by(horizon) %>%
  summarise(
    n_total = n(),
    retention = 1.0,
    n_present = n(),
    med = median(share_tp, na.rm = TRUE),
    p10 = quantile(share_tp, 0.10, na.rm = TRUE),
    p90 = quantile(share_tp, 0.90, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(source = "simulation")

micro_summary_k12 <- bind_rows(emp_summary_k12, sim_summary_k12)

micro_summary_k12
```

```{r micro-k12-plot}
ggplot(micro_summary_k12, aes(x = horizon, y = med, group = source, linetype = source)) +
  geom_ribbon(
    aes(ymin = p10, ymax = p90, fill = source),
    alpha = 0.35,
    color = NA
  ) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = horizons_micro) +
  labs(
    title = "Micro movement for rank k = 12",
    subtitle = "Median share at t+h with 10-90% band (empirical vs simulation)",
    x = "Horizon (weeks ahead)",
    y = "Share at t+h"
  ) +
  theme_bw()
```

### 14.3 Empirical rank-slot micro comparator (new)

This uses the *rank-slot* share at t+h (not tracking endpoint identity), to make an apples-to-apples comparison
with the simulation object.

```{r micro-emp-rankslot}
emp_rankslot_micro_for_k <- function(endpoint_weekly, k, horizons = c(1L, 4L, 12L)) {
  base <- endpoint_weekly %>%
    filter(rank == k) %>%
    transmute(week_t = week, share_t = share_topK)

  fut <- endpoint_weekly %>%
    filter(rank == k) %>%
    transmute(week_tp = week, share_tp = share_topK)

  purrr::map_dfr(horizons, function(h) {
    base %>%
      mutate(week_tp = week_t + weeks(as.integer(h)), horizon = as.integer(h)) %>%
      left_join(fut, by = "week_tp") %>%
      filter(!is.na(share_tp))
  })
}

emp_rs_k12 <- emp_rankslot_micro_for_k(
  endpoint_weekly = endpoint_weekly %>% filter(rank <= K_cut),
  k = 12L,
  horizons = horizons_micro
)

emp_rs_summary_k12 <- emp_rs_k12 %>%
  group_by(horizon) %>%
  summarise(
    n_total = n(),
    med = median(share_tp, na.rm = TRUE),
    p10 = quantile(share_tp, 0.10, na.rm = TRUE),
    p90 = quantile(share_tp, 0.90, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(source = "emp_rankslot")

micro_summary_k12_3way <- bind_rows(
  emp_summary_k12 %>% select(horizon, med, p10, p90, source),
  emp_rs_summary_k12,
  sim_summary_k12 %>% select(horizon, med, p10, p90, source)
)

ggplot(micro_summary_k12_3way, aes(x = horizon, y = med, group = source, linetype = source)) +
  geom_ribbon(
    aes(ymin = p10, ymax = p90, fill = source),
    alpha = 0.25,
    color = NA
  ) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = horizons_micro) +
  labs(
    title = "Micro movement (k=12): identity vs rank-slot vs simulation",
    subtitle = "Empirical identity-tracking differs conceptually from rank-slot tracking",
    x = "Horizon (weeks ahead)",
    y = "Share at t+h"
  ) +
  theme_bw()
```

## Micro fit across ranks (v12-style)

This reproduces the v12 idea of scoring micro-movement fit across a grid of ranks (here: 10, 20, …, 200).
It compares empirical **identity-based** future shares (conditioned on presence at t+h) to the simulated
**rank-slot** future shares.

```{r micro-fit-grid}
k_grid_micro <- seq(10L, 200L, by = 10L)

micro_fit_tbl <- purrr::map_dfr(k_grid_micro, function(k) {

  emp <- emp_micro_for_k(
    endpoint_weekly = endpoint_weekly %>% filter(rank <= K_cut),
    k = k,
    horizons = horizons_micro
  ) %>%
    group_by(horizon) %>%
    summarise(
      emp_retention = mean(present_tp),
      emp_med = median(share_tp[present_tp], na.rm = TRUE),
      .groups = "drop"
    )

  sim <- sim_micro_for_k(
    sim_snapshots = sim_out$snapshots,
    k = k,
    horizons = horizons_micro
  ) %>%
    group_by(horizon) %>%
    summarise(
      sim_med = median(share_tp, na.rm = TRUE),
      .groups = "drop"
    )

  emp %>%
    inner_join(sim, by = "horizon") %>%
    mutate(
      k = k,
      sq_err = (emp_med - sim_med)^2
    )
})

micro_rmse_by_k <- micro_fit_tbl %>%
  group_by(k) %>%
  summarise(
    rmse = sqrt(mean(sq_err, na.rm = TRUE)),
    avg_retention = mean(emp_retention, na.rm = TRUE),
    .groups = "drop"
  )

micro_rmse_by_k
```

```{r micro-fit-grid-plot}
ggplot(micro_rmse_by_k, aes(k, rmse)) +
  geom_line() +
  geom_point(size = 1.5) +
  labs(
    title = "Micro-movement RMSE by rank (k)",
    subtitle = "RMSE over horizons between empirical identity-based median(share_tp) and simulated median(share_tp)",
    x = "Rank k",
    y = "RMSE"
  ) +
  theme_bw()

ggplot(micro_rmse_by_k, aes(k, avg_retention)) +
  geom_line() +
  geom_point(size = 1.5) +
  scale_y_continuous(labels = percent_format(accuracy = 0.1)) +
  labs(
    title = "Empirical retention by rank (k)",
    subtitle = "Average across horizons",
    x = "Rank k",
    y = "Retention"
  ) +
  theme_bw()
```

## Sigma calibration loop (v10) (optional but included)

This reproduces the v10 approach: scan `sigma` and pick the value that minimizes durable mismatch, with a small CDC penalty.

```{r sigma-calibration}
calibrate_sigma <- function(sigma, w0_ext, mean_vec_s, sd_vec_s) {
  sim_res <- simulate_rank_paths(
    w0 = w0_ext,
    K_cut = K_cut,
    K_max = K_max,
    T = sim_T_weeks,
    n_paths = sim_n_paths,
    mu = sim_mu,
    sigma = sigma,
    entry_frac = sim_entry_frac,
    mean_vec = mean_vec_s,
    sd_vec = sd_vec_s,
    horizons = horizons_durable,
    K_xi = NULL
  )

  sim_cdc_i <- sim_res$snapshots %>%
    group_by(rank) %>%
    summarise(w_bar = mean(share, na.rm = TRUE), .groups = "drop")

  rmse_dur <- durable_rmse(emp_targets, sim_res$growth, bucket_def)
  rmse_cdc <- cdc_rmse(emp_cdc, sim_cdc_i)

  tibble(
    sigma = sigma,
    rmse_dur = rmse_dur,
    rmse_cdc = rmse_cdc,
    score = rmse_dur + 0.25 * rmse_cdc
  )
}

sigma_grid <- seq(0.15, 0.40, by = 0.01)

sigma_results <- purrr::map_dfr(
  sigma_grid,
  calibrate_sigma,
  w0_ext = w0_ext,
  mean_vec_s = mean_vec_s,
  sd_vec_s = sd_vec_s
) %>%
  arrange(score)

best_sigma <- sigma_results %>% slice(1)
best_sigma

ggplot(sigma_results, aes(x = sigma, y = score)) +
  geom_line() +
  geom_vline(xintercept = best_sigma$sigma, linetype = "dashed") +
  labs(
    title = "Sigma calibration (v10-style)",
    x = "sigma",
    y = "score = durable RMSE + 0.25 * CDC RMSE"
  ) +
  theme_bw()
```

## Sigma calibration by horizon (v10-style) (optional)

This is the v10 extension: pick a separate `sigma` for each horizon by minimizing the horizon-specific
durable-change mismatch.

```{r sigma-calibration-by-horizon}
if (isTRUE(params$run_grid_search)) {

  sigma_by_horizon <- function(sigma, w0_ext, mean_vec_s, sd_vec_s, bucket_def) {

    sim_res <- simulate_rank_paths(
      w0 = w0_ext,
      K_cut = K_cut,
      K_max = K_max,
      T = sim_T_weeks,
      n_paths = sim_n_paths,
      mu = sim_mu,
      sigma = sigma,
      entry_frac = sim_entry_frac,
      mean_vec = mean_vec_s,
      sd_vec = sd_vec_s,
      horizons = horizons_durable,
      K_xi = NULL
    )

    sim_targets <- sim_res$growth %>%
      mutate(bucket = assign_bucket(rank, bucket_def)) %>%
      filter(!is.na(bucket)) %>%
      group_by(bucket, horizon) %>%
      summarise(med_abs_sim = median(abs(log_growth), na.rm = TRUE), .groups = "drop")

    emp_targets %>%
      inner_join(sim_targets, by = c("bucket", "horizon")) %>%
      group_by(horizon) %>%
      summarise(
        sigma = sigma,
        rmse_h = sqrt(mean((med_abs_emp - med_abs_sim)^2, na.rm = TRUE)),
        .groups = "drop"
      )
  }

  sigma_h_tbl <- purrr::map_dfr(
    sigma_grid,
    sigma_by_horizon,
    w0_ext = w0_ext,
    mean_vec_s = mean_vec_s,
    sd_vec_s = sd_vec_s,
    bucket_def = bucket_def
  )

  best_sigma_by_horizon <- sigma_h_tbl %>%
    group_by(horizon) %>%
    arrange(rmse_h, .by_group = TRUE) %>%
    slice(1) %>%
    ungroup()

  best_sigma_by_horizon

  ggplot(sigma_h_tbl, aes(x = sigma, y = rmse_h)) +
    geom_line() +
    facet_wrap(~ horizon, scales = "free_y") +
    labs(
      title = "Sigma calibration by horizon",
      x = "sigma",
      y = "RMSE within horizon"
    ) +
    theme_bw()
}
```

## Grid search over (mu, sigma, h) (v9-style) (optional)

This reproduces v9’s idea: jointly tune a global drift `mu`, global volatility scale `sigma`,
and smoothing half-width `h`, using a weighted score combining CDC and durable-change mismatch.

```{r grid-search-mu-sigma-h}
if (isTRUE(params$run_grid_search)) {

  mu_grid <- c(-0.01, -0.005, 0.0, 0.005)
  sigma_grid2 <- c(0.1, 0.3, 0.5, 0.7, 1.0)
  h_grid <- c(0L, 2L, 5L)

  lambda_cdc <- 0.7

  param_grid <- tidyr::expand_grid(mu = mu_grid, sigma = sigma_grid2, h = h_grid)
  grid_results <- vector("list", nrow(param_grid))

  for (i in seq_len(nrow(param_grid))) {

    mu_i <- param_grid$mu[i]
    sigma_i <- param_grid$sigma[i]
    h_i <- as.integer(param_grid$h[i])

    sm_i <- build_gaussian_params_smoothed(h = h_i)

    sim_i <- simulate_rank_paths(
      w0 = w0_ext,
      K_cut = K_cut,
      K_max = K_max,
      T = sim_T_weeks,
      n_paths = sim_n_paths,
      mu = mu_i,
      sigma = sigma_i,
      entry_frac = sim_entry_frac,
      mean_vec = sm_i$mean_dlogw_s,
      sd_vec = sm_i$sd_dlogw_s,
      horizons = horizons_durable,
      K_xi = NULL
    )

    sim_cdc_i <- sim_i$snapshots %>%
      group_by(rank) %>%
      summarise(w_bar = mean(share, na.rm = TRUE), .groups = "drop")

    rmse_cdc_i <- cdc_rmse(emp_cdc, sim_cdc_i)
    rmse_dur_i <- durable_rmse(emp_targets, sim_i$growth, bucket_def)

    score_i <- lambda_cdc * rmse_cdc_i + (1 - lambda_cdc) * rmse_dur_i

    grid_results[[i]] <- tibble(
      mu = mu_i, sigma = sigma_i, h = h_i,
      rmse_cdc = rmse_cdc_i,
      rmse_durable = rmse_dur_i,
      score = score_i
    )
  }

  grid_tbl <- bind_rows(grid_results) %>% arrange(score)
  grid_tbl %>% slice(1:10)

  ggplot(grid_tbl, aes(x = score)) +
    geom_histogram(bins = 30) +
    labs(
      title = "Grid-search score distribution",
      x = "score",
      y = "count"
    ) +
    theme_bw()
}
```

## Alternative model: bootstrap increment samplers (best-practice)

This section generalizes the bootstrap simulator to preserve heavy tails and local rank dependence,
with optional common-shock factors and week-profile resampling.

```{r bootstrap-samplers}
build_pool_by_rank <- function(rank_panel, K_max, value_col = "dlogw") {
  pools <- vector("list", K_max)
  if (!value_col %in% names(rank_panel)) {
    stop("value_col not in rank_panel: ", value_col)
  }
  rp <- rank_panel %>% filter(rank <= K_max)
  split_vals <- split(rp[[value_col]], rp$rank)
  for (r in seq_len(K_max)) {
    v <- split_vals[[as.character(r)]]
    if (is.null(v)) v <- numeric(0)
    pools[[r]] <- v
  }
  pools
}

resolve_rank_bandwidth <- function(rank_bandwidth, r) {
  h <- if (is.function(rank_bandwidth)) {
    rank_bandwidth(r)
  } else if (length(rank_bandwidth) > 1L) {
    rank_bandwidth[pmin(r, length(rank_bandwidth))]
  } else {
    rank_bandwidth
  }
  h <- as.integer(round(h))
  if (!is.finite(h) || h < 0L) h <- 0L
  h
}

kernel_weights <- function(dist, h, kernel) {
  if (length(dist) == 0) return(numeric(0))
  if (h <= 0L) return(rep(1, length(dist)))
  if (kernel == "uniform") return(rep(1, length(dist)))
  if (kernel == "triangular") return(pmax(0, 1 - dist / (h + 1)))
  if (kernel == "exp") {
    tau <- max(1e-6, h / 2)
    return(exp(-dist / tau))
  }
  rep(1, length(dist))
}

build_rank_neighborhoods <- function(K_max, rank_bandwidth, kernel = "uniform") {
  neighbors <- vector("list", K_max)
  probs <- vector("list", K_max)
  for (r in seq_len(K_max)) {
    h <- resolve_rank_bandwidth(rank_bandwidth, r)
    lo <- max(1L, r - h)
    hi <- min(K_max, r + h)
    nbrs <- lo:hi
    dist <- abs(nbrs - r)
    w <- kernel_weights(dist, h, kernel)
    if (length(w) == 0 || sum(w) <= 0) w <- rep(1, length(nbrs))
    w <- w / sum(w)
    neighbors[[r]] <- nbrs
    probs[[r]] <- w
  }
  list(neighbors = neighbors, probs = probs)
}

stationary_bootstrap_indices <- function(n, L, size) {
  if (n <= 0) stop("n must be positive")
  if (L <= 1) return(sample.int(n, size = size, replace = TRUE))
  p <- 1 / L
  idx <- integer(size)
  idx[1] <- sample.int(n, size = 1)
  for (i in 2:size) {
    if (runif(1) < p) {
      idx[i] <- sample.int(n, size = 1)
    } else {
      idx[i] <- idx[i - 1L] + 1L
      if (idx[i] > n) idx[i] <- 1L
    }
  }
  idx
}

gpd_fit_mle <- function(excess) {
  excess <- excess[is.finite(excess) & excess >= 0]
  if (length(excess) < 5) return(NULL)

  nll <- function(par) {
    xi <- par[1]
    beta <- par[2]
    if (!is.finite(beta) || beta <= 0) return(Inf)
    if (any(1 + xi * excess / beta <= 0)) return(Inf)
    if (abs(xi) < 1e-6) {
      return(length(excess) * log(beta) + sum(excess) / beta)
    }
    length(excess) * log(beta) + (1 / xi + 1) * sum(log(1 + xi * excess / beta))
  }

  init <- c(0.1, max(mean(excess), 1e-6))
  fit <- tryCatch(
    optim(
      init,
      nll,
      method = "L-BFGS-B",
      lower = c(-0.5, 1e-6),
      upper = c(5, Inf)
    ),
    error = function(e) NULL
  )

  if (is.null(fit) || fit$convergence != 0 || any(!is.finite(fit$par))) return(NULL)
  list(xi = fit$par[1], beta = fit$par[2])
}

rgpd <- function(n, xi, beta) {
  if (n <= 0) return(numeric(0))
  u <- runif(n)
  if (abs(xi) < 1e-6) return(rexp(n, rate = 1 / beta))
  beta / xi * (u^(-xi) - 1)
}

build_tail_splice_fits <- function(residual_pools, bucket_assigner, tail_q, tail_fit_min_n) {
  if (is.null(bucket_assigner)) return(NULL)
  bucket_by_rank <- bucket_assigner(seq_along(residual_pools))
  bucket_levels <- unique(bucket_by_rank[!is.na(bucket_by_rank)])
  if (length(bucket_levels) == 0) return(NULL)

  fits <- list()
  for (bucket in bucket_levels) {
    ranks <- which(bucket_by_rank == bucket)
    resids <- unlist(residual_pools[ranks], use.names = FALSE)
    resids <- resids[is.finite(resids)]
    if (length(resids) == 0) next

    abs_resids <- abs(resids)
    u <- as.numeric(quantile(abs_resids, tail_q, na.rm = TRUE))
    exceed <- abs_resids[abs_resids > u] - u
    if (length(exceed) < tail_fit_min_n) next

    fit <- gpd_fit_mle(exceed)
    if (is.null(fit)) next

    fits[[as.character(bucket)]] <- list(
      u = u,
      p_tail = mean(abs_resids > u, na.rm = TRUE),
      sign_prob = mean(resids[abs_resids > u] >= 0, na.rm = TRUE),
      xi = fit$xi,
      beta = fit$beta
    )
  }

  list(fits = fits, bucket_by_rank = bucket_by_rank)
}

sample_from_pool <- function(pool) {
  if (length(pool) == 0) return(0)
  pool[sample.int(length(pool), 1L)]
}

sample_from_pool_tail_splice <- function(pool, tail_fit) {
  if (length(pool) == 0) return(0)
  if (is.null(tail_fit)) return(sample_from_pool(pool))

  u <- tail_fit$u
  body <- pool[abs(pool) <= u]
  if (length(body) == 0) body <- pool

  if (runif(1) < tail_fit$p_tail) {
    sign <- ifelse(runif(1) < tail_fit$sign_prob, 1, -1)
    return(sign * (u + rgpd(1, tail_fit$xi, tail_fit$beta)))
  }

  body[sample.int(length(body), 1L)]
}

build_sampler_iid_rank <- function(rank_panel, cfg) {
  pool_by_rank <- build_pool_by_rank(rank_panel, cfg$K_max, "dlogw")

  sample_one_week <- function(mu = 0, sigma = 1) {
    inc <- numeric(cfg$K_max)
    for (r in seq_len(cfg$K_max)) {
      inc[r] <- sample_from_pool(pool_by_rank[[r]])
    }
    mu + sigma * inc
  }

  list(
    method = "iid_rank",
    K_max = cfg$K_max,
    sample_one_week = sample_one_week,
    debug_info = list()
  )
}

build_sampler_local_rank <- function(rank_panel, cfg) {
  pool_by_rank <- build_pool_by_rank(rank_panel, cfg$K_max, "dlogw")
  neighborhoods <- build_rank_neighborhoods(cfg$K_max, cfg$rank_bandwidth, cfg$kernel)

  sample_one_week <- function(mu = 0, sigma = 1) {
    inc <- numeric(cfg$K_max)
    for (r in seq_len(cfg$K_max)) {
      nbrs <- neighborhoods$neighbors[[r]]
      probs <- neighborhoods$probs[[r]]
      r_prime <- if (length(nbrs) == 1L) nbrs else sample(nbrs, size = 1L, prob = probs)
      inc[r] <- sample_from_pool(pool_by_rank[[r_prime]])
    }
    mu + sigma * inc
  }

  list(
    method = "local_rank",
    K_max = cfg$K_max,
    sample_one_week = sample_one_week,
    debug_info = list(neighborhoods = neighborhoods)
  )
}

build_sampler_week_vector <- function(rank_panel, cfg) {
  max_rank_by_week <- rank_panel %>%
    group_by(week) %>%
    summarise(max_rank = max(rank), .groups = "drop")

  K_use <- min(cfg$K_max, min(max_rank_by_week$max_rank, na.rm = TRUE))

  rank_wide <- rank_panel %>%
    filter(rank <= K_use) %>%
    select(week, rank, dlogw) %>%
    tidyr::pivot_wider(names_from = rank, values_from = dlogw) %>%
    arrange(week)

  X <- as.matrix(rank_wide %>% select(-week))
  complete <- complete.cases(X)
  X_complete <- X[complete, , drop = FALSE]

  if (nrow(X_complete) < 2) {
    stop("Not enough complete weeks for week_vector bootstrap.")
  }

  sample_path <- function(T_sim, mu = 0, sigma = 1) {
    idx <- if (isTRUE(cfg$block_bootstrap)) {
      stationary_bootstrap_indices(nrow(X_complete), cfg$block_length, T_sim)
    } else {
      sample.int(nrow(X_complete), T_sim, replace = TRUE)
    }

    mat <- X_complete[idx, , drop = FALSE]
    if (K_use < cfg$K_max) {
      mat <- cbind(mat, matrix(0, nrow = T_sim, ncol = cfg$K_max - K_use))
    }

    mu + sigma * mat
  }

  list(
    method = "week_vector",
    K_max = cfg$K_max,
    sample_path = sample_path,
    debug_info = list(K_use = K_use)
  )
}

build_sampler_factor_local <- function(rank_panel, cfg, bucket_assigner = NULL) {
  K_max <- cfg$K_max
  K_pca <- min(cfg$K_pca, K_max)

  rp <- rank_panel %>% filter(rank <= K_max)

  rank_wide <- rp %>%
    filter(rank <= K_pca) %>%
    select(week, rank, dlogw) %>%
    tidyr::pivot_wider(names_from = rank, values_from = dlogw) %>%
    arrange(week)

  X <- as.matrix(rank_wide %>% select(-week))
  complete <- complete.cases(X)
  X_complete <- X[complete, , drop = FALSE]

  if (nrow(X_complete) < 2) {
    stop("Not enough complete weeks for factor PCA.")
  }

  pca_fit <- prcomp(X_complete, center = TRUE, scale. = TRUE)
  J <- min(cfg$n_factors, ncol(pca_fit$x))
  if (J < 1) stop("bootstrap_n_factors must be >= 1.")

  F_emp <- pca_fit$x[, seq_len(J), drop = FALSE]
  factor_weeks <- rank_wide$week[complete]

  rp_factor <- rp %>% filter(week %in% factor_weeks)
  split_by_rank <- split(rp_factor, rp_factor$rank)

  a_vec <- numeric(K_max)
  b_mat <- matrix(0, nrow = K_max, ncol = J)
  residual_pools <- vector("list", K_max)

  for (r in seq_len(K_max)) {
    df_r <- split_by_rank[[as.character(r)]]
    v_all <- rp %>% filter(rank == r) %>% pull(dlogw)

    fallback <- function() {
      if (length(v_all) == 0) {
        a_vec[r] <<- 0
        residual_pools[[r]] <<- numeric(0)
      } else {
        a_vec[r] <<- mean(v_all, na.rm = TRUE)
        residual_pools[[r]] <<- v_all - a_vec[r]
      }
    }

    if (is.null(df_r) || nrow(df_r) < (J + 2L)) {
      fallback()
      next
    }

    idx <- match(df_r$week, factor_weeks)
    valid <- !is.na(idx)
    if (sum(valid) < (J + 2L)) {
      fallback()
      next
    }

    y <- df_r$dlogw[valid]
    Xr <- cbind(1, F_emp[idx[valid], , drop = FALSE])
    fit <- lm.fit(x = Xr, y = y)
    coef <- fit$coefficients

    if (any(!is.finite(coef))) {
      fallback()
      next
    }

    a_vec[r] <- coef[1]
    b_mat[r, ] <- coef[-1]
    residual_pools[[r]] <- as.numeric(y - Xr %*% coef)
  }

  neighborhoods <- build_rank_neighborhoods(K_max, cfg$rank_bandwidth, cfg$kernel)
  tail_fits <- NULL
  if (isTRUE(cfg$tailsplice) && !is.null(bucket_assigner)) {
    tail_fits <- build_tail_splice_fits(
      residual_pools,
      bucket_assigner,
      cfg$tail_q,
      cfg$tail_fit_min_n
    )
  }

  sample_residuals <- function() {
    e <- numeric(K_max)
    for (r in seq_len(K_max)) {
      nbrs <- neighborhoods$neighbors[[r]]
      probs <- neighborhoods$probs[[r]]
      r_prime <- if (length(nbrs) == 1L) nbrs else sample(nbrs, size = 1L, prob = probs)
      pool <- residual_pools[[r_prime]]
      if (length(pool) == 0) {
        e[r] <- 0
        next
      }

      if (!is.null(tail_fits) && length(tail_fits$fits) > 0) {
        bucket <- tail_fits$bucket_by_rank[r_prime]
        fit <- tail_fits$fits[[as.character(bucket)]]
        e[r] <- sample_from_pool_tail_splice(pool, fit)
      } else {
        e[r] <- sample_from_pool(pool)
      }
    }
    e
  }

  sample_path <- function(T_sim, mu = 0, sigma = 1) {
    F_boot <- if (isTRUE(cfg$block_bootstrap)) {
      idx <- stationary_bootstrap_indices(nrow(F_emp), cfg$block_length, T_sim)
      F_emp[idx, , drop = FALSE]
    } else {
      F_emp[sample.int(nrow(F_emp), T_sim, replace = TRUE), , drop = FALSE]
    }

    out <- matrix(0, nrow = T_sim, ncol = K_max)
    for (t in seq_len(T_sim)) {
      e <- sample_residuals()
      inc_raw <- a_vec + as.vector(b_mat %*% F_boot[t, ]) + e
      out[t, ] <- mu + sigma * inc_raw
    }
    out
  }

  list(
    method = if (isTRUE(cfg$tailsplice)) "factor_local_tailsplice" else "factor_local",
    K_max = K_max,
    sample_path = sample_path,
    debug_info = list(
      a = a_vec,
      b = b_mat,
      F_emp = F_emp,
      factor_weeks = factor_weeks
    )
  )
}

build_increment_sampler <- function(rank_panel, cfg, bucket_assigner = NULL) {
  method <- cfg$bootstrap_method
  if (method == "factor_local_tailsplice") {
    cfg$tailsplice <- TRUE
    method <- "factor_local"
  }

  if (method == "iid_rank") {
    return(build_sampler_iid_rank(rank_panel, cfg))
  }
  if (method == "local_rank") {
    return(build_sampler_local_rank(rank_panel, cfg))
  }
  if (method == "factor_local") {
    return(build_sampler_factor_local(rank_panel, cfg, bucket_assigner))
  }
  if (method == "week_vector") {
    return(build_sampler_week_vector(rank_panel, cfg))
  }

  stop("Unknown bootstrap method: ", method)
}

simulate_rank_paths_generic <- function(
  w0, K_cut, K_max, T = 52, n_paths = 200,
  mu = 0, sigma = 1, entry_frac = 0.10,
  increment_generator,
  horizons = c(4L, 8L),
  K_xi = NULL,
  store_increments = FALSE,
  increments_paths = 0L,
  increments_K = K_cut
) {
  stopifnot(is.list(increment_generator))
  if (!is.null(increment_generator$K_max)) {
    stopifnot(increment_generator$K_max == K_max)
  }

  top_n_sorted <- function(w, n) w[order(w, decreasing = TRUE)][seq_len(n)]

  horizons <- as.integer(horizons)
  stopifnot(all(horizons >= 1L), all(horizons <= T))

  if (!is.null(K_xi)) {
    K_xi <- as.integer(K_xi)
    stopifnot(K_xi >= 2L, K_xi <= K_cut)
  }

  increments_paths <- as.integer(increments_paths)
  increments_K <- as.integer(min(increments_K, K_max))

  snapshots <- vector("list", n_paths)
  growth_collect <- setNames(vector("list", length(horizons)), paste0("h", horizons))
  xi_collect <- vector("list", n_paths)
  increments_collect <- vector("list", if (store_increments) min(n_paths, increments_paths) else 0L)

  for (p in seq_len(n_paths)) {

    if (length(w0) < K_max) {
      w <- c(w0, entrant_sampler(K_max - length(w0)))
      w <- w / sum(w)
    } else {
      w <- w0[1:K_max]
      w <- w / sum(w)
    }

    w_baseK <- top_n_sorted(w, K_cut)
    w_baseK <- w_baseK / sum(w_baseK)

    snaps_p <- vector("list", T)
    xi_p <- vector("list", T)
    inc_p <- if (store_increments && p <= increments_paths) vector("list", T) else NULL

    dlogw_mat <- NULL
    if (!is.null(increment_generator$sample_path)) {
      dlogw_mat <- increment_generator$sample_path(T, mu = mu, sigma = sigma)
      stopifnot(is.matrix(dlogw_mat), nrow(dlogw_mat) == T, ncol(dlogw_mat) == K_max)
    }

    for (t in seq_len(T)) {
      dlogw <- if (!is.null(dlogw_mat)) {
        dlogw_mat[t, ]
      } else if (!is.null(increment_generator$sample_one_week)) {
        increment_generator$sample_one_week(mu = mu, sigma = sigma)
      } else {
        stop("increment_generator must provide sample_path or sample_one_week")
      }

      stopifnot(length(dlogw) == K_max)

      w <- w * exp(dlogw)
      w <- pmax(w, 1e-18)
      w <- w / sum(w)

      w_top <- top_n_sorted(w, K_cut)
      n_tail <- K_max - K_cut

      if (n_tail > 0) {
        n_add <- max(1L, floor(n_tail * entry_frac))
        w <- c(w_top, entrant_sampler(n_add), entrant_sampler(n_tail - n_add))
      } else {
        w <- w_top
      }

      w <- w / sum(w)

      wK <- top_n_sorted(w, K_cut)
      wK <- wK / sum(wK)

      snaps_p[[t]] <- tibble(
        path = p,
        t = t,
        rank = seq_len(K_cut),
        share = wK
      )

      if (store_increments && p <= increments_paths) {
        inc_p[[t]] <- tibble(
          path = p,
          t = t,
          rank = seq_len(increments_K),
          dlogw = dlogw[seq_len(increments_K)]
        )
      }

      if (!is.null(K_xi)) {
        w_top_now <- top_n_sorted(w, K_cut) %>% pmax(1e-18)
        xi_vals <- log(w_top_now[seq_len(K_xi - 1L)] / w_top_now[seq(2L, K_xi)])
        xi_p[[t]] <- tibble(
          path = p,
          t = t,
          k = seq_len(K_xi - 1L),
          xi = xi_vals
        )
      }

      if (t %in% horizons) {
        lg <- log(wK / w_baseK)
        lbl <- paste0("h", t)
        growth_collect[[lbl]] <- rbind(
          growth_collect[[lbl]],
          cbind(
            path = p,
            rank = seq_len(K_cut),
            horizon = t,
            log_growth = lg
          )
        )
      }
    }

    snapshots[[p]] <- bind_rows(snaps_p)
    if (!is.null(K_xi)) xi_collect[[p]] <- bind_rows(xi_p)
    if (store_increments && p <= increments_paths) {
      increments_collect[[p]] <- bind_rows(inc_p)
    }
  }

  growth_df <- bind_rows(lapply(growth_collect, as_tibble))

  out <- list(
    snapshots = bind_rows(snapshots),
    growth = growth_df
  )

  if (!is.null(K_xi)) out$xi <- bind_rows(xi_collect)
  if (store_increments && length(increments_collect) > 0) {
    out$increments <- bind_rows(increments_collect)
  }

  out
}
```

```{r bootstrap-model}
if (isTRUE(params$run_bootstrap_model)) {

  bootstrap_cfg <- list(
    K_cut = K_cut,
    K_max = K_max,
    T_sim = sim_T_weeks,
    n_paths = sim_n_paths,
    entry_frac = sim_entry_frac,
    bootstrap_method = bootstrap_method,
    rank_bandwidth = bootstrap_rank_bandwidth,
    kernel = bootstrap_kernel,
    block_bootstrap = bootstrap_block_bootstrap,
    block_length = bootstrap_block_length,
    n_factors = bootstrap_n_factors,
    K_pca = bootstrap_K_pca,
    tailsplice = bootstrap_tailsplice,
    tail_q = bootstrap_tail_q,
    tail_fit_min_n = bootstrap_tail_fit_min_n
  )

  increment_sampler <- build_increment_sampler(
    rank_panel,
    cfg = bootstrap_cfg,
    bucket_assigner = function(r) assign_bucket(r, bucket_def)
  )

  sim_boot <- simulate_rank_paths_generic(
    w0 = w0_ext,
    K_cut = K_cut,
    K_max = K_max,
    T = sim_T_weeks,
    n_paths = sim_n_paths,
    mu = sim_mu,
    sigma = sim_sigma,
    entry_frac = sim_entry_frac,
    increment_generator = increment_sampler,
    horizons = horizons_durable,
    K_xi = sim_K_xi,
    store_increments = bootstrap_diag_paths > 0,
    increments_paths = bootstrap_diag_paths,
    increments_K = K_cut
  )

  sim_boot_cdc <- sim_boot$snapshots %>%
    group_by(rank) %>%
    summarise(w_bar = mean(share, na.rm = TRUE), .groups = "drop")

  score_boot <- tibble(
    model = paste0("bootstrap_", increment_sampler$method),
    rmse_cdc = cdc_rmse(emp_cdc, sim_boot_cdc),
    rmse_durable = durable_rmse(emp_targets, sim_boot$growth, bucket_def),
    rmse_xi = xi_rmse(
      rho_by_k_med %>% filter(k <= sim_K_xi - 1L),
      sim_boot$xi %>% filter(k <= sim_K_xi - 1L)
    )
  )

  score_gauss <- score_tbl %>% mutate(model = "gaussian_smoothed")

  bind_rows(score_gauss, score_boot) %>%
    select(model, everything())

  bind_rows(
    emp_cdc %>% mutate(source = "Empirical"),
    sim_cdc %>% mutate(source = "Gaussian"),
    sim_boot_cdc %>% mutate(source = paste0("Bootstrap (", increment_sampler$method, ")"))
  ) %>%
    ggplot(aes(rank, w_bar, linetype = source)) +
    geom_line() +
    scale_x_log10() +
    scale_y_log10(labels = percent_format(accuracy = 0.01)) +
    labs(
      title = "CDC comparison across models",
      x = "Rank (log scale)",
      y = "Mean share within top-K (log scale)",
      linetype = NULL
    ) +
    theme_bw()

  if (!is.null(sim_boot$increments)) {
    tail_summ_boot <- tail_diagnostics(sim_boot$increments, bucket_def) %>%
      mutate(source = paste0("Bootstrap (", increment_sampler$method, ")"))

    bind_rows(
      tail_summ %>% mutate(source = "Empirical"),
      tail_summ_boot
    ) %>%
      select(source, everything())
  }

  # Minimal invariants / smoke checks
  stopifnot(nrow(sim_boot$snapshots) == sim_n_paths * sim_T_weeks * K_cut)
  stopifnot(all(is.finite(sim_boot$snapshots$share)))
  stopifnot(min(sim_boot$snapshots$share, na.rm = TRUE) > 0)

  sum_by_pt <- tapply(
    sim_boot$snapshots$share,
    list(sim_boot$snapshots$path, sim_boot$snapshots$t),
    sum
  )
  stopifnot(all(abs(sum_by_pt - 1) < 1e-6))

  # Deterministic with fixed seed
  seed_state <- if (exists(".Random.seed", envir = .GlobalEnv)) {
    get(".Random.seed", envir = .GlobalEnv)
  } else {
    NULL
  }

  set.seed(params$seed)
  if (!is.null(increment_sampler$sample_one_week)) {
    test1 <- increment_sampler$sample_one_week(mu = sim_mu, sigma = sim_sigma)
    set.seed(params$seed)
    test2 <- increment_sampler$sample_one_week(mu = sim_mu, sigma = sim_sigma)
    stopifnot(isTRUE(all.equal(test1, test2)))
  } else {
    test1 <- increment_sampler$sample_path(T_sim = 5L, mu = sim_mu, sigma = sim_sigma)
    set.seed(params$seed)
    test2 <- increment_sampler$sample_path(T_sim = 5L, mu = sim_mu, sigma = sim_sigma)
    stopifnot(isTRUE(all.equal(test1, test2)))
  }

  if (is.null(seed_state)) {
    rm(".Random.seed", envir = .GlobalEnv)
  } else {
    assign(".Random.seed", seed_state, envir = .GlobalEnv)
  }
}
```

## Notes for SPT-facing interpretation (non-quantitative)

- This document computes both **rank-slot** objects (ranked shares and their increments) and
  **identity-tracking** objects (follow the endpoint currently at rank k forward).
- For SPT / rank-based diffusion discussions, the rank-slot objects are the most direct analog.
- Identity-tracking objects are essential for **turnover** and will often motivate extensions
  (birth/death, regime switching, common shocks, jumps).
