## Micro-level movement validation (v12) + rank-slot comparator

### 14.1 Empirical micro object (identity-based, v12)

```{r micro-setup, include=FALSE}
if (!exists("endpoint_bundle")) {
  endpoint_bundle <- load_endpoint_weekly(cfg = CFG)
}
endpoint_weekly <- endpoint_bundle$endpoint_weekly
K_cut <- endpoint_bundle$K_cut
K_max <- endpoint_bundle$K_max
max_rank_seen <- endpoint_bundle$max_rank_seen
sim_K_xi <- endpoint_bundle$sim_K_xi

horizons_micro <- CFG$horizons_micro

if (!exists("rank_panel")) {
  rank_panel <- load_rank_panel(endpoint_weekly, cfg = CFG)
}
if (!exists("gauss_params_raw")) {
  gauss_params_raw <- load_gauss_params_raw(rank_panel, K_max, cfg = CFG)
}
if (!exists("sm_params")) {
  sm_params <- load_gauss_params_smoothed(gauss_params_raw, K_max, CFG$smoothing_h, cfg = CFG)
}

entrant_pool <- build_entrant_pool(endpoint_weekly, K_cut, CFG$K_tail_buffer, max_rank_seen)
entrant_sampler <- make_entrant_sampler(entrant_pool)
w0_ext <- build_w0_ext(endpoint_weekly, K_max, entrant_sampler)

mean_vec_s <- sm_params$mean_dlogw_s
sd_vec_s <- sm_params$sd_dlogw_s

if (!exists("sim_out")) {
  sim_out <- load_sim_baseline(
    w0_ext = w0_ext,
    mean_vec_s = mean_vec_s,
    sd_vec_s = sd_vec_s,
    K_cut = K_cut,
    K_max = K_max,
    sim_K_xi = sim_K_xi,
    cfg = CFG,
    entrant_sampler = entrant_sampler
  )
}
```

```{r micro-emp}
emp_micro_k12 <- emp_micro_for_k(
  endpoint_weekly = endpoint_weekly %>% filter(rank <= K_cut),
  k = 12L,
  horizons = horizons_micro
)

emp_summary_k12 <- emp_micro_k12 %>%
  group_by(horizon) %>%
  summarise(
    n_total = n(),
    retention = mean(present_tp),
    n_present = sum(present_tp),
    med = median(share_tp[present_tp], na.rm = TRUE),
    p10 = quantile(share_tp[present_tp], 0.10, na.rm = TRUE),
    p90 = quantile(share_tp[present_tp], 0.90, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(source = "empirical")
```

### 14.2 Simulation micro object (rank-slot, v12 style)

```{r micro-sim}
sim_micro_k12 <- sim_micro_for_k(
  sim_snapshots = sim_out$snapshots,
  k = 12L,
  horizons = horizons_micro
)

sim_summary_k12 <- sim_micro_k12 %>%
  group_by(horizon) %>%
  summarise(
    n_total = n(),
    retention = 1.0,
    n_present = n(),
    med = median(share_tp, na.rm = TRUE),
    p10 = quantile(share_tp, 0.10, na.rm = TRUE),
    p90 = quantile(share_tp, 0.90, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(source = "simulation")

micro_summary_k12 <- bind_rows(emp_summary_k12, sim_summary_k12)

micro_summary_k12
```

```{r micro-k12-plot}
ggplot(micro_summary_k12, aes(x = horizon, y = med, group = source, linetype = source)) +
  geom_ribbon(
    aes(ymin = p10, ymax = p90, fill = source),
    alpha = 0.35,
    color = NA
  ) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = horizons_micro) +
  labs(
    title = "Micro movement for rank k = 12",
    subtitle = "Median share at t+h with 10-90% band (empirical vs simulation)",
    x = "Horizon (weeks ahead)",
    y = "Share at t+h"
  ) +
  theme_bw()
```

### 14.3 Empirical rank-slot micro comparator (new)

```{r micro-emp-rankslot}
emp_rs_k12 <- emp_rankslot_micro_for_k(
  endpoint_weekly = endpoint_weekly %>% filter(rank <= K_cut),
  k = 12L,
  horizons = horizons_micro
)

emp_rs_summary_k12 <- emp_rs_k12 %>%
  group_by(horizon) %>%
  summarise(
    n_total = n(),
    med = median(share_tp, na.rm = TRUE),
    p10 = quantile(share_tp, 0.10, na.rm = TRUE),
    p90 = quantile(share_tp, 0.90, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(source = "emp_rankslot")

micro_summary_k12_3way <- bind_rows(
  emp_summary_k12 %>% select(horizon, med, p10, p90, source),
  emp_rs_summary_k12,
  sim_summary_k12 %>% select(horizon, med, p10, p90, source)
)

ggplot(micro_summary_k12_3way, aes(x = horizon, y = med, group = source, linetype = source)) +
  geom_ribbon(
    aes(ymin = p10, ymax = p90, fill = source),
    alpha = 0.25,
    color = NA
  ) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = horizons_micro) +
  labs(
    title = "Micro movement (k=12): identity vs rank-slot vs simulation",
    subtitle = "Empirical identity-tracking differs conceptually from rank-slot tracking",
    x = "Horizon (weeks ahead)",
    y = "Share at t+h"
  ) +
  theme_bw()
```

## Micro fit across ranks (v12-style)

This reproduces the v12 idea of scoring micro-movement fit across a grid of ranks (here: 10, 20, …, 200).
It compares empirical **identity-based** future shares (conditioned on presence at t+h) to the simulated
**rank-slot** future shares.

```{r micro-fit-grid}
k_grid_micro <- seq(10L, 200L, by = 10L)

micro_fit_tbl <- purrr::map_dfr(k_grid_micro, function(k) {

  emp <- emp_micro_for_k(
    endpoint_weekly = endpoint_weekly %>% filter(rank <= K_cut),
    k = k,
    horizons = horizons_micro
  ) %>%
    group_by(horizon) %>%
    summarise(
      emp_retention = mean(present_tp),
      emp_med = median(share_tp[present_tp], na.rm = TRUE),
      .groups = "drop"
    )

  sim <- sim_micro_for_k(
    sim_snapshots = sim_out$snapshots,
    k = k,
    horizons = horizons_micro
  ) %>%
    group_by(horizon) %>%
    summarise(
      sim_med = median(share_tp, na.rm = TRUE),
      .groups = "drop"
    )

  emp %>%
    inner_join(sim, by = "horizon") %>%
    mutate(
      k = k,
      sq_err = (emp_med - sim_med)^2
    )
})

micro_rmse_by_k <- micro_fit_tbl %>%
  group_by(k) %>%
  summarise(
    rmse = sqrt(mean(sq_err, na.rm = TRUE)),
    avg_retention = mean(emp_retention, na.rm = TRUE),
    .groups = "drop"
  )

micro_rmse_by_k
```

```{r micro-fit-grid-plot}
ggplot(micro_rmse_by_k, aes(k, rmse)) +
  geom_line() +
  geom_point(size = 1.5) +
  labs(
    title = "Micro-movement RMSE by rank (k)",
    subtitle = "RMSE over horizons between empirical identity-based median(share_tp) and simulated median(share_tp)",
    x = "Rank k",
    y = "RMSE"
  ) +
  theme_bw()

ggplot(micro_rmse_by_k, aes(k, avg_retention)) +
  geom_line() +
  geom_point(size = 1.5) +
  scale_y_continuous(labels = percent_format(accuracy = 0.1)) +
  labs(
    title = "Empirical retention by rank (k)",
    subtitle = "Average across horizons",
    x = "Rank k",
    y = "Retention"
  ) +
  theme_bw()
```
