## Alternative model: bootstrap increment samplers (best-practice) (optional)

This generalizes the early nonparametric bootstrap simulator to support:

- `iid_rank` (per-rank iid pool; the original simple bootstrap)
- `local_rank` (pooled neighborhoods in rank; recommended baseline)
- `factor_local` (common-shock factors + local residual bootstrap)
- `week_vector` (resample entire weekly increment profiles; preserves cross-rank dependence)
- `factor_local_tailsplice` (optional: semi-parametric tail splicing for residuals)

Control these with `CFG$bootstrap_*` params (see `report/master_report.Rmd` YAML).

```{r bootstrap-setup, include=FALSE}
if (!exists("endpoint_bundle")) {
  endpoint_bundle <- load_endpoint_weekly(cfg = CFG)
}
endpoint_weekly <- endpoint_bundle$endpoint_weekly
K_cut <- endpoint_bundle$K_cut
K_max <- endpoint_bundle$K_max
max_rank_seen <- endpoint_bundle$max_rank_seen

if (!exists("rank_panel")) {
  rank_panel <- load_rank_panel(endpoint_weekly, cfg = CFG)
}

entrant_pool <- build_entrant_pool(endpoint_weekly, K_cut, CFG$K_tail_buffer, max_rank_seen)
entrant_sampler <- make_entrant_sampler(entrant_pool)
w0_ext <- build_w0_ext(endpoint_weekly, K_max, entrant_sampler)
```

```{r bootstrap-model}
if (isTRUE(CFG$run_bootstrap_model)) {

  bootstrap_cfg <- list(
    K_cut = K_cut,
    K_max = K_max,
    T_sim = CFG$sim_T_weeks,
    n_paths = CFG$sim_n_paths,
    entry_frac = CFG$sim_entry_frac,
    bootstrap_method = CFG$bootstrap_method,
    rank_bandwidth = CFG$bootstrap_rank_bandwidth,
    kernel = CFG$bootstrap_kernel,
    block_bootstrap = CFG$bootstrap_block_bootstrap,
    block_length = CFG$bootstrap_block_length,
    n_factors = CFG$bootstrap_n_factors,
    K_pca = CFG$bootstrap_K_pca,
    tailsplice = CFG$bootstrap_tailsplice,
    tail_q = CFG$bootstrap_tail_q,
    tail_fit_min_n = CFG$bootstrap_tail_fit_min_n
  )

  increment_sampler <- build_increment_sampler(
    rank_panel,
    cfg = bootstrap_cfg,
    bucket_assigner = function(r) assign_bucket(r, CFG$bucket_def)
  )

  sim_boot <- simulate_rank_paths_generic(
    w0 = w0_ext,
    K_cut = K_cut,
    K_max = K_max,
    T = CFG$sim_T_weeks,
    n_paths = CFG$sim_n_paths,
    mu = CFG$sim_mu,
    sigma = CFG$sim_sigma,
    entry_frac = CFG$sim_entry_frac,
    increment_generator = increment_sampler,
    horizons = CFG$horizons_durable,
    K_xi = NULL,
    store_increments = CFG$bootstrap_diag_paths > 0,
    increments_paths = CFG$bootstrap_diag_paths,
    increments_K = K_cut,
    entrant_sampler = entrant_sampler
  )

  sim_boot
}
```

## Notes for SPT-facing interpretation (non-quantitative)

(Place SPT-facing narrative notes here.)
