---
title: "FB weekly: fast analysis for SPT-style talk"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
params:
  fb_weekly_parquet: "Facebook/fb_ranked_weekly.parquet"
  out_dir: "../assets"   # should match the TeX deck's assets/ directory
  seed: 1823

  # Universe settings
  K_cut_target: 12000
  K_tail_buffer: 2000
  K_xi_est: 2000
  min_week_ranks_keep: 12000

  # Rank smoothing (half-width)
  smoothing_h: 5

  # Buckets used repeatedly in the exploration PDFs
  bucket_breaks: !r c(0, 25, 250, Inf)
  bucket_labels: ["large", "midsize", "small"]

  # Micro horizons and representative ranks
  horizons_micro: [1, 4, 12]
  k_focus: 12
  k_grid: [1, 5, 12, 25, 100, 500, 1000]

  # Fast simulation settings (tuned for <= ~20 minutes on an M1 MacBook Pro)
  sim_T_weeks: 52
  sim_n_paths: 60
  sim_sigma: 0.16
  sim_mu: 0.0
  sim_entry_frac: 0.10

  # Bootstrap settings for the quarterly CI in v11
  bootstrap_period_unit: "quarter"
  bootstrap_B_fast: 75
  bootstrap_k_min: 10

  # Optional expensive steps
  run_sigma_calibration: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

suppressPackageStartupMessages({
  library(tidyverse)
  library(arrow)
  library(lubridate)
  library(scales)
  library(here)
})

source(here::here("R", "utils.R"))

set.seed(params$seed)

out_dir <- params$out_dir
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)
if (!dir.exists(out_dir)) {
  fallback_dir <- "assets"
  dir.create(fallback_dir, recursive = TRUE, showWarnings = FALSE)
  if (dir.exists(fallback_dir)) {
    out_dir <- fallback_dir
  }
}

`%||%` <- function(x, y) if (is.null(x)) y else x
```

# 1. Load data (weekly snapshots) and compute shares

```{r load-data}
fb_weekly_parquet_path <- params$fb_weekly_parquet
if (!file.exists(fb_weekly_parquet_path)) {
  candidates <- c(
    here::here("fb_ranked_weekly.parquet"),
    here::here("Facebook", "fb_ranked_weekly.parquet")
  )
  fb_weekly_parquet_path <- candidates[file.exists(candidates)][1]
}

if (is.na(fb_weekly_parquet_path) || !file.exists(fb_weekly_parquet_path)) {
  stop(
    "Weekly parquet not found. Tried: ",
    paste(c(params$fb_weekly_parquet,
            here::here("fb_ranked_weekly.parquet"),
            here::here("Facebook", "fb_ranked_weekly.parquet")),
          collapse = ", ")
  )
}

raw <- read_parquet(fb_weekly_parquet_path)

endpoint_weekly <- raw %>%
  transmute(
    week = as.Date(date),
    endpoint_id,
    metric_value,
    rank = rank
  ) %>%
  group_by(week) %>%
  arrange(desc(metric_value), .by_group = TRUE) %>%
  mutate(
    rank = row_number(),
    share_global = metric_value / sum(metric_value, na.rm = TRUE)
  ) %>%
  ungroup()

filter_result <- filter_bad_weeks(
  endpoint_weekly,
  min_week_ranks_keep = params$min_week_ranks_keep,
  verbose = TRUE
)
endpoint_weekly <- filter_result$endpoint_weekly
bad_weeks <- filter_result$bad_weeks
week_counts <- filter_result$week_counts

max_rank_seen <- endpoint_weekly %>%
  summarise(max_rank = max(rank, na.rm = TRUE)) %>%
  pull(max_rank)

K_cut <- min(as.integer(params$K_cut_target), as.integer(max_rank_seen))
K_max <- min(as.integer(K_cut + params$K_tail_buffer), as.integer(max_rank_seen))

endpoint_weekly <- endpoint_weekly %>%
  group_by(week) %>%
  mutate(
    share_topK = if_else(
      rank <= K_cut,
      share_global / sum(share_global[rank <= K_cut], na.rm = TRUE),
      NA_real_
    )
  ) %>%
  ungroup()

endpoint_weekly %>%
  summarise(
    min_week = min(week),
    max_week = max(week),
    n_weeks = n_distinct(week),
    n_endpoints = n_distinct(endpoint_id),
    K_cut = K_cut,
    K_max = K_max
  )
```

# 2. Core empirical objects for the talk

## 2.1 Capital distribution curve (CDC)

```{r emp-cdc}
emp_cdc <- endpoint_weekly %>%
  filter(rank <= K_cut) %>%
  group_by(rank) %>%
  summarise(w_bar = mean(share_topK, na.rm = TRUE), .groups = "drop") %>%
  arrange(rank)
```

## 2.2 Weekly rank increments (drift/volatility by rank)

```{r rank-increments}
rank_panel <- endpoint_weekly %>%
  filter(rank <= K_max) %>%
  group_by(rank) %>%
  arrange(week, .by_group = TRUE) %>%
  mutate(
    share = pmax(share_topK, 1e-12),
    log_w = log(share),
    dlogw = log_w - lag(log_w)
  ) %>%
  ungroup() %>%
  filter(is.finite(dlogw))

gauss_params_raw <- rank_panel %>%
  group_by(rank) %>%
  summarise(
    mean_dlogw = mean(dlogw, na.rm = TRUE),
    sd_dlogw = sd(dlogw, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  right_join(tibble(rank = 1:K_max), by = "rank") %>%
  arrange(rank) %>%
  mutate(
    mean_dlogw = replace_na(mean_dlogw, 0),
    sd_dlogw = replace_na(sd_dlogw, 0),
    sd_dlogw = if_else(sd_dlogw <= 0 | !is.finite(sd_dlogw), 1e-6, sd_dlogw)
  )

moving_average_rank <- function(x, h) {
  stopifnot(h >= 0L)
  n <- length(x)
  if (h == 0L) return(x)
  cs <- c(0, cumsum(replace_na(x, 0)))
  out <- numeric(n)
  for (r in seq_len(n)) {
    lo <- max(1L, r - h)
    hi <- min(n, r + h)
    out[r] <- (cs[hi + 1L] - cs[lo]) / (hi - lo + 1L)
  }
  out
}

build_gaussian_params_smoothed <- function(h) {
  tibble(
    rank = 1:K_max,
    mean_dlogw_s = moving_average_rank(gauss_params_raw$mean_dlogw, h = h),
    sd_dlogw_s = pmax(moving_average_rank(gauss_params_raw$sd_dlogw, h = h), 1e-6)
  )
}

sm <- build_gaussian_params_smoothed(h = as.integer(params$smoothing_h))
```

## 2.3 Gap process: \(\Xi_k(t)=\log(\mu_{(k)}(t)/\mu_{(k+1)}(t))\)

```{r xi-weekly}
# Note: Xi ratios are invariant to renormalization by a constant, but we compute
# them from share_global to stay close to v11.
xi_weekly <- endpoint_weekly %>%
  group_by(week) %>%
  arrange(rank, .by_group = TRUE) %>%
  mutate(
    share = pmax(share_global, 1e-18),
    share_next = lead(share),
    xi = log(share / share_next),
    k = rank
  ) %>%
  ungroup() %>%
  filter(!is.na(share_next), is.finite(xi), xi > 0)

xi_summary_by_k <- xi_weekly %>%
  group_by(k) %>%
  summarise(
    xi_med = median(xi, na.rm = TRUE),
    xi_p10 = quantile(xi, 0.10, na.rm = TRUE),
    xi_p90 = quantile(xi, 0.90, na.rm = TRUE),
    .groups = "drop"
  )

p_xi <- ggplot(filter(xi_summary_by_k, k <= K_cut - 1L), aes(x = k, y = xi_med)) +
  geom_line(linewidth = 0.8) +
  geom_ribbon(aes(ymin = xi_p10, ymax = xi_p90), alpha = 0.2) +
  scale_x_continuous(trans = "log10") +
  ylim(0, 1) +
  labs(
    title = "Empirical log-gaps by rank (median + 10-90% band)",
    x = "Rank (log)",
    y = "Xi_k"
  ) +
  theme_bw()

ggsave(file.path(out_dir, "xi_median_band.png"), p_xi,
       width = 7.2, height = 5.6, dpi = 200)
```

```{r xi-bucket-time}
assign_bucket <- function(rank, breaks, labels) {
  cut(
    rank,
    breaks = breaks,
    labels = labels,
    include.lowest = TRUE,
    right = TRUE
  )
}

xi_bucket_week <- xi_weekly %>%
  mutate(bucket = assign_bucket(k, params$bucket_breaks, params$bucket_labels)) %>%
  group_by(week, bucket) %>%
  summarise(xi_med = median(xi, na.rm = TRUE), .groups = "drop")

p_bucket <- ggplot(xi_bucket_week, aes(x = week, y = xi_med, color = bucket)) +
  geom_line(alpha = 0.8) +
  geom_smooth(se = TRUE) +
  labs(
    title = "Weekly median Xi by rank bucket",
    subtitle = "Loess smoother highlights temporal trends",
    x = NULL,
    y = "Median Xi"
  ) +
  theme_bw()

ggsave(file.path(out_dir, "xi_bucket_time.png"), p_bucket,
       width = 7.2, height = 5.6, dpi = 200)
```

# 3. The v11 ``implied volatility'' diagnostic: \(\widehat{\sigma}(k)\)

```{r sigma-hat}
# rho_k: median Xi by rank
rho_by_k_med <- xi_weekly %>%
  filter(k <= (as.integer(params$K_xi_est) - 1L)) %>%
  group_by(k) %>%
  summarise(rho_k = median(xi, na.rm = TRUE), .groups = "drop")

# g_k from smoothed drift estimates, then S_k = cumulative drift
K_xi_est <- as.integer(params$K_xi_est)

g_by_k <- sm %>%
  transmute(k = rank, g_k = mean_dlogw_s) %>%
  filter(k <= (K_xi_est + 1L)) %>%
  arrange(k)

S_by_k <- g_by_k %>%
  mutate(S_k = cumsum(g_k))

sigma_hat_by_k <- rho_by_k_med %>%
  inner_join(S_by_k %>% select(k, S_k), by = "k") %>%
  mutate(
    sigma2_hat = -2 * S_k * rho_k,
    sigma_hat = if_else(sigma2_hat > 0, sqrt(sigma2_hat), NA_real_)
  )

p_sigk <- ggplot(
  filter(sigma_hat_by_k, k >= params$bootstrap_k_min, !is.na(sigma_hat)),
  aes(x = k, y = sigma_hat)
) +
  geom_line(linewidth = 0.9) +
  scale_x_continuous(trans = "log10") +
  labs(x = "k", y = "sigma_hat") +
  theme_bw()

ggsave(file.path(out_dir, "sigma_hat_k.png"), p_sigk,
       width = 7.2, height = 5.6, dpi = 200)
```

```{r sigma-hat-ci}
bootstrap_sigma_period <- function(xi_df_period, S_by_k, B, k_min, K_xi_est) {
  weeks <- unique(xi_df_period$week)
  boot_stats <- numeric(B)

  for (b in seq_len(B)) {
    w_samp <- sample(weeks, size = length(weeks), replace = TRUE)

    rho_b <- xi_df_period %>%
      filter(week %in% w_samp, k <= (K_xi_est - 1L)) %>%
      group_by(k) %>%
      summarise(rho_k = median(xi, na.rm = TRUE), .groups = "drop")

    sig_b <- rho_b %>%
      inner_join(S_by_k %>% select(k, S_k), by = "k") %>%
      mutate(
        sigma2_hat = -2 * S_k * rho_k,
        sigma_hat = if_else(sigma2_hat > 0, sqrt(sigma2_hat), NA_real_)
      ) %>%
      filter(k >= k_min, k <= (K_xi_est - 1L))

    boot_stats[b] <- median(sig_b$sigma_hat, na.rm = TRUE)
  }

  tibble(
    sigma_hat_med = median(boot_stats, na.rm = TRUE),
    ci_lo = quantile(boot_stats, 0.025, na.rm = TRUE),
    ci_hi = quantile(boot_stats, 0.975, na.rm = TRUE)
  )
}

B <- if (isTRUE(getOption("knitr.in.progress"))) as.integer(params$bootstrap_B_fast) else as.integer(params$bootstrap_B_fast)

xi_weekly_q <- xi_weekly %>%
  mutate(period = floor_date(week, params$bootstrap_period_unit))

sigma_ci_by_period <- xi_weekly_q %>%
  group_by(period) %>%
  group_modify(~ bootstrap_sigma_period(
    .x,
    S_by_k,
    B = B,
    k_min = as.integer(params$bootstrap_k_min),
    K_xi_est = K_xi_est
  )) %>%
  ungroup()

p_ci <- ggplot(sigma_ci_by_period, aes(x = period, y = sigma_hat_med)) +
  geom_line() +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), alpha = 0.2) +
  labs(
    title = "Bootstrap CI for sigma_hat over time (quarterly)",
    x = NULL,
    y = "sigma_hat (median across k)"
  ) +
  theme_bw()

ggsave(file.path(out_dir, "sigma_hat_ci_time.png"), p_ci,
       width = 7.2, height = 5.6, dpi = 200)
```

# 4. Fast simulation outputs needed for the talk

This section produces the two simulated objects we show on slides:

1. Simulated CDC (compared to empirical)
2. Simulated micro movement for rank slots (used as a complement to the empirical cohort-tracking object)

```{r sim-minimal}
# Seed curve w0 up to K_max
w0_raw <- emp_cdc$w_bar[seq_len(min(K_max, nrow(emp_cdc)))]
w0_raw <- w0_raw / sum(w0_raw)

entrant_pool <- endpoint_weekly %>%
  filter(rank > K_cut, rank <= K_max) %>%
  transmute(share = share_global) %>%
  filter(is.finite(share), share > 0)

entrant_sampler <- function(n) {
  if (n <= 0L) return(numeric(0))
  sample(entrant_pool$share, size = n, replace = TRUE)
}

# Minimal simulation: accumulate CDC + store rank-slot series for a few k's
simulate_rank_paths_minimal <- function(
    w0, K_cut, K_max,
    T, n_paths,
    mean_vec, sd_vec,
    mu = 0, sigma = 1,
    entry_frac = 0.10,
    k_track = c(12L, 1L),
    horizons = c(1L, 4L, 12L)
) {
  stopifnot(length(mean_vec) == K_max, length(sd_vec) == K_max)
  k_track <- sort(unique(as.integer(k_track)))
  horizons <- sort(unique(as.integer(horizons)))

  cdc_sum <- numeric(K_cut)
  n_cdc <- 0L

  # rank-slot shares for tracked k
  slot <- array(NA_real_, dim = c(n_paths, T, length(k_track)))

  for (p in seq_len(n_paths)) {
    if (length(w0) < K_max) {
      w <- c(w0, entrant_sampler(K_max - length(w0)))
    } else {
      w <- w0[seq_len(K_max)]
    }
    w <- pmax(w, 1e-18)
    w <- w / sum(w)

    for (t in seq_len(T)) {
      dlogw <- rnorm(K_max, mean = mu + mean_vec, sd = sigma * sd_vec)
      w <- w * exp(dlogw)
      w <- pmax(w, 1e-18)
      w <- w / sum(w)

      # exit/entry: keep top K_cut, refill to K_max
      w_sorted <- sort(w, decreasing = TRUE)
      w_top <- w_sorted[seq_len(K_cut)]
      n_tail <- K_max - K_cut

      if (n_tail > 0) {
        n_add <- max(1L, floor(n_tail * entry_frac))
        w <- c(w_top, entrant_sampler(n_add), entrant_sampler(n_tail - n_add))
      } else {
        w <- w_top
      }
      w <- pmax(w, 1e-18)
      w <- w / sum(w)

      # top-K renormalized for CDC and micro slot tracking
      wK <- w[seq_len(K_cut)]
      wK <- wK / sum(wK)

      cdc_sum <- cdc_sum + wK
      n_cdc <- n_cdc + 1L

      for (j in seq_along(k_track)) {
        kk <- k_track[j]
        if (kk <= K_cut) slot[p, t, j] <- wK[kk]
      }
    }
  }

  # CDC mean
  cdc_mean <- cdc_sum / n_cdc

  # micro: slot-based future shares at t+h
  micro <- purrr::map_dfr(seq_along(k_track), function(j) {
    kk <- k_track[j]
    purrr::map_dfr(horizons, function(h) {
      vals <- as.vector(slot[, (1:(T - h)) + h, j])
      tibble(k = kk, horizon = h, share_tp = vals)
    })
  })

  list(cdc_mean = cdc_mean, micro = micro)
}

k_track <- sort(unique(c(params$k_focus, params$k_grid)))

sim_min <- simulate_rank_paths_minimal(
  w0 = w0_raw,
  K_cut = K_cut,
  K_max = K_max,
  T = as.integer(params$sim_T_weeks),
  n_paths = as.integer(params$sim_n_paths),
  mean_vec = sm$mean_dlogw_s,
  sd_vec = sm$sd_dlogw_s,
  mu = params$sim_mu,
  sigma = params$sim_sigma,
  entry_frac = params$sim_entry_frac,
  k_track = k_track,
  horizons = as.integer(params$horizons_micro)
)
```

```{r cdc-plot}
sim_cdc <- tibble(rank = 1:K_cut, w_bar = sim_min$cdc_mean)

cdc_plot_df <- bind_rows(
  emp_cdc %>% transmute(rank, w_bar, source = "Empirical"),
  sim_cdc %>% transmute(rank, w_bar, source = "Simulated")
)

p_cdc <- ggplot(cdc_plot_df, aes(x = rank, y = w_bar, linetype = source)) +
  geom_line(linewidth = 0.9) +
  scale_x_continuous(trans = "log10") +
  scale_y_continuous(trans = "log10") +
  labs(
    title = paste0("CDC: Empirical vs Simulated (sigma = ", params$sim_sigma, ")"),
    x = "Rank (log)",
    y = "Mean share (log)"
  ) +
  theme_bw() +
  theme(legend.position = "top")

ggsave(file.path(out_dir, "cdc_emp_vs_sim.png"), p_cdc,
       width = 7.2, height = 5.6, dpi = 200)
```

# 5. Micro validation objects (empirical cohort tracking vs simulated rank-slot)

```{r emp-micro}
emp_micro_for_k <- function(endpoint_weekly, k, horizons = c(1L, 4L, 12L)) {
  base <- endpoint_weekly %>%
    filter(rank == k) %>%
    transmute(
      week_t = week,
      endpoint_id,
      share_t = share_topK
    )

  out <- purrr::map_dfr(horizons, function(h) {
    fut <- endpoint_weekly %>%
      transmute(
        week_tp = week,
        endpoint_id,
        share_tp = share_topK
      )

    base %>%
      mutate(week_tp = week_t + weeks(h), horizon = h) %>%
      left_join(fut, by = c("week_tp", "endpoint_id")) %>%
      mutate(present_tp = !is.na(share_tp))
  })

  out
}

k_focus <- as.integer(params$k_focus)
horizons <- as.integer(params$horizons_micro)

emp_micro_k <- emp_micro_for_k(endpoint_weekly, k = k_focus, horizons = horizons)
```

```{r micro-plot}
# Empirical cohort summary (condition on still being present)
emp_summary_k <- emp_micro_k %>%
  group_by(horizon) %>%
  summarise(
    n_total = n(),
    retention = mean(present_tp),
    n_present = sum(present_tp),
    med = median(share_tp[present_tp], na.rm = TRUE),
    p10 = quantile(share_tp[present_tp], 0.10, na.rm = TRUE),
    p90 = quantile(share_tp[present_tp], 0.90, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(source = "empirical")

# Simulation rank-slot summary (always present by construction)
sim_micro_k <- sim_min$micro %>%
  filter(k == k_focus) %>%
  group_by(horizon) %>%
  summarise(
    n_total = n(),
    retention = 1.0,
    n_present = n(),
    med = median(share_tp, na.rm = TRUE),
    p10 = quantile(share_tp, 0.10, na.rm = TRUE),
    p90 = quantile(share_tp, 0.90, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(source = "simulation")

micro_summary_k <- bind_rows(emp_summary_k, sim_summary_k <- sim_micro_k)

p_micro <- ggplot(micro_summary_k, aes(x = horizon, y = med,
                                      group = source, linetype = source)) +
  geom_ribbon(aes(ymin = p10, ymax = p90, fill = source), alpha = 0.50, color = NA) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = horizons) +
  labs(
    title = paste0("Micro movement for rank k = ", k_focus),
    subtitle = "Median share at t+h with 10-90% band",
    x = "Horizon h (weeks ahead)",
    y = "Share at t+h"
  ) +
  theme_bw()

ggsave(file.path(out_dir, "micro_k12.png"), p_micro,
       width = 7.2, height = 5.6, dpi = 200)
```

```{r micro-rmse-by-rank}
# Repeat micro summaries across k_grid and compute RMSE over horizons
k_grid <- as.integer(params$k_grid)

micro_fit_for_k <- function(k_val) {
  emp <- emp_micro_for_k(endpoint_weekly, k = k_val, horizons = horizons) %>%
    group_by(horizon) %>%
    summarise(
      emp_retention = mean(present_tp),
      emp_med = median(share_tp[present_tp], na.rm = TRUE),
      .groups = "drop"
    )

  sim <- sim_min$micro %>%
    filter(k == k_val) %>%
    group_by(horizon) %>%
    summarise(sim_med = median(share_tp, na.rm = TRUE), .groups = "drop")

  joined <- emp %>%
    inner_join(sim, by = "horizon")

  rmse <- sqrt(mean((joined$emp_med - joined$sim_med)^2, na.rm = TRUE))
  tibble(k = k_val, rmse_med = rmse)
}

fit_rmse <- map_dfr(k_grid, micro_fit_for_k)

p_rmse <- ggplot(fit_rmse, aes(x = k, y = rmse_med)) +
  geom_line() +
  geom_point(size = 2) +
  scale_x_continuous(trans = "log10") +
  labs(
    title = "RMSE of median shares across horizons",
    subtitle = "Computed from empirical vs simulated medians over h = {1,4,12}",
    x = "Rank k (log scale)",
    y = "RMSE (median share)"
  ) +
  theme_bw()

ggsave(file.path(out_dir, "micro_rmse_by_rank.png"), p_rmse,
       width = 7.2, height = 5.6, dpi = 200)
```

# 6. Optional: sigma calibration curve (expensive)

This reproduces the spirit of v10 but is 
**intentionally disabled by default** because it can be expensive.

```{r sigma-calibration, eval=params$run_sigma_calibration}
# Placeholder: if enabled, implement a coarse grid scan using a reduced K_cut
# and reduced paths (or reuse cached results). Keep the output file name
# fixed to match the TeX deck.

stop("Sigma calibration scan is disabled by default. Set run_sigma_calibration: true and implement your preferred scan here.")
```

# 7. Outputs

The analysis writes the figures used by the beamer deck into:

- `r out_dir`

Expected filenames:

- `cdc_emp_vs_sim.png`
- `sigma_calibration.png` (from v10; regenerate only if desired)
- `xi_median_band.png`
- `xi_bucket_time.png`
- `sigma_hat_k.png`
- `sigma_hat_ci_time.png`
- `micro_k12.png`
- `micro_rmse_by_rank.png`


```{r ensure-sigma-calibration-figure}
# Ensure the TeX deck always has a sigma_calibration.png to include.
# If you want the true v10-style scan, set run_sigma_calibration: true above
# and replace the placeholder implementation in that chunk.

sigma_path <- file.path(out_dir, "sigma_calibration.png")

if (!file.exists(sigma_path)) {
  p_placeholder <- ggplot(tibble(x = 0, y = 0), aes(x, y)) +
    geom_blank() +
    geom_vline(xintercept = params$sim_sigma, linetype = "dashed", color = "red") +
    scale_x_continuous(limits = c(0.14, 0.42)) +
    labs(
      title = "Sigma calibration overall (placeholder)",
      subtitle = paste0("Using fixed sigma = ", params$sim_sigma,
                        "; set run_sigma_calibration: true to scan"),
      x = "sigma",
      y = "score"
    ) +
    theme_bw()

  ggsave(sigma_path, p_placeholder, width = 8, height = 5, dpi = 200)
}
```


## Provenance

```{r provenance, echo=FALSE}
cat("Rendered at:", as.character(Sys.time()), "\n\n")

git_hash <- tryCatch(system("git rev-parse HEAD", intern = TRUE), error = function(e) NA_character_)
if (!is.na(git_hash) && length(git_hash) > 0) {
  cat("Git commit:", git_hash, "\n")
}
```
