---
title: "Rank Diffusion and Simulation of FB Data (v4)"
format: pdf
execute:
  echo: true
  warning: false
  message: false
---

## Introduction

This script builds on the simple simulation from v4 by 1) considering the full data set (5563 max ranks observed), adn then 2) conduct a grid search over varying growth rates to attempt to reconstruct the empirical distribution as closely as possible. 

First, read in the data and load some libraries as before. 

```{r}
library(tidyverse)
library(arrow)
library(scales)
library(here)
library(lubridate)

set.seed(1823)

raw <- read_parquet(
  here::here("/Users/philipwaggoner/Dropbox (Personal)/GW - IDDP/Projects/Rank Diffusion (Matt)/Facebook/fb_top2000_ranked_daily.parquet")
)
```

## Construct endpoint-level weekly series

```{r}
endpoint_weekly <- raw |>
  mutate(week = floor_date(date, unit = "week", week_start = 1)) |>
  group_by(week, endpoint_id) |>
  summarise(metric_value = sum(metric_value, na.rm = TRUE), .groups = "drop") |>
  group_by(week) |>
  arrange(desc(metric_value), .by_group = TRUE) |>
  mutate(
    rank  = row_number(),
    share = metric_value / sum(metric_value, na.rm = TRUE)
    ) |>
  ungroup()
```

In the above chunk we are collapsing the raw post-level engagement into endpoint-level weekly totals. Then, we rank endpoints within each week (rank = row_number()), with rank 1 being the most engaged endpoint that week. Then, compute weekly share, which equals endpoint's `metric_value / total metric_value` across all endpoints for that week.

This gives a consistent panel with one row per endpoint per week, with both rank and relative share.

For the sim, then, we have `r summary(endpoint_weekly$rank)[6]` unique ranks to work with. 

So, now, we set the baseline curve, which is the mean share by rank: 

```{r}
mean_share <- endpoint_weekly |>
  group_by(rank) |>
  summarise(w_bar = mean(share, na.rm = TRUE), .groups = "drop") |>
  arrange(rank)
```

From here, define the rank-wise delta log(share) pools (weekly):

```{r}
rank_panel <- endpoint_weekly |>
  group_by(rank) |>
  arrange(week, .by_group = TRUE) |>
  mutate(
    share = pmax(share, 1e-12),
    log_w = log(share),
    dlogw = log_w - lag(log_w)
  ) |>
  ungroup() |>
  filter(is.finite(dlogw))
```

## Setup for simulation

Now, 1) allow a large top-K cutoff to essentially include the entire sample, and then 2) build entrant pools for entry/exit. After this, we can update the CDC plot for comparison as in v3. 

```{r}
K_target_cut <- 5000L
K_target_max <- 6000L # max simulated "universe" size (top-K we keep + extra tail)
buffer_tail <- 2000L # how far below the cutoff to "harvest" entrants
T_weeks <- 52
N_paths <- 250

# limits given the data
max_rank_seen <- max(endpoint_weekly$rank, na.rm = TRUE)
K_cut  <- min(K_target_cut,  max_rank_seen)
K_max  <- min(K_target_max,  max_rank_seen + buffer_tail)

# baseline w0 (length = min(K_max, nrow(mean_share)))
w0_raw <- mean_share$w_bar[seq_len(min(K_max, nrow(mean_share)))]
w0_raw <- w0_raw / sum(w0_raw)

# tail entrant pool: ranks just below the cutoff (or as deep as available)
entrant_pool <- endpoint_weekly |>
filter(rank > K_cut, 
       rank <= min(K_cut + buffer_tail, max_rank_seen)) |>
  transmute(share) |>
  filter(is.finite(share), share > 0)

entrant_sampler <- function(n) sample(entrant_pool$share, 
                                      size = n, 
                                      replace = TRUE)

# rank-wise increment lists
increments_by_rank <- rank_panel |>
  filter(rank <= K_max) |>
  group_by(rank) |>
  summarise(increments = list(dlogw), .groups = "drop") |>
  arrange(rank)

# sample one delta log(share) per rank, with a scale on volatility (sigma_scale); come back to this later
sample_increments <- function(sigma_scale = 1) {
  vapply(increments_by_rank$increments, function(x) {
    if (length(x) == 0) return(0)
    sigma_scale * sample(x, size = 1, replace = TRUE)
    }, numeric(1))
}
```

## Simulation

This version updates v3 subtly. Steps include: 1) simulate K_max rank slots each step, 2) renormalize across K_max, 3) keep top K_cut as the tracked set, and 4) record snapshots of the top-K distribution at each time for CDC plot.

```{r}
simulate_rank_paths <- function(
    w0, K_cut, K_max, T = 52, n_paths = 200,
    sigma_scale = 1, entry_frac = 0.10) {
  
  stopifnot(length(w0) <= K_max, K_max >= K_cut)
  horizons <- c(1, 7, 28)
  growth_collect <- vector("list", length(horizons))
  names(growth_collect) <- paste0("h", horizons)
  snapshots <- vector("list", n_paths)
  final_top <- matrix(NA_real_, nrow = n_paths, ncol = K_cut)

top_n_sorted <- function(w, n) { w[order(w, decreasing = TRUE)][seq_len(n)] }

for (p in seq_len(n_paths)) {
  if (length(w0) < K_max) {
    w <- c(w0, entrant_sampler(K_max - length(w0)))
    w <- w / sum(w)
    } else {
      w <- w0[1:K_max]; w <- w / sum(w)
      }
  
  snaps_p <- vector("list", T)

  for (t in seq_len(T)) {
    dlogw <- sample_increments(sigma_scale = sigma_scale)
    if (length(dlogw) != length(w)) dlogw <- rep_len(dlogw, length.out = length(w))

    w <- w * exp(dlogw)
    w <- pmax(w, 1e-18)
    w <- w / sum(w) # normalize across K_max

  # keep top-K_cut, then refill tail to K_max with entrants
  w_top <- top_n_sorted(w, K_cut)
  n_tail <- K_max - K_cut
  if (n_tail > 0) {
    n_add <- max(1L, floor(n_tail * entry_frac))
    w <- c(w_top, entrant_sampler(n_add), entrant_sampler(n_tail - n_add))
  } else {
    w <- w_top
  }
  w <- w / sum(w)

  # CDC snapshot at time t: top-K_cut, renormalized within K_cut
  wK <- top_n_sorted(w, K_cut)
  wK <- wK / sum(wK)
  snaps_p[[t]] <- tibble(path = p, t = t, rank = seq_len(K_cut), share = wK)

  # growth (relative to start) at horizons
  if (t %in% horizons) {
    w0K <- top_n_sorted(w0, K_cut); w0K <- w0K / sum(w0K)
    lg  <- log(wK / w0K)
    growth_collect[[paste0("h", t)]] <- rbind(
      growth_collect[[paste0("h", t)]],
      cbind(path = p, rank = seq_len(K_cut), horizon = t, log_growth = lg)
    )
  }
}

  final_top[p, ] <- top_n_sorted(w, K_cut)
  snapshots[[p]] <- bind_rows(snaps_p)

}

  list(
    # n_paths x K_cut
    final = final_top,
    # snapshots for CDC
    snapshots = bind_rows(snapshots),  
    # time horizons
    growth = bind_rows(growth_collect)
  )
}
```

## Run sim at K and compare CDC visually

```{r eval = FALSE}
# build an initial w0 for K_max
if (length(w0_raw) < K_max) {
  w0_ext <- c(w0_raw, entrant_sampler(K_max - length(w0_raw)))
  w0_ext <- w0_ext / sum(w0_ext)
  } else {
    w0_ext <- w0_raw[1:K_max]; w0_ext <- w0_ext / sum(w0_ext)
    }

set.seed(123)

sim_v4 <- simulate_rank_paths(
  w0 = w0_ext, K_cut = K_cut, K_max = K_max,
  T = T_weeks, n_paths = N_paths, entry_frac  = 0.10
)

# empirical CDC (top-K_cut, normalized within K)
emp_cdc <- endpoint_weekly |>
  group_by(week) |>
  arrange(desc(share), .by_group = TRUE) |>
  mutate(rank_in_week = row_number()) |>
  filter(rank_in_week <= K_cut) |>
  mutate(shareK = share / sum(share)) |>
  ungroup() |>
  group_by(rank = rank_in_week) |>
  summarise(w_bar = mean(shareK, na.rm = TRUE), 
            .groups = "drop")

# sim CDC
sim_cdc <- sim_v4$snapshots |>
  group_by(rank) |>
  summarise(w_bar = mean(share, na.rm = TRUE), 
            .groups = "drop")

write_rds(sim_v4, here::here("Desktop", "sim_v4.rds"))
write_rds(emp_cdc, here::here("Desktop", "emp_cdc.rds"))
write_rds(sim_cdc, here::here("Desktop", "sim_cdc.rds"))
```

Now visualize: 

```{r}

emp_cdc <- read_rds(here::here("/Users/philipwaggoner/Dropbox (Personal)/GW - IDDP/Projects/Rank Diffusion (Matt)/Facebook/shared w Matt/v4/emp_cdc.rds"))
sim_cdc <- read_rds(here::here("/Users/philipwaggoner/Dropbox (Personal)/GW - IDDP/Projects/Rank Diffusion (Matt)/Facebook/shared w Matt/v4/sim_cdc.rds"))

# compare CDC for empirical and simulated
bind_rows(
  emp_cdc |> mutate(source = "Empirical"),
  sim_cdc |> mutate(source = "Simulated")
  ) |>
  ggplot(aes(rank, w_bar, linetype = source)) +
  geom_line() +
  scale_x_log10() +
  scale_y_log10(labels = percent_format(accuracy = 0.01)) +
  labs(
    title = paste0("CDC: Empirical (mean share) vs Sim (K_cut=", K_cut, ", K_max=", K_max, ")"),
    x = "Rank (log scale)", 
    y = "Mean share (log scale)", 
    linetype = NULL
  ) +
  theme_minimal()
```

## Grid Search

Next, let's conduct a grid search over growth rates. We will vary the `sigma_scale` parameter, which multiplies every rank’s change in log(share) draw. The idea is:

  - $<$ 1 shrinks volatility (flatter CDC tail)
  - $>$ 1 inflates volatility (heavier tail)

For each of assessment, we'll compute RMSE in log-share space, and then plot all curves against each other for direct comparison. 

```{r eval = FALSE}
# distance metric between two CDC curves (log-space RMSE)
cdc_rmse <- function(emp, sim) {
  df <- emp |> 
    inner_join(sim, 
               by = "rank", 
               suffix = c("_emp", "_sim"))
  sqrt(mean((log(df$w_bar_emp) - log(df$w_bar_sim))^2, na.rm = TRUE))
}

# try out a grid of sigma scales
sigma_grid <- c(0.6, 0.75, 0.9, 1.0, 1.1, 1.25, 1.5)

results <- vector("list", length(sigma_grid))

for (i in seq_along(sigma_grid)) {
  s <- sigma_grid[i]
  set.seed(500 + i)
  sim_i <- simulate_rank_paths(
    w0 = w0_ext, K_cut = K_cut, K_max = K_max,
    T = T_weeks, n_paths = N_paths,
    sigma_scale = s, entry_frac = 0.10
    )

  sim_cdc_i <- sim_i$snapshots |>
    group_by(rank) |>
    summarise(w_bar = mean(share, na.rm = TRUE), 
              .groups = "drop")

  rmse_i <- cdc_rmse(emp_cdc, sim_cdc_i)
  
  results[[i]] <- list(scale = s, 
                       cdc = sim_cdc_i, 
                       rmse = rmse_i)
}

grid_tbl <- map_dfr(results, ~ .x$cdc |> 
                      mutate(sigma_scale = .x$scale))

rmse_tbl <- tibble(
  sigma_scale = map_dbl(results, "scale"),
  rmse = map_dbl(results, "rmse")
) |>
  arrange(rmse)

best_sigma <- rmse_tbl$sigma_scale[1]

#best_sigma
#rmse_tbl
```

```{r eval = FALSE, echo = FALSE}
write_rds(rmse_tbl, here::here("Desktop", "rmse_tbl.rds"))
write_rds(grid_tbl, here::here("Desktop", "grid_tbl.rds"))
write_rds(results, here::here("Desktop", "results.rds"))
```

Now, visualize: 

```{r}
results <- read_rds(here::here("/Users/philipwaggoner/Dropbox (Personal)/GW - IDDP/Projects/Rank Diffusion (Matt)/Facebook/shared w Matt/v4/results.rds"))
grid_tbl <- read_rds(here::here("/Users/philipwaggoner/Dropbox (Personal)/GW - IDDP/Projects/Rank Diffusion (Matt)/Facebook/shared w Matt/v4/grid_tbl.rds"))
rmse_tbl <- read_rds(here::here("/Users/philipwaggoner/Dropbox (Personal)/GW - IDDP/Projects/Rank Diffusion (Matt)/Facebook/shared w Matt/v4/rmse_tbl.rds"))

best_sigma <- 0.6

emp_line <- emp_cdc |> mutate(source = "Empirical")

sims_all <- grid_tbl |>
  mutate(source = paste0("Sim sigma =", sigma_scale))

ggplot() +
  geom_line(data = sims_all, aes(rank, w_bar, group = source),
            color = "grey70", alpha = 0.7) +
  geom_line(
    data = filter(sims_all, sigma_scale == best_sigma),
    aes(rank, w_bar), color = "steelblue", linewidth = 1.1) +
  geom_line(data = emp_line, aes(rank, w_bar), color = "black") +
  scale_x_log10() +
  scale_y_log10(labels = percent_format(accuracy = 0.01)) +
  labs(
    title = paste0("CDC grid search (K_cut=", K_cut, ", K_max=", K_max, ")"),
    subtitle = paste0("Blue = best fit from sims (sigma = 0.6); black = empirical"),
    x = "Rank (log scale)", 
    y = "Mean share (log scale)") +
  theme_minimal()
```

So from this we can see the grid search found that scaling empirical growth by `sigma = 0.6` gives the most realistic long-run capital distribution. This means rank diffusion on this platform is moderately damped, i.e., about 60% as volatile as unconstrained random diffusion. **Is this a correct understanding?**

If so, here's how to unpack it a little more. When the simulated log-share increments were scaled to 60% of their original empirical variance, the resulting CDC best matched the real Facebook data. So, in other words, the rank diffusion model reproduces the long-run shape of the empirical power-law best when we "dampen" the volatility by about 40%. If correct, this implies at least 2 things:

  1. The raw empirical increment distribution, if applied directly, over-diffuses, s.t. it erodes the heavy tail too quickly.
  
  2. But applying a scaling of `sigma = 0.6` stabilizes the system, so to speak, letting the simulated "capital"/share concentrate near the top ranks in a way consistent with reality/Facebook data

