---
title: "Rank Diffusion and Simulation of FB Data (v6)"
format: pdf
execute:
  echo: true
  warning: false
  message: false
---

## Introduction

This version switches to the weekly-organized file and keeps all rows per weekly snapshot (no per-week de-duplication). We first confirm coverage and weekly scale (~14k rows per week), then run our basic entry/exit simulation and compare the simulated capital-distribution curve (CDC) to the empirical CDC. From there, we can make decisions about a good grid search (e.g., over mean and variances, etc).

## Setup

```{r}
library(tidyverse)
library(arrow)
library(scales)
library(here)
library(lubridate)
library(glue)

set.seed(1823)

raw <- read_parquet(
  here::here("/Users/philipwaggoner/Dropbox (Personal)/GW - IDDP/Projects/Rank Diffusion (Matt)/Facebook/fb_ranked_weekly.parquet")
)

glimpse(raw)
```

## Weekly coverage & scale checks

The weekly file represents weekly snapshots. But beofre going further, we should validate the span and the weekly row counts (which should align with ~14k rows on typical weeks).

```{r}
# date span and overall distinct pages seen
raw %>%
  summarise(
    min_date = min(date),
    max_date = max(date),
    n_pages  = n_distinct(endpoint_id),
    weeks    = n_distinct(date)
    )
```

```{r}
# weekly totals (no de-dup per week; i.e., each row = one endpoint's weekly record)

weekly_totals <- raw %>%
count(week = date, name = "n_rows") %>%
arrange(week)

summary(weekly_totals$n_rows)
head(weekly_totals, 5); tail(weekly_totals, 5)

ggplot(weekly_totals, aes(week, n_rows)) +
geom_line() +
scale_y_continuous(labels = comma) +
labs(
title = "Weekly row counts (no de-duplication, weekly snapshots)",
x = "Week", y = "Rows"
) +
theme_bw()
```

Note: in the weekly file, each row is already one endpoint's weekly record, so "unique per week" and "all rows per week" (per earlier versions of this script) are effectively the same number.

## Build endpoint-weekly series for CDC & simulation

We re-compute share within each week.

```{r}

endpoint_weekly <- raw %>%
  transmute(
    week = date,
    endpoint_id,
    metric_value,
    rank = rank
  ) %>%
  group_by(week) %>%
  arrange(desc(metric_value), .by_group = TRUE) %>%
  mutate(
    rank  = row_number(),
    share = metric_value / sum(metric_value, na.rm = TRUE)
  ) %>%
  ungroup()

# per-week observed rank maxima 
# (should go well above 10k on many weeks; 
# again, counter to what we were previously seeing with the day-level data)

max_rank_by_week <- endpoint_weekly %>%
  group_by(week) %>%
  summarise(max_rank = max(rank), .groups = "drop")

max_rank_seen <- max(max_rank_by_week$max_rank, na.rm = TRUE)
max_rank_seen
```


## Empirical CDC (top-K mean share)

Empirical CDC here means the mean (across weeks) of the top-K shares, normalized within the top-K each week.

```{r}
K_cut <- min(12000L, max_rank_seen)

emp_cdc <- endpoint_weekly %>%
  group_by(week) %>%
  arrange(desc(share), .by_group = TRUE) %>%
  mutate(rank_in_week = row_number()) %>%
  filter(rank_in_week <= K_cut) %>%
  mutate(shareK = share / sum(share)) %>%
  ungroup() %>%
  group_by(rank = rank_in_week) %>%
  summarise(w_bar = mean(shareK, na.rm = TRUE), .groups = "drop")

head(emp_cdc, 10)
```

## Simulation (entry/exit, weekly increments)

We'll seed with the empirical mean share by rank, define rank-wise increment pools (weekly delta_log share), then simulate K_max slots with entry/exit and compare the CDC.

```{r}
# baseline seed curve is mean share by rank across weeks
mean_share <- endpoint_weekly %>%
  group_by(rank) %>%
  summarise(w_bar = mean(share, na.rm = TRUE), .groups = "drop") %>%
  arrange(rank)

# increment pools: per-rank weekly delta_log(share)
rank_panel <- endpoint_weekly %>%
  group_by(rank) %>%
  arrange(week, .by_group = TRUE) %>%
  mutate(
    share = pmax(share, 1e-12),
    log_w = log(share),
    dlogw = log_w - lag(log_w)
  ) %>%
  ungroup() %>%
  filter(is.finite(dlogw))

# universe size (top-K we track; K_max adds tail for entry/exit)
buffer_tail <- 2000L
K_max <- min(K_cut + buffer_tail, max_rank_seen)

w0_raw <- mean_share$w_bar[seq_len(min(K_max, nrow(mean_share)))]
w0_raw <- w0_raw / sum(w0_raw)

# entrant pool: ranks just below the top-K_cut
entrant_pool <- endpoint_weekly %>%
  filter(rank > K_cut, rank <= min(K_cut + buffer_tail, max_rank_seen)) %>%
  transmute(share) %>%
  filter(is.finite(share), share > 0)

entrant_sampler <- function(n) {
  sample(entrant_pool$share, size = n, replace = TRUE)
}

# rank-wise increment lists
increments_by_rank <- rank_panel %>%
  filter(rank <= K_max) %>%
  group_by(rank) %>%
  summarise(increments = list(dlogw), .groups = "drop") %>%
  arrange(rank)

sample_increments <- function(sigma_scale = 1) {
  vapply(increments_by_rank$increments, function(x) {
    if (length(x) == 0) return(0)
    sigma_scale * sample(x, size = 1, replace = TRUE)
  }, numeric(1))
}

simulate_rank_paths <- function(
  w0, K_cut, K_max, T = 52, n_paths = 150,
  sigma_scale = 1.0, entry_frac = 0.10
  ) {
  stopifnot(length(w0) <= K_max, K_max >= K_cut)

  top_n_sorted <- function(w, n) w[order(w, decreasing = TRUE)][seq_len(n)]

  snapshots <- vector("list", n_paths)

  for (p in seq_len(n_paths)) {
    if (length(w0) < K_max) { # initialize to full universe length K_max
    w <- c(w0, entrant_sampler(K_max - length(w0)))
    w <- w / sum(w)
    } else {
      w <- w0[1:K_max]; w <- w / sum(w)
      }

  snaps_p <- vector("list", T)

for (t in seq_len(T)) {
  dlogw <- sample_increments(sigma_scale = sigma_scale)
  if (length(dlogw) != length(w)) dlogw <- rep_len(dlogw, length.out = length(w))

  w <- w * exp(dlogw)
  w <- pmax(w, 1e-18)
  w <- w / sum(w)

  # keep top-K_cut, refill tail from entrant pool to K_max
  w_top  <- top_n_sorted(w, K_cut)
  n_tail <- K_max - K_cut
  if (n_tail > 0) {
    n_add <- max(1L, floor(n_tail * entry_frac))
    w <- c(w_top, entrant_sampler(n_add), entrant_sampler(n_tail - n_add))
  } else {
    w <- w_top
  }
  w <- w / sum(w)

  # CDC snapshot: top-K_cut renormalized within K_cut
  wK <- top_n_sorted(w, K_cut)
  wK <- wK / sum(wK)
  snaps_p[[t]] <- tibble(path = p, t = t, rank = seq_len(K_cut), share = wK)
}

snapshots[[p]] <- bind_rows(snaps_p)

}

list(snapshots = bind_rows(snapshots))
}
```

## Run a single baseline sim for CDC comparison

```{r, echo = TRUE, eval = FALSE}
T_weeks <- 52
N_paths <- 150

# extend w0 to K_max once
if (length(w0_raw) < K_max) {
  w0_ext <- c(w0_raw, entrant_sampler(K_max - length(w0_raw)))
  w0_ext <- w0_ext / sum(w0_ext)
  } else {
    w0_ext <- w0_raw[1:K_max]; w0_ext <- w0_ext / sum(w0_ext)
    }

set.seed(123)

sim_out <- simulate_rank_paths(
  w0 = w0_ext, K_cut = K_cut, K_max = K_max,
  T = T_weeks, n_paths = N_paths,
  sigma_scale = 0.1, entry_frac = 0.10
  )
```

```{r, echo = TRUE, eval = FALSE}
# sim CDC: average across snapshots and paths
sim_cdc <- sim_out$snapshots %>%
  group_by(rank) %>%
  summarise(w_bar = mean(share, na.rm = TRUE), .groups = "drop")
```

```{r, echo = FALSE, eval = FALSE}
write_rds(emp_cdc, here::here("Desktop", "emp_cdc.rds"))
write_rds(sim_cdc, here::here("Desktop", "sim_cdc.rds"))
```

```{r, echo = FALSE, eval = TRUE}
emp_cdc <- read_rds("/Users/philipwaggoner/Desktop/emp_cdc.rds")
sim_cdc <- read_rds("/Users/philipwaggoner/Desktop/sim_cdc.rds")
```

Now, visualize the empirical vs sim comparison:

```{r}
#| fig-width: 8
#| fig-height: 6

bind_rows(
  emp_cdc %>% mutate(source = "Empirical"),
  sim_cdc %>% mutate(source = "Simulated")
  ) %>%
  ggplot(aes(rank, w_bar, linetype = source)) +
  geom_line() +
  scale_x_log10() +
  scale_y_log10(labels = percent_format(accuracy = 0.01)) +
  labs(
    title = glue("Capital Distribution Curve (CDC): Empirical vs Simulated (K_cut={K_cut}, K_max={K_max})"),
    x = "Rank (log scale)",
    y = "Mean share within top-K (log scale)",
    linetype = NULL
  ) +
  theme_bw()
```

Using the weekly file resolves the earlier 5,563 cap we saw. That's great news!

The simulation's current parameters are:

- `sigma_scale` (volatility on delta_log(share)); currently set at 0.1, and gets near perfect match (earlier testing showed the higher the value, the worse the fit gets, e.g., 1.0 looked like the old, poorly fitting day-based data simulation)
- `entry_frac` (how much of the tail is comprised of "fresh entrants" each step)
- `K_cut`, `K_max` (tracked top-K and universe size)

So, from here we could, e.g., run a grid over `sigma_scale` and a mean-shift by rank, or test lag structure as previously discussed (e.g., martingale diagnostics), and/or explicitly model downward jump components if the tails remain too heavy relative to empirical. A lot of options how that a key data QA step is addressed and the CDC comparison looks more reasonable. 
