---
title: "Rank Diffusion and Simulation of FB Data (v2)"
format: pdf
execute:
  echo: true
  warning: false
  message: false
---

## Introduction

In this script we walk through the steps for moving from raw Facebook post-level engagement data to endpoint-level time series (calculating growth rate distributions across multiple time horizons), and offer a first attempt at simulating rank diffusion based on empirical patterns/data. 

## Setup

```{r}
library(tidyverse)
library(arrow)
library(scales)
library(here)

set.seed(1823)

raw <- read_parquet(here("/Users/philipwaggoner/Dropbox (Personal)/GW - IDDP/Projects/Rank Diffusion (Matt)/Facebook/fb_top2000_ranked_daily.parquet"))

glimpse(raw)
```

## Step 1: Construct endpoint-level weekly series

```{r}
endpoint_weekly <- raw |>
  mutate(week = floor_date(date, unit = "week", week_start = 1)) |>
  group_by(week, endpoint_id) |>
  summarise(metric_value = sum(metric_value, na.rm = TRUE), .groups = "drop") |>
  group_by(week) |>
  arrange(desc(metric_value), .by_group = TRUE) |>
  mutate(
    rank = row_number(),
    share = metric_value / sum(metric_value, na.rm = TRUE)
  ) |>
  ungroup()

head(endpoint_weekly)
```

In the above chunk we are collapsing the raw post-level engagement into endpoint-level weekly totals. Then, we rank endpoints within each week (rank = row_number()), with rank 1 being the most engaged endpoint that week. Then, compute weekly share, which equals endpoint's `metric_value / total metric_value` across all endpoints for that week.

This gives a consistent panel with one row per endpoint per week, with both rank and relative share.

## Step 2: Compute forward-looking log growth

```{r}
endpoint_panel <- endpoint_weekly |>
  group_by(endpoint_id) |>
  arrange(week, .by_group = TRUE) |>
  mutate(
    log_growth_1w  = log(lead(share, 1)  / share),
    log_growth_7w  = log(lead(share, 7)  / share),
    log_growth_28w = log(lead(share, 28) / share)
  ) |>
  ungroup()

head(endpoint_panel)
```

Now, within each endpoint's time series, compute forward growth rates at horizons of 1, 7, and 28 weeks *ahead*, and express growth in log differences. These values are meant to capture volatility and drift in relative shares.

## Step 3: Categorize Endpoints by Size

```{r}
endpoint_panel <- endpoint_panel |>
  mutate(
    rank_group = case_when(
      rank <= 25 ~ "large",
      rank <= 250 ~ "midsize",
      TRUE ~ "small"
    )
  )

head(endpoint_panel)
```

Now, endpoints are bucketed by rank:

  - Top 25 = `large`
  - 26–250 = `midsize`
  - 251+ = `small`

This categorization is a bit arbitrary at the moment, but is meant to mirror the categorization idea from the book, which is useful for analyzing differences in volatility by some scale.

Now, take a look at the growth distributions: 

```{r}
#| fig-width: 8
#| fig-height: 6

plot_df <- endpoint_panel |>
  pivot_longer(
    cols = starts_with("log_growth_"),
    names_to = "horizon",
    values_to = "log_growth"
  ) |>
  mutate(
    horizon = recode(horizon,
                     log_growth_1w = "1 Week",
                     log_growth_7w = "7 Week",
                     log_growth_28w = "28 Week"),
    horizon = factor(horizon, levels = c("1 Week", "7 Week", "28 Week"))
  ) |>
  filter(!is.na(log_growth), !is.infinite(log_growth))

ggplot(plot_df, aes(x = log_growth, fill = rank_group, col = rank_group)) +
  geom_density(alpha = 0.3) +
  facet_wrap(~ horizon, scales = "fixed") +
  labs(
    title = "Distribution of log growth rates by rank group",
    x = "Log growth rate",
    y = "Density",
    fill = "Rank group", color = "Rank group"
  ) +
  theme_bw() +
  theme(legend.position = "bottom")
```

From here, we get:

  - At 1-week horizon: distributions are tight, centered near zero
  - At longer horizons: distributions widen reflecting potentially higher volatility

\newpage

## Simulation

Now, we will make a first attempt at simulating rank diffusion using the empirical weekly increments data.

### Seed simulation with mean shares

```{r}
mean_share <- endpoint_weekly |>
  group_by(rank) |>
  summarise(w_bar = mean(share, na.rm = TRUE), 
            .groups = "drop") |>
  arrange(rank)
```

From this code, we are computing average weekly share for each rank, which gives a baseline distribution (`w0`) from which simulations start.

### Construct increment pools

```{r}
rank_panel <- endpoint_weekly |>
  group_by(rank) |>
  arrange(week, .by_group = TRUE) |>
  mutate(
    share = pmax(share, 1e-12),
    log_w = log(share),
    dlogw = log_w - lag(log_w)
  ) |>
  ungroup() |>
  filter(!is.na(dlogw), is.finite(dlogw))
```

Now, for each rank position, compute weekly change in log share (`dlogw`), which forms the empirical pool of increments we will sample from. Also adding a small epsilon (`pmax(share, 1e-12)`) to prevent issues with zeros.

### Choose K, build w₀, increment pools, and a sampler

```{r}
K <- min(500, nrow(mean_share))
ranks <- 1:K

w0 <- mean_share$w_bar[ranks]
w0 <- w0 / sum(w0)

increments_by_rank <- rank_panel |>
  filter(rank %in% ranks, is.finite(dlogw)) |>
  group_by(rank) |>
  summarise(increments = list(dlogw), .groups = "drop") |>
  arrange(rank)

# note: draw one delta_log(share) for every rank slot (length K) for a simulated week
sample_increments <- function() {
  vapply(increments_by_rank$increments, function(x) {
    if (length(x) == 0) return(0)
    sample(x, size = 1, replace = TRUE) # nonparametric bootstrap per Matt suggestion
  }, numeric(1))
}
```

- `w_0` anchors all paths at a realistic starting curve

- `increments_by_rank` preserves rank-specific volatility structure

- `sample_increments()` is the core nonparametric resampling step

### Define sim

```{r}
simulate_rank_paths <- function(w0, K, T = 52, n_paths = 200, block = 1) {
  paths_final <- matrix(NA_real_, nrow = n_paths, ncol = K)
  horizons <- c(1, 7, 28)  # in weeks (was days originally)
  growth_collect <- vector("list", length(horizons))
  names(growth_collect) <- paste0("h", horizons)

  for (p in seq_len(n_paths)) {
    w <- w0
    for (t in seq_len(T)) {
      dlogw <- sample_increments()
      w <- w * exp(dlogw)
      w <- w / sum(w)
      if (t %in% horizons) {
        log_growth <- log(w / w0)
        growth_collect[[paste0("h", t)]] <- rbind(
          growth_collect[[paste0("h", t)]],
          cbind(path = p, 
                rank = ranks,
                horizon = t, 
                log_growth = log_growth)
        )
      }
    }
    paths_final[p, ] <- w
  }

  list(final = paths_final, 
       growth = bind_rows(growth_collect))
}
```

This function simulates `n_paths` independent evolutions of the rank-share distribution. So, for each week, `block`, (or some other time interval), sample increments (`dlogw`) from historical pools. Then, update shares (`w * exp(dlogw)`), then renormalize to ensure shares sum to 1. Collect growth outcomes at horizons of 1, 7, and 28 weeks in the future. From this we get shares and a dataset of growth results.

### Run sim

```{r}
set.seed(42)

sim <- simulate_rank_paths(w0 = w0, 
                           K = K, 
                           T = 52, 
                           n_paths = 300, 
                           block = 1)
```

### Simulated growth

```{r}
extract_horizon <- function(growth_df, col_name, label) {
  mat <- growth_df[[col_name]]
  df  <- as_tibble(mat)

  req <- c("rank", "log_growth")
  if (!all(req %in% names(df))) {
    stop("Horizon ", col_name, " is missing required columns: ",
         paste(setdiff(req, names(df)), collapse = ", "))
  }

  # build a path vector (if missing, fill NAs of correct length)
  path_vec <- if ("path" %in% names(df)) as.integer(df$path) else rep(NA_integer_, nrow(df))

  tibble(
    path       = path_vec,
    rank_slot  = as.integer(df$rank),
    horizon    = factor(label, levels = c("1 Week","7 Week","28 Week")),
    log_growth = as.numeric(df$log_growth)
  )
}

sim_growth <- bind_rows(
  extract_horizon(sim$growth, "h1",  "1 Week"),
  extract_horizon(sim$growth, "h7",  "7 Week"),
  extract_horizon(sim$growth, "h28", "28 Week")
) |>
  mutate(
    rank_group = case_when(
      rank_slot <= 25  ~ "large",
      rank_slot <= 250 ~ "midsize",
      TRUE             ~ "small"
    ),
    rank_group = factor(rank_group, levels = c("large","midsize","small"))
  )
```

Now, we collect simulated outcomes for horizons 1, 7, 28 weeks, and assign each rank slot to a rank group. Quick note note the structure: we grab the horizon-specific matrix (`growth_df[[col_name]]`) and convert it into a df. Output includes:

  - `path` (simulation path ID)
  - `rank_slot` (the rank position)
  - `horizon` (e.g., "1 Week")
  - `log_growth` (the simulated growth value)

Now, take a look at simulated distributions:

```{r}
#| fig-width: 8
#| fig-height: 6

ggplot(sim_growth, aes(x = log_growth, color = rank_group, 
                       fill = rank_group)) +
  geom_density(alpha = 0.25) +
  facet_wrap(~ horizon, scales = "fixed") +
  labs(
    title = "Simulated log growth distributions by rank group & time",
    subtitle = "Fixed scales for direct comparison",
    x = "Log growth", 
    y = "Density",
    color = "Rank group", 
    fill = "Rank group"
  ) +
  theme_bw() +
  theme(legend.position = "bottom")
```


```{r}
#| fig-width: 8
#| fig-height: 6

ggplot(sim_growth, aes(x = log_growth, color = rank_group, 
                       fill = rank_group)) +
  geom_density(alpha = 0.25) +
  facet_wrap(~ horizon, scales = "free_y") +
  labs(
    title = "Simulated log growth distributions by rank group & time",
    subtitle = "Variable scales for easier viewing of nuance",
    x = "Log growth", 
    y = "Density",
    color = "Rank group", 
    fill = "Rank group"
  ) +
  theme_bw() +
  theme(legend.position = "bottom")
```


From this:

  - At 1-week: sharp spike near 0 (i.e., most weeks involve tiny changes)
  - At 7-week and 28-week: distributions get wider
