---
title: "Rank Diffusion and Simulation of FB Data (v3)"
format: pdf
execute:
  echo: true
  warning: false
  message: false
---

## Introduction

This script builds on the simple simulation from v2:

1. Aggregate post-level data to weekly endpoint-level shares and rank them each week  
2. Compute forward log-growth distributions across 1, 7, and 28 weeks
3. Run a simulation that resamples change in log(share) by rank slot
4. Drops mass below a cutoff and replenish from an empirical entrant pool 
5. Compare simulated vs empirical capital distributions for rank-stationarity assessment

---

## Setup

```{r}
library(tidyverse)
library(arrow)
library(scales)
library(here)
library(lubridate)

set.seed(1823)

raw <- read_parquet(
  here("/Users/philipwaggoner/Dropbox (Personal)/GW - IDDP/Projects/Rank Diffusion (Matt)/Facebook/fb_top2000_ranked_daily.parquet")
)
```

## Step 1: Construct endpoint-level weekly series

```{r}
endpoint_weekly <- raw |>
  mutate(week = floor_date(date, 
                           unit = "week", 
                           week_start = 1)) |>
  group_by(week, endpoint_id) |>
  summarise(metric_value = sum(metric_value, na.rm = TRUE), 
            .groups = "drop") |>
  group_by(week) |>
  arrange(desc(metric_value), .by_group = TRUE) |>
  mutate(
    rank  = row_number(),
    share = metric_value / sum(metric_value, na.rm = TRUE)
  ) |>
  ungroup()

head(endpoint_weekly)
```

In the above chunk we are collapsing the raw post-level engagement into endpoint-level weekly totals. Then, we rank endpoints within each week (rank = row_number()), with rank 1 being the most engaged endpoint that week. Then, compute weekly share, which equals endpoint's `metric_value / total metric_value` across all endpoints for that week.

This gives a consistent panel with one row per endpoint per week, with both rank and relative share.

## Step 2: Compute forward-looking log growth

```{r}
endpoint_panel <- endpoint_weekly |>
  group_by(endpoint_id) |>
  arrange(week, .by_group = TRUE) |>
  mutate(
    log_growth_1w  = log(lead(share, 1)  / share),
    log_growth_7w  = log(lead(share, 7)  / share),
    log_growth_28w = log(lead(share, 28) / share)
  ) |>
  ungroup()

head(endpoint_panel)
```

Now, within each endpoint's time series, compute forward growth rates at horizons of 1, 7, and 28 weeks *ahead*, and show growth in terms of log differences. These values are supposed to capture volatility and drift in relative shares.

## Step 3: Categorize Endpoints by Size

```{r}
endpoint_panel <- endpoint_panel |>
  mutate(
    rank_group = case_when(
      rank <= 25  ~ "large",
      rank <= 250 ~ "midsize",
      TRUE        ~ "small"
    )
  )
```

Now, endpoints are bucketed by rank:

  - Top 25 = `large`
  - 26–250 = `midsize`
  - 251+ = `small`

This categorization is a bit arbitrary at the moment, but is meant to mirror the categorization idea from the book, which is helpful for analyzing differences in volatility by some scale.

Now, take a look at the growth distributions: 

```{r}
#| fig-width: 8
#| fig-height: 6

plot_df <- endpoint_panel |>
  pivot_longer(
    cols = starts_with("log_growth_"),
    names_to = "horizon",
    values_to = "log_growth"
  ) |>
  mutate(
    horizon = recode(horizon,
                     log_growth_1w = "1 Week",
                     log_growth_7w = "7 Week",
                     log_growth_28w = "28 Week"),
    horizon = factor(horizon, levels = c("1 Week", "7 Week", "28 Week"))
  ) |>
  filter(is.finite(log_growth))

ggplot(plot_df, aes(x = log_growth, fill = rank_group, col = rank_group)) +
  geom_density(alpha = 0.3) +
  facet_wrap(~ horizon, scales = "fixed") +
  labs(
    title = "Empirical: distribution of log growth by rank group",
    x = "Log growth rate", y = "Density",
    fill = "Rank group", color = "Rank group"
  ) +
  theme_bw() +
  theme(legend.position = "bottom")
```

From here, we get:

  - At 1-week horizon: distributions are tight, centered near zero
  - At longer horizons: distributions widen reflecting potentially higher volatility

\newpage

## Simulation

Now, we simulate weekly evolution of the rank-share curve by resampling delta log(share) from the empirical pool for each rank slot, renormalizing shares each step.

### Seed curve (mean share by rank) and increment pools

```{r}
# mean share by rank -> baseline w0 (see below)
mean_share <- endpoint_weekly |>
  group_by(rank) |>
  summarise(w_bar = mean(share, na.rm = TRUE), 
            .groups = "drop") |>
  arrange(rank)

# per-rank change in log(share) pool
rank_panel <- endpoint_weekly |>
  group_by(rank) |>
  arrange(week, .by_group = TRUE) |>
  mutate(
    share = pmax(share, 1e-12),
    log_w = log(share),
    dlogw = log_w - lag(log_w)
  ) |>
  ungroup() |>
  filter(is.finite(dlogw))
```

Now, for each rank position, compute weekly change in log share (`dlogw`), which forms the empirical pool of increments we will sample from.

### Choose K, build w_0, increment pools, and a sampler

```{r}
K <- min(500L, nrow(mean_share))
ranks <- 1:K

w0 <- mean_share$w_bar[ranks]
w0 <- w0 / sum(w0)

increments_by_rank <- rank_panel |>
  filter(rank %in% ranks) |>
  group_by(rank) |>
  summarise(increments = list(dlogw), .groups = "drop") |>
  arrange(rank)

sample_increments <- function() {
  vapply(increments_by_rank$increments, function(x) {
    if (length(x) == 0) return(0)
    sample(x, size = 1, replace = TRUE)
  }, numeric(1))
}
```

- `w_0` anchors all paths at a realistic starting curve

- `increments_by_rank` preserves rank-specific volatility structure

- `sample_increments()` is the core nonparametric resampling step

### Baseline simulator (no entry/exit)

```{r}
simulate_rank_paths <- function(w0, K, T = 52, n_paths = 200) {
  paths_final <- matrix(NA_real_, nrow = n_paths, ncol = K)
  horizons <- c(1, 7, 28)
  growth_collect <- vector("list", length(horizons))
  names(growth_collect) <- paste0("h", horizons)

  for (p in seq_len(n_paths)) {
    w <- w0
    for (t in seq_len(T)) {
      dlogw <- sample_increments()
      w <- w * exp(dlogw)
      w <- w / sum(w)
      if (t %in% horizons) {
        log_growth <- log(w / w0)
        growth_collect[[paste0("h", t)]] <- rbind(
          growth_collect[[paste0("h", t)]],
          cbind(path = p, 
                rank = ranks, 
                horizon = t,
                log_growth = log_growth)
        )
      }
    }
    paths_final[p, ] <- w
  }

  list(final = paths_final, 
       growth = bind_rows(growth_collect))
}

set.seed(42)
sim <- simulate_rank_paths(w0 = w0, K = K, T = 52, n_paths = 300)
```

This function simulates `n_paths` independent evolutions of the rank-share distribution. So, for each week (`block`), sample increments (`dlogw`) from historical pools. Then, update shares (`w * exp(dlogw)`), then renormalize to ensure shares sum to 1. Collect growth outcomes at horizons of 1, 7, and 28 weeks in the future. From this we get shares and a dataset of growth results.

### Baseline simulated growth & plot

```{r}
extract_horizon <- function(growth_df, col_name, label) {
  mat <- growth_df[[col_name]]
  df  <- as_tibble(mat)
  req <- c("rank", "log_growth")
  if (!all(req %in% names(df))) stop("Missing columns in ", col_name)
  path_vec <- if ("path" %in% names(df)) as.integer(df$path) else rep(NA_integer_, nrow(df))

  tibble(
    path       = path_vec,
    rank_slot  = as.integer(df$rank),
    horizon    = factor(label, 
                        levels = c("1 Week","7 Week","28 Week")),
    log_growth = as.numeric(df$log_growth)
  )
}

sim_growth <- bind_rows(
  extract_horizon(sim$growth, "h1",  "1 Week"),
  extract_horizon(sim$growth, "h7",  "7 Week"),
  extract_horizon(sim$growth, "h28", "28 Week")
) |>
  mutate(
    rank_group = case_when(
      rank_slot <= 25  ~ "large",
      rank_slot <= 250 ~ "midsize",
      TRUE             ~ "small"
    ),
    rank_group = factor(rank_group, 
                        levels = c("large","midsize","small"))
  )

ggplot(sim_growth, aes(x = log_growth, 
                       color = rank_group, fill = rank_group)) +
  geom_density(alpha = 0.25) +
  facet_wrap(~ horizon, scales = "fixed") +
  labs(
    title = "SIM (baseline): log growth by rank group & horizon",
    x = "Log growth", y = "Density",
    color = "Rank group", fill = "Rank group"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

\clearpage

## Entry/Exit extension (top-K cutoff)

Now, here's a build out on top of the baseline simulation. 

### Build an empirical entrant pool

```{r}
K_cut   <- 1000L # top-K we track
buffer  <- 300L # look below cutoff for near-entrants

entrant_pool <- endpoint_weekly |>
  filter(rank > K_cut, rank <= K_cut + buffer) |>
  transmute(share) |>
  filter(is.finite(share), share > 0)

stopifnot(nrow(entrant_pool) > 100)
entrant_sampler <- function(n) sample(entrant_pool$share, size = n, replace = TRUE)
```

### Simulator with entry/exit

```{r}
simulate_rank_paths_entry <- function(
  w0, K_cut = 1000L, K_max = length(w0), T = 52, n_paths = 200,
  entry_frac = 0.10
) {
  stopifnot(K_max >= K_cut)
  horizons <- c(1, 7, 28)
  growth_collect <- vector("list", length(horizons))
  names(growth_collect) <- paste0("h", horizons)

  top_n_sorted <- function(w, n) {
    o <- order(w, decreasing = TRUE)
    w[o][seq_len(n)]
  }

  paths_final <- matrix(NA_real_, nrow = n_paths, ncol = K_cut)

  for (p in seq_len(n_paths)) {
    w <- w0
    if (length(w) < K_max) {
      w <- c(w, entrant_sampler(K_max - length(w)))
      w <- w / sum(w)
      w <- top_n_sorted(w, K_max)
    }
    w_base <- w

    for (t in seq_len(T)) {
      dlogw <- sample_increments()
      if (length(dlogw) != length(w)) dlogw <- rep_len(dlogw, length.out = length(w))

      w <- w * exp(dlogw)
      w <- pmax(w, 1e-18)
      w <- w / sum(w)

      # exit: keep top-K_cut
      w_top <- top_n_sorted(w, K_cut)

      # entry: refill tail to K_max
      n_fill <- K_max - K_cut
      if (n_fill > 0) {
        n_add <- max(1L, floor(n_fill * entry_frac))
        w <- c(w_top, entrant_sampler(n_add), entrant_sampler(n_fill - n_add))
      } else {
        w <- w_top
      }

      w <- w / sum(w)
      w <- top_n_sorted(w, K_max)

      if (t %in% horizons) {
        w_now   <- top_n_sorted(w, K_cut)
        w_baseK <- top_n_sorted(w_base, K_cut)
        log_growth <- log(w_now / w_baseK)
        growth_collect[[paste0("h", t)]] <- rbind(
          growth_collect[[paste0("h", t)]],
          cbind(path = p, rank = seq_len(K_cut), horizon = t, log_growth = log_growth)
        )
      }
    }
    paths_final[p, ] <- top_n_sorted(w, K_cut)
  }

  list(final = paths_final, growth = bind_rows(growth_collect))
}
```


### Run entry/exit sim & plot

```{r}
#| fig-width: 8
#| fig-height: 6

K_cut <- 1000L
K_max <- 1200L

# extend w0 to K_max if needed
if (length(w0) < K_max) {
  w0_ext <- c(w0, entrant_sampler(K_max - length(w0)))
  w0_ext <- w0_ext / sum(w0_ext)
} else {
  w0_ext <- w0[1:K_max]; w0_ext <- w0_ext / sum(w0_ext)
}

set.seed(777)
sim2 <- simulate_rank_paths_entry(
  w0 = w0_ext, K_cut = K_cut, K_max = K_max,
  T = 52, n_paths = 300, entry_frac = 0.10
)

# ---- extract horizon helper ----
extract_horizon <- function(growth_df, col_name, label) {
  mat <- growth_df[[col_name]]
  df  <- as_tibble(mat)
  names(df) <- c("path", "rank", "horizon", "log_growth")  # enforce correct col names
  
  tibble(
    path       = as.integer(df$path),
    rank_slot  = as.integer(df$rank),
    horizon    = factor(label, levels = c("1 Week","7 Week","28 Week")),
    log_growth = as.numeric(df$log_growth)
  )
}

# ---- rebuild tidy sim growth ----
sim2_growth <- bind_rows(
  extract_horizon(sim2$growth, "h1",  "1 Week"),
  extract_horizon(sim2$growth, "h7",  "7 Week"),
  extract_horizon(sim2$growth, "h28", "28 Week")
) |>
  mutate(
    rank_group = case_when(
      rank_slot <= 25  ~ "large",
      rank_slot <= 250 ~ "midsize",
      TRUE             ~ "small"
    ),
    rank_group = factor(rank_group, levels = c("large","midsize","small"))
  )


# results
ggplot(sim2_growth, aes(x = log_growth, color = rank_group, fill = rank_group)) +
  geom_density(alpha = 0.25) +
  facet_wrap(~ horizon, scales = "free_y") +
  labs(
    title = "SIM v3 (entry/exit): log growth by rank group & horizon",
    x = "Log growth", y = "Density",
    color = "Rank group", fill = "Rank group"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")


# same bandwidth across facets + zoom around 0
ggplot(sim2_growth, aes(log_growth, color = rank_group, fill = rank_group)) +
  geom_density(alpha = 0.25, adjust = 1) + 
  facet_wrap(~ horizon, scales = "free_y") +
  geom_vline(xintercept = 0, linetype = 3) +
  coord_cartesian(xlim = c(-0.6, 0.6)) +
  labs(title = "SIM v3 (entry/exit): log-growth by rank group & horizon",
       subtitle = "zoom in around 0",
       x = "Log growth", y = "Density",
       color = "Rank group", fill = "Rank group") +
  theme_minimal() + theme(legend.position = "bottom")
```

We now see a bimodal distribution among *small* size ranks (i.e., rank >= 250). 

Importantly, this is largely a function of how the simulation is now setup:

  - we are sampling delta_log(share) (i.e., the "increment") separately for each rank slot; then renormalize after each time step to make sure all shares sum to 1
  - this allows entry and exit of endpoints around the rank cutoff (e.g., top 1000)
  - as a result, when new pages enter the ranking, they seem to do so with a burst of initial share (a short-lived viral spike), followed by a quick decay toward zero (once the spike fades)
  - so, when we aggregate all "small" ranks (again those > 250), that group contains a mix of steady low-share incumbents, and *also* a group of higher volatility new entrants who shoot up briefly, then crash

So, we can think of this mixture as producing two distributions in the log-growth distribution:

  1. a narrow group around 0; small incumbents with stable low shares
  2. a second group comprised of entrants whose share collapses after a short-lived spike

### Rank-stationarity check (capital distribution)

```{r}
#| fig-width: 6
#| fig-height: 6

# EMPIRICAL: mean share by rank within top-K (weekly)
K <- 1000

emp_capdist <- endpoint_weekly |>
  group_by(week) |>
  arrange(desc(share), .by_group = TRUE) |>
  mutate(rank_in_week = row_number()) |>
  filter(rank_in_week <= K) |>
  mutate(share_K = share / sum(share)) |> 
  ungroup() |>
  group_by(rank_in_week) |>
  summarise(w_bar = mean(share_K, na.rm = TRUE), .groups = "drop") |>
  rename(rank = rank_in_week)

# sim$final: n_paths x K matrix
sim_capdist <- as_tibble(sim$final) |>
  mutate(path = row_number()) |>
  pivot_longer(-path, names_to = "col", values_to = "share") |>
  mutate(col = as.integer(gsub("\\D", "", col))) |>
  group_by(path) |>
  arrange(desc(share), .by_group = TRUE) |>
  mutate(rank = row_number()) |>
  filter(rank <= K) |>
  group_by(path) |>
  mutate(shareK = share / sum(share)) |>
  ungroup() |>
  group_by(rank) |>
  summarise(w_bar = mean(shareK, na.rm = TRUE), .groups = "drop")

# viz
bind_rows(
  emp_capdist |> mutate(source = "Empirical"),
  sim_capdist |> mutate(source = "Sim v3")
) |>
  ggplot(aes(rank, w_bar, linetype = source)) +
  geom_line() +
  scale_x_log10() +
  scale_y_log10(labels = scales::percent_format(accuracy = 0.01)) +
  labs(
    title = "Capital-distribution (mean share)",
    subtitle = "Empirical vs Simulated",
    x = "Rank (log scale)", y = "Mean share (log scale)",
    linetype = ""
  ) +
  theme_minimal()

```

