---
title: "Rank Diffusion and Simulation of FB Data (v5)"
format: pdf
execute:
  echo: true
  warning: false
  message: false
---

## Introduction

This script builds on the simple simulation from v4 by first investigating why we ended up last time with 5563 max endpoints per week, compared to Matt's snapshot that showed a max of 14k, and much bigger numbers than mine. Then, once explored, we re-run the v4 simulation (no grid search yet) to be able to compare CDC curves to see whether and how different the empirical and simulated curves are. This gives a good foundation for moving forward as we decide how to define the sample. In future versions, once decided, we will continue the analysis to, e.g., conduct a grid search over mean and variance, and also martingales, etc. per Matt email. 

```{r}
library(tidyverse)
library(arrow)
library(scales)
library(here)
library(lubridate)
library(glue)

set.seed(1823)

raw <- read_parquet(
  here::here("/Users/philipwaggoner/Dropbox (Personal)/GW - IDDP/Projects/Rank Diffusion (Matt)/Facebook/fb_top2000_ranked_daily.parquet")
)
```

## Pages per week

```{r}
#| fig-width: 8
#| fig-height: 6

# Weekly count of unique endpoints
weekly_counts <- raw %>%
  mutate(week = floor_date(date, "week", week_start = 1)) %>%
  distinct(week, endpoint_id) %>%
  count(week, name = "n_endpoints")

summary_weekly <- weekly_counts %>%
  summarise(
    start = min(week), end = max(week),
    weeks = n(),
    min  = min(n_endpoints), which_min  = week[which.min(n_endpoints)],
    max  = max(n_endpoints), which_max  = week[which.max(n_endpoints)],
    median = median(n_endpoints), mean = mean(n_endpoints)
  )

#print(summary_weekly)

ggplot(weekly_counts, aes(week, n_endpoints)) +
  geom_line() +
  geom_smooth(se = FALSE, method = "loess", span = 0.2) +
  scale_y_continuous(labels = comma) +
  labs(title = "Weekly unique endpoints",
       x = "Week", 
       y = "Count of unique endpoint_id") + 
  theme_bw()
```

Still only seeing max of 5563. Why? 

```{r}
# 0) date range and overall unique pages
raw %>%
  summarise(min_date = min(date), max_date = max(date),
            n_pages  = n_distinct(endpoint_id))
```

Ok, we do see 14k + rows and 176 unique weeks and the same dates as the snapshot from Matt

So now, recreate that snapshot from Matt. Note the use of the term "unique". So we will start there: 

```{r}
# 1) Weekly unique pages (Monday-based weeks)
weekly_counts <- raw %>%
  mutate(
    endpoint_id = str_squish(endpoint_id),
    week        = floor_date(date, unit = "week", week_start = 1) # Monday
  ) %>%
  distinct(week, endpoint_id) %>%     # unique pages within each week
  count(week, name = "n_endpoints") %>%
  arrange(week)

# 2) Summary stats + coverage
summary_weekly <- weekly_counts %>%
  summarise(
    start   = min(week),
    end     = max(week),
    weeks   = n(),                           # should be 176
    max_n   = max(n_endpoints),
    max_wk  = week[which.max(n_endpoints)],
    min_n   = min(n_endpoints),
    min_wk  = week[which.min(n_endpoints)],
    median  = median(n_endpoints),
    mean    = mean(n_endpoints)
  )

# 3) First/last 5 weeks
first5 <- weekly_counts %>% slice_head(n = 5)
last5  <- weekly_counts %>% slice_tail(n = 5)

# 4) Pretty-print to match the style you pasted
fmt <- function(x) formatC(x, big.mark = ",", format = "f", digits = 0)
fmt1 <- function(x) formatC(x, big.mark = ",", format = "f", digits = 1)

cat(glue(
"Weekly Unique Pages

  - Coverage: {summary_weekly$start} >>> {summary_weekly$end} ({summary_weekly$weeks} weeks)
  - Metric: unique endpoint_id per week (Monday-based snapshot)
  - Summary:
      - Max: {fmt(summary_weekly$max_n)} on {summary_weekly$max_wk}
      - Min: {fmt(summary_weekly$min_n)} on {summary_weekly$min_wk}
      - Median: {fmt(summary_weekly$median)}
      - Mean: {fmt1(summary_weekly$mean)}
  - First 5 weeks:
{paste0('      - ', first5$week, ' >>> ', fmt(first5$n_endpoints), collapse = '\n')}
  - Last 5 weeks:
{paste0('      - ', last5$week,  ' >>> ', fmt(last5$n_endpoints), collapse = '\n')}
"))
```

Ok, from *unique* endpoints per week, we are back to 5563 (not 14k). 

So, now, let's relax "unique" and try with all endpoints:

```{r}
weekly_counts_all <- raw %>%
  mutate(
    endpoint_id = str_squish(endpoint_id),
    week        = floor_date(date, unit = "week", week_start = 1)  # Monday
  ) %>%
  count(week, name = "n_rows") %>%     # <-- count all rows, not unique pages
  arrange(week)

summary_all <- weekly_counts_all %>%
  summarise(
    start   = min(week),
    end     = max(week),
    weeks   = n(),
    max_n   = max(n_rows),
    max_wk  = week[which.max(n_rows)],
    min_n   = min(n_rows),
    min_wk  = week[which.min(n_rows)],
    median  = median(n_rows),
    mean    = mean(n_rows)
  )

# first/last 5 weeks to match Matt
first5_all <- weekly_counts_all %>% slice_head(n = 5)
last5_all  <- weekly_counts_all %>% slice_tail(n = 5)

fmt  <- function(x) formatC(x, big.mark = ",", format = "f", digits = 0)
fmt1 <- function(x) formatC(x, big.mark = ",", format = "f", digits = 1)

cat(glue(
"Weekly Totals (no de-duplication)

  - Coverage: {summary_all$start} >>> {summary_all$end} ({summary_all$weeks} weeks)
  - Metric: total rows per week (Monday-based snapshot)
  - Summary:
      - Max: {fmt(summary_all$max_n)} on {summary_all$max_wk}
      - Min: {fmt(summary_all$min_n)} on {summary_all$min_wk}
      - Median: {fmt(summary_all$median)}
      - Mean: {fmt1(summary_all$mean)}
  - First 5 weeks:
{paste0('      - ', first5_all$week, ' >>> ', fmt(first5_all$n_rows), collapse = '\n')}
  - Last 5 weeks:
{paste0('      - ', last5_all$week,  ' >>> ', fmt(last5_all$n_rows), collapse = '\n')}
"))
```

Ok, so these values are quite tidy and feel odd. They are close to Matt's but don't match exactly. Look a bit more. Start with looking at daily row counts:

```{r}
daily_rows <- raw %>%
  mutate(date = as.Date(date)) %>%
  count(date, name = "n_rows_day") %>%
  arrange(date)

summary(daily_rows$n_rows_day)
```

This proves we are working with the "top 2000 daily" file, as days are capped at 2000.

Now, any head/tail unique patterns? 

```{r}
head(daily_rows, 10); tail(daily_rows, 10)
```

No. Capped at 2k.

So, what about weekly totals, now, in light of the above? Look next at sum of daily rows in each Monday-based week (i.e., *no* de-dup):

```{r}
weekly_totals <- daily_rows %>%
  mutate(week = floor_date(date, "week", week_start = 1)) %>%
  group_by(week) %>%
  summarise(n_rows = sum(n_rows_day), .groups = "drop") %>%
  arrange(week)

weekly_unique <- raw %>%
  mutate(week = floor_date(date, "week", week_start = 1),
         endpoint_id = str_squish(endpoint_id)) %>%
  distinct(week, endpoint_id) %>%
  count(week, name = "n_unique") %>%
  arrange(week)

# side-by-side (rows vs unique endpoints)
weekly_compare <- weekly_totals %>%
  left_join(weekly_unique, by = "week")

head(weekly_compare, 10)
```

```{r}
summary(weekly_compare$n_rows); summary(weekly_compare$n_unique)
```

Ok, so we have confirmed that there's a real difference in looking at unique pages per week versus all pages (non-unique) per week. 

**Question at hand: which approach should we take?**

The v4 simulation took the unique approach, which is why we maxed out at 5,563 (instead of 14k as in Matt's version). 

So, to round out the first main sim approach, the following updates the v4 sim, and re-runs on the full (non-unique) universe and then looks at the capital plot to see if things look different or not. Once we decide on what should be included in the sample (unique vs. non-unique), we can update, and diagnose downstream accordingly. 

## Simulation on full data

Here, we'll do: a simple, large-k simulation and then a visual CDC comparison of the simulated curve vs the empirical curve: 

```{r}
# endpoint-level weekly series from all rows/endpoints
endpoint_weekly <- raw %>%
  mutate(week = floor_date(date, unit = "week", week_start = 1)) %>%
  group_by(week) %>%
  arrange(desc(metric_value), .by_group = TRUE) %>%
  mutate(
    rank  = row_number(),
    share = metric_value / sum(metric_value, na.rm = TRUE)
  ) %>%
  ungroup()

# baselines for the sim
# first, mean share by rank to seed curve
mean_share <- endpoint_weekly |>
  group_by(rank) |>
  summarise(w_bar = mean(share, na.rm = TRUE), .groups = "drop") |>
  arrange(rank)

# next, rank-wise weekly delta log(share) pools
rank_panel <- endpoint_weekly |>
  group_by(rank) |>
  arrange(week, .by_group = TRUE) |>
  mutate(
    share = pmax(share, 1e-12),
    log_w = log(share),
    dlogw = log_w - lag(log_w)
  ) |>
  ungroup() |>
  filter(is.finite(dlogw))

# def max (which is 14k to match Matt's approach)
max_rank_seen <- max(endpoint_weekly$rank, na.rm = TRUE)

# def "universe" (note: tail gets refilled by entrants)
K_cut       <- min(12000L, max_rank_seen)
K_max       <- min(K_cut + 2000L, max_rank_seen)
buffer_tail <- 2000L # entrant pool depth below cutoff

# seed curve w0 (length up to K_max)
w0_raw <- mean_share$w_bar[seq_len(min(K_max, nrow(mean_share)))]
w0_raw <- w0_raw / sum(w0_raw)

# def entrant pool: ranks just below the cutoff
entrant_pool <- endpoint_weekly |>
  filter(rank > K_cut,
         rank <= min(K_cut + buffer_tail, max_rank_seen)) |>
  transmute(share) |>
  filter(is.finite(share), share > 0)

# bootstrapping helper
entrant_sampler <- function(n) {
  sample(entrant_pool$share, size = n, replace = TRUE)
}

# rank-wise increment lists and a sampler
increments_by_rank <- rank_panel |>
  filter(rank <= K_max) |>
  group_by(rank) |>
  summarise(increments = list(dlogw), .groups = "drop") |>
  arrange(rank)

sample_increments <- function(sigma_scale = 0.1) {
  vapply(increments_by_rank$increments, function(x) {
    if (length(x) == 0) return(0)
    sigma_scale * sample(x, size = 1, replace = TRUE)
  }, numeric(1))
}

# def the simulator (entry/exit, track top-K_cut, etc. for CDC)
simulate_rank_paths <- function(
  w0, K_cut, K_max, T = 52, n_paths = 200,
  sigma_scale = 1, entry_frac = 0.10
) {
  stopifnot(length(w0) <= K_max, K_max >= K_cut)

  top_n_sorted <- function(w, n) w[order(w, decreasing = TRUE)][seq_len(n)]

  snapshots <- vector("list", n_paths)

  for (p in seq_len(n_paths)) {
    if (length(w0) < K_max) {
      w <- c(w0, entrant_sampler(K_max - length(w0)))
      w <- w / sum(w)
    } else {
      w <- w0[1:K_max]; w <- w / sum(w)
    }

    snaps_p <- vector("list", T)

    for (t in seq_len(T)) {
      dlogw <- sample_increments(sigma_scale = sigma_scale)
      if (length(dlogw) != length(w)) dlogw <- rep_len(dlogw, length.out = length(w))

      w <- w * exp(dlogw)
      w <- pmax(w, 1e-18)
      w <- w / sum(w)

      # keep top-K_cut, refill tail from entrant pool to K_max
      w_top  <- top_n_sorted(w, K_cut)
      n_tail <- K_max - K_cut
      if (n_tail > 0) {
        n_add <- max(1L, floor(n_tail * entry_frac))
        w <- c(w_top, entrant_sampler(n_add), entrant_sampler(n_tail - n_add))
      } else {
        w <- w_top
      }
      w <- w / sum(w)

      # CDC snapshot at time t: top-K_cut only, renormalized within K_cut
      wK <- top_n_sorted(w, K_cut)
      wK <- wK / sum(wK)
      snaps_p[[t]] <- tibble(path = p, t = t, rank = seq_len(K_cut), share = wK)
    }

    snapshots[[p]] <- bind_rows(snaps_p)
  }

  list(snapshots = bind_rows(snapshots))
}

# run a single, simple sim and build CDCs for comparison
set.seed(123)

T_weeks <- 52
N_paths <- 150

if (length(w0_raw) < K_max) { # extend w0 to K_max once
  w0_ext <- c(w0_raw, entrant_sampler(K_max - length(w0_raw)))
  w0_ext <- w0_ext / sum(w0_ext)
} else {
  w0_ext <- w0_raw[1:K_max]; w0_ext <- w0_ext / sum(w0_ext)
}
```

```{r echo = TRUE, eval = FALSE}
sim_out <- simulate_rank_paths(
  w0 = w0_ext, K_cut = K_cut, K_max = K_max,
  T = T_weeks, n_paths = N_paths,
  sigma_scale = 1.0, entry_frac = 0.10
)
```

```{r echo = FALSE, eval=TRUE}
sim_out <- read_rds("sim_out.rds")
```

```{r}
#| fig-width: 8
#| fig-height: 6

# first: empirical CDC showing average share curve within top-K_cut
emp_cdc <- endpoint_weekly |>
  group_by(week) |>
  arrange(desc(share), .by_group = TRUE) |>
  mutate(rank_in_week = row_number()) |>
  filter(rank_in_week <= K_cut) |>
  mutate(shareK = share / sum(share)) |>
  ungroup() |>
  group_by(rank = rank_in_week) |>
  summarise(w_bar = mean(shareK, na.rm = TRUE), .groups = "drop")

# second: sim CDC showing the average across snapshots and paths per sim
sim_cdc <- sim_out$snapshots |>
  group_by(rank) |>
  summarise(w_bar = mean(share, na.rm = TRUE), .groups = "drop")

bind_rows(
  emp_cdc |> mutate(source = "Empirical"),
  sim_cdc |> mutate(source = "Simulated")
) |>
  ggplot(aes(rank, w_bar, linetype = source)) +
  geom_line() +
  scale_x_log10() +
  scale_y_log10(labels = percent_format(accuracy = 0.01)) +
  labs(
    title = paste0("CDC: Empirical vs Simulated  (K_cut=", K_cut, ", K_max=", K_max, ")"),
    x = "Rank (log scale)",
    y = "Mean share (log scale)",
    linetype = NULL
  ) +
  theme_bw()
```
