## Setup

This section initializes parameters and loads the weekly Facebook data into a core panel.

```{r setup-vars, include=FALSE}
K_cut_target <- CFG$K_cut_target
K_tail_buffer <- CFG$K_tail_buffer
Smoothing_h <- CFG$smoothing_h

horizons_growth <- CFG$horizons_growth
horizons_durable <- CFG$horizons_durable
horizons_micro <- CFG$horizons_micro
turnover_K_list <- CFG$turnover_K_list
turnover_horizons <- CFG$turnover_horizons

bucket_def <- CFG$bucket_def
bootstrap_B <- CFG$bootstrap_B
bootstrap_k_min <- CFG$bootstrap_k_min
bootstrap_period_unit <- CFG$bootstrap_period_unit

sim_T_weeks <- CFG$sim_T_weeks
sim_n_paths <- CFG$sim_n_paths
sim_mu <- CFG$sim_mu
sim_sigma <- CFG$sim_sigma
sim_entry_frac <- CFG$sim_entry_frac
sim_K_xi <- CFG$sim_K_xi
pca_K <- CFG$pca_K

if (!exists("endpoint_bundle")) {
  endpoint_bundle <- load_endpoint_weekly(cfg = CFG)
}
endpoint_weekly <- endpoint_bundle$endpoint_weekly
max_rank_by_week <- endpoint_bundle$max_rank_by_week
max_rank_seen <- endpoint_bundle$max_rank_seen
K_cut <- endpoint_bundle$K_cut
K_max <- endpoint_bundle$K_max
sim_K_xi <- endpoint_bundle$sim_K_xi
```

## Load weekly Facebook data and construct core panel

The exploration treats the weekly Facebook file as weekly snapshots with (at least):
`date`, `endpoint_id`, `metric_value`, `rank` (rank is recomputed defensively below).

We construct:

- `share_global`: each endpoint's share of total weekly attention (sums to 1 within week).
- `share_topK`: same share, renormalized within the top-K (`K_cut`) each week.

```{r load-data}
endpoint_weekly %>% glimpse()
```
