## Stylized fact B: Rank-dependent drift and volatility (rank-slot increments)

We work with rank-slot series: for each rank `k`, the sequence of shares of the endpoint
occupying that rank each week. This is the object that rank-based diffusion models target.

```{r model-setup, include=FALSE}
if (!exists("endpoint_bundle")) {
  endpoint_bundle <- load_endpoint_weekly(cfg = CFG)
}
endpoint_weekly <- endpoint_bundle$endpoint_weekly
K_cut <- endpoint_bundle$K_cut
K_max <- endpoint_bundle$K_max
max_rank_seen <- endpoint_bundle$max_rank_seen
sim_K_xi <- endpoint_bundle$sim_K_xi

bucket_def <- CFG$bucket_def
Smoothing_h <- CFG$smoothing_h
turnover_K_list <- CFG$turnover_K_list
turnover_horizons <- CFG$turnover_horizons
horizons_micro <- CFG$horizons_micro
bootstrap_period_unit <- CFG$bootstrap_period_unit
bootstrap_B <- CFG$bootstrap_B
bootstrap_k_min <- CFG$bootstrap_k_min
pca_K <- CFG$pca_K

if (!exists("rank_panel")) {
  rank_panel <- load_rank_panel(endpoint_weekly, cfg = CFG)
}
if (!exists("gauss_params_raw")) {
  gauss_params_raw <- load_gauss_params_raw(rank_panel, K_max, cfg = CFG)
}
if (!exists("sm_params")) {
  sm_params <- load_gauss_params_smoothed(gauss_params_raw, K_max, Smoothing_h, cfg = CFG)
}
```

```{r rank-panel}
rank_panel %>% glimpse()

ggplot(gauss_params_raw, aes(rank, sd_dlogw)) +
  geom_line() +
  scale_x_continuous(trans = "log10") +
  labs(
    title = "Rank-dependent volatility proxy",
    subtitle = "sd(d log w) by rank (rank-slot series, global shares)",
    x = "Rank (log scale)",
    y = "sd(d log w)"
  ) +
  theme_bw()
```

### Smoothed per-rank moments (v9+)

```{r smoothing-helpers}
sm_plot <- sm_params %>%
  select(rank, mean_dlogw_s, sd_dlogw_s) %>%
  left_join(gauss_params_raw %>% select(rank, mean_dlogw, sd_dlogw), by = "rank") %>%
  pivot_longer(
    cols = c(mean_dlogw, mean_dlogw_s, sd_dlogw, sd_dlogw_s),
    names_to = "series",
    values_to = "value"
  ) %>%
  mutate(kind = case_when(
    str_detect(series, "^mean") ~ "mean",
    TRUE ~ "sd"
  ))

ggplot(filter(sm_plot, kind == "mean"), aes(rank, value, color = series)) +
  geom_line() +
  scale_x_continuous(trans = "log10") +
  labs(
    title = "Per-rank drift proxy: raw vs smoothed",
    x = "Rank (log scale)",
    y = "mean(d log w)"
  ) +
  theme_bw() +
  guides(color = "none")

ggplot(filter(sm_plot, kind == "sd"), aes(rank, value, color = series)) +
  geom_line() +
  scale_x_continuous(trans = "log10") +
  labs(
    title = "Per-rank volatility proxy: raw vs smoothed",
    x = "Rank (log scale)",
    y = "sd(d log w)"
  ) +
  theme_bw() +
  guides(color = "none")
```

## Stylized fact D: Gap structure via Xi_k(t) = log(w_k / w_{k+1}) (v11)

We compute Xi using *global* shares, with ranks recomputed within each week.

```{r xi-weekly}
xi_weekly <- endpoint_weekly %>%
  group_by(week) %>%
  arrange(rank, .by_group = TRUE) %>%
  mutate(
    w = pmax(share_global, 1e-18),
    w_next = lead(w),
    xi = log(w / w_next),
    k = rank
  ) %>%
  ungroup() %>%
  filter(is.finite(xi), is.finite(k))

K_xi_est <- min(2000L, K_cut)

# scatter (restricted for clarity)
ggplot(filter(xi_weekly, k <= 500L), aes(x = k, y = xi)) +
  geom_point(alpha = 0.35, size = 0.4) +
  scale_x_continuous(trans = "log10") +
  labs(
    title = "Log gaps (Xi) by rank: Xi_k(t) = log(w_k / w_{k+1})",
    subtitle = "Restricted to top 500 ranks for visibility",
    x = "Rank k (log scale)",
    y = "Xi"
  ) +
  theme_bw()

rho_by_k_med <- xi_weekly %>%
  filter(k <= (K_xi_est - 1L)) %>%
  group_by(k) %>%
  summarise(rho_k = median(xi, na.rm = TRUE), .groups = "drop")

ggplot(rho_by_k_med, aes(k, rho_k)) +
  geom_line() +
  theme_bw() +
  labs(
    title = "Median Xi by rank (rho_k)",
    x = "Rank k",
    y = "rho_k"
  )
```

### v11-style sigma_hat(k) construction

Following v11:

- Let `g_k` be smoothed per-rank drift proxy (mean of `d log w`) for ranks.
- Define `S_k = cumsum(g_k)`.
- Combine with `rho_k` to form `sigma_hat(k) = sqrt( -2 S_k rho_k )` when positive.

```{r sigma-hat}
g_by_k <- sm_params %>%
  transmute(k = rank, g_k = mean_dlogw_s) %>%
  filter(k <= (K_xi_est + 1L)) %>%
  arrange(k)

S_by_k <- g_by_k %>%
  mutate(S_k = cumsum(g_k))

sigma_hat_by_k <- rho_by_k_med %>%
  inner_join(S_by_k %>% select(k, S_k), by = "k") %>%
  mutate(
    sigma2_hat = -2 * S_k * rho_k,
    sigma_hat = if_else(sigma2_hat > 0, sqrt(sigma2_hat), NA_real_)
  )

sigma_hat_by_k %>%
  summarise(med_sigma = median(sigma_hat, na.rm = TRUE))

ggplot(
  filter(sigma_hat_by_k, k >= 10L, !is.na(sigma_hat)),
  aes(x = k, y = sigma_hat)
) +
  geom_line() +
  scale_x_continuous(trans = "log10") +
  labs(
    title = "sigma_hat(k) vs rank (v11-style)",
    x = "Rank k (log scale)",
    y = "sigma_hat(k)"
  ) +
  theme_bw()
```

### Stability across time (quarterly summary)

```{r sigma-hat-over-time}
xi_by_period <- xi_weekly %>%
  mutate(period = floor_date(week, bootstrap_period_unit)) %>%
  filter(k <= (K_xi_est - 1L))

rho_period_by_k <- xi_by_period %>%
  group_by(period, k) %>%
  summarise(rho_k = median(xi, na.rm = TRUE), .groups = "drop")

sigma_hat_period <- rho_period_by_k %>%
  inner_join(S_by_k %>% select(k, S_k), by = "k") %>%
  mutate(
    sigma2_hat = -2 * S_k * rho_k,
    sigma_hat = if_else(sigma2_hat > 0, sqrt(sigma2_hat), NA_real_)
  )

sigma_hat_period_summ <- sigma_hat_period %>%
  group_by(period) %>%
  summarise(
    sigma_med = median(sigma_hat, na.rm = TRUE),
    sigma_p25 = quantile(sigma_hat, 0.25, na.rm = TRUE),
    sigma_p75 = quantile(sigma_hat, 0.75, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(sigma_hat_period_summ, aes(x = period, y = sigma_med)) +
  geom_line() +
  geom_ribbon(aes(ymin = sigma_p25, ymax = sigma_p75), alpha = 0.2) +
  labs(
    title = "sigma_hat over time (median across ranks) with IQR band",
    x = NULL,
    y = "sigma_hat (median)"
  ) +
  theme_bw()
```

### Bootstrap CI over time (v11-style)

```{r bootstrap-ci}
sigma_ci_by_period <- xi_by_period %>%
  group_by(period) %>%
  group_modify(~ bootstrap_sigma_period(
    xi_df_period = .x,
    S_by_k = S_by_k,
    B = as.integer(bootstrap_B),
    k_min = as.integer(bootstrap_k_min),
    K_xi_est = K_xi_est
  )) %>%
  ungroup()

ggplot(sigma_ci_by_period, aes(x = period, y = sigma_hat_med)) +
  geom_line() +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), alpha = 0.2) +
  labs(
    title = "Bootstrap CI for sigma_hat over time",
    x = NULL,
    y = "sigma_hat (median across k)"
  ) +
  theme_bw()
```

## Stylized fact C (expanded): Turnover and rank mobility (new)

### 9.1 Top-K membership turnover (set overlap)

```{r topk-turnover}
turnover_tbl <- purrr::map_dfr(
  turnover_K_list,
  ~ topK_overlap(endpoint_weekly, K = .x, horizons = turnover_horizons)
)

ggplot(turnover_tbl, aes(x = week, y = overlap_frac, color = factor(horizon))) +
  geom_line(alpha = 0.6) +
  facet_wrap(~ K, scales = "free_y") +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(
    title = "Top-K overlap over time",
    subtitle = "Fraction of endpoints in top-K at week t that also appear in top-K at t+h",
    x = "Week",
    y = "Overlap fraction",
    color = "Horizon (weeks)"
  ) +
  theme_bw()
```

### 9.2 Rank-conditioned mobility: where does the endpoint at rank k go?

```{r micro-emp-extended}
k_grid_mobility <- c(1L, 5L, 12L, 25L, 100L, 500L, 1000L)

emp_micro_mobility <- purrr::map_dfr(
  k_grid_mobility,
  ~ emp_micro_for_k(endpoint_weekly, k = .x, horizons = horizons_micro) %>%
    mutate(k = .x)
)

mobility_summ <- emp_micro_mobility %>%
  group_by(k, horizon) %>%
  summarise(
    retention = mean(present_tp),
    p_within_1 = mean(abs(rank_change) <= 1, na.rm = TRUE),
    p_within_5 = mean(abs(rank_change) <= 5, na.rm = TRUE),
    med_abs_rank_change = median(abs(rank_change), na.rm = TRUE),
    .groups = "drop"
  )

mobility_summ
```

```{r mobility-plots}
ggplot(mobility_summ, aes(x = horizon, y = retention, group = factor(k))) +
  geom_line() +
  geom_point(size = 1.5) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_x_continuous(breaks = horizons_micro) +
  labs(
    title = "Retention by horizon for rank-conditioned cohorts",
    x = "Horizon (weeks ahead)",
    y = "Fraction still observed",
    caption = "Cohort: endpoint that sits at rank k at week t"
  ) +
  theme_bw()

ggplot(mobility_summ, aes(x = horizon, y = med_abs_rank_change, group = factor(k))) +
  geom_line() +
  geom_point(size = 1.5) +
  scale_x_continuous(breaks = horizons_micro) +
  labs(
    title = "Median |rank change| for rank-conditioned cohorts",
    x = "Horizon (weeks ahead)",
    y = "Median absolute rank change"
  ) +
  theme_bw()
```

### 9.3 Rank transition matrix (deciles + out-of-topK)

```{r transition-matrix}
rank_bins <- endpoint_weekly %>%
  select(week, endpoint_id, rank) %>%
  mutate(bin = make_rank_bin(rank, K = 1000L, n_bins = 10L))

trans_1w <- rank_bins %>%
  transmute(
    endpoint_id,
    week_t = week,
    bin_t = bin,
    week_tp = week + lubridate::weeks(1L)
  ) %>%
  left_join(
    rank_bins %>% transmute(endpoint_id, week_tp = week, bin_tp = bin),
    by = c("endpoint_id", "week_tp")
  ) %>%
  mutate(bin_tp = replace_na(bin_tp, "out")) %>%
  count(bin_t, bin_tp) %>%
  group_by(bin_t) %>%
  mutate(p = n / sum(n)) %>%
  ungroup()

ggplot(trans_1w, aes(x = bin_tp, y = bin_t, fill = p)) +
  geom_tile() +
  scale_fill_continuous(labels = percent_format(accuracy = 0.1)) +
  labs(
    title = "Rank-bin transition matrix (1 week)",
    subtitle = "Bins are deciles within top 1000 ranks; 'out' means rank>1000 or missing",
    x = "Bin at t+1",
    y = "Bin at t"
  ) +
  theme_bw()
```

## Stylized fact E: Common shocks vs idiosyncratic movement (new)

A pragmatic diagnostic: PCA on rank-slot increments for the top `pca_K` ranks.

```{r common-shocks-pca}
K_pca <- min(pca_K, K_cut)

if (is.na(K_pca) || K_pca < 2) {
  message("Skipping PCA: K_pca < 2 (K_pca=", K_pca, ").")
} else {
  rank_slot_wide <- rank_panel %>%
    filter(rank <= K_pca) %>%
    select(week, rank, dlogw) %>%
    pivot_wider(names_from = rank, values_from = dlogw)

  rank_slot_mat <- rank_slot_wide %>%
    drop_na() %>%
    arrange(week)

  X <- rank_slot_mat %>%
    select(-week) %>%
    as.matrix()

  if (nrow(X) < 2 || ncol(X) < 2) {
    message(
      "Skipping PCA: need at least 2 complete weeks and 2 rank columns; got ",
      nrow(X), "x", ncol(X), "."
    )
  } else {
    pca_fit <- prcomp(X, center = TRUE, scale. = TRUE)

    var_explained <- (pca_fit$sdev^2) / sum(pca_fit$sdev^2)

    tibble(
      component = seq_along(var_explained),
      var_explained = var_explained
    ) %>%
      slice(1:10)

    ggplot(
      tibble(component = seq_along(var_explained), var_explained = var_explained) %>% slice(1:20),
      aes(component, var_explained)
    ) +
      geom_line() +
      geom_point(size = 1.5) +
      scale_y_continuous(labels = percent_format(accuracy = 0.1)) +
      labs(
        title = "PCA on rank-slot increments: variance explained",
        subtitle = paste0("Top ", K_pca, " ranks; complete weeks only"),
        x = "Principal component",
        y = "Share of variance explained"
      ) +
      theme_bw()

    pc1 <- pca_fit$x[, 1]
    pc1_df <- tibble(week = rank_slot_mat$week, pc1 = pc1)

    ggplot(pc1_df, aes(week, pc1)) +
      geom_line() +
      labs(
        title = "First principal component over time (common shock proxy)",
        x = "Week",
        y = "PC1 score"
      ) +
      theme_bw()
  }
}
```

## Heavy-tail / jump diagnostics for increments (new)

This section checks whether rank-slot increments have tails that are substantially heavier than Gaussian.

```{r tail-diagnostics}
inc_by_bucket <- rank_panel %>%
  filter(rank <= K_cut) %>%
  mutate(bucket = assign_bucket(rank, bucket_def)) %>%
  filter(!is.na(bucket)) %>%
  group_by(bucket) %>%
  mutate(
    z = (dlogw - mean(dlogw, na.rm = TRUE)) / sd(dlogw, na.rm = TRUE)
  ) %>%
  ungroup()

tail_summ <- inc_by_bucket %>%
  group_by(bucket) %>%
  summarise(
    n = n(),
    p_gt_5sd = mean(abs(z) >= 5, na.rm = TRUE),
    p_gt_3sd = mean(abs(z) >= 3, na.rm = TRUE),
    q_0.001 = quantile(z, 0.001, na.rm = TRUE),
    q_0.999 = quantile(z, 0.999, na.rm = TRUE),
    .groups = "drop"
  )

tail_summ
```

```{r qq-plot}
ggplot(inc_by_bucket, aes(sample = z)) +
  stat_qq() +
  stat_qq_line() +
  facet_wrap(~ bucket, scales = "free") +
  labs(
    title = "QQ plot of standardized increments by bucket",
    subtitle = "Visual check for heavy tails vs Gaussian"
  ) +
  theme_bw()
```

## Simulation framework (v8–v11) + unified fit metrics

### 12.1 Empirical entrant pool for entry/exit

```{r entrant-pool}
entrant_pool <- build_entrant_pool(endpoint_weekly, K_cut, CFG$K_tail_buffer, max_rank_seen)

stopifnot(nrow(entrant_pool) > 100)

entrant_sampler <- make_entrant_sampler(entrant_pool)

w0_ext <- build_w0_ext(endpoint_weekly, K_max, entrant_sampler)
```

### 12.2 Simulator (Gaussian increments with smoothed moments; optional Xi output)

This merges v9’s simulator with v11’s Xi storage.

### 12.3 Fit metrics (CDC, durable change, Xi)

```{r fit-metrics, include=FALSE}
if (!exists("emp_cdc")) {
  emp_cdc <- load_emp_cdc(endpoint_weekly, K_cut, cfg = CFG)
}
if (!exists("emp_targets")) {
  emp_targets <- load_emp_targets(endpoint_weekly, K_cut, CFG$horizons_durable, bucket_def, cfg = CFG)
}
```

## Baseline simulation run + comparison (CDC, durable, Xi)

```{r run-sim}
mean_vec_s <- sm_params$mean_dlogw_s
sd_vec_s <- sm_params$sd_dlogw_s

if (!exists("sim_out")) {
  sim_out <- load_sim_baseline(
    w0_ext = w0_ext,
    mean_vec_s = mean_vec_s,
    sd_vec_s = sd_vec_s,
    K_cut = K_cut,
    K_max = K_max,
    sim_K_xi = sim_K_xi,
    cfg = CFG,
    entrant_sampler = entrant_sampler
  )
}

# CDC comparison
sim_cdc <- sim_out$snapshots %>%
  group_by(rank) %>%
  summarise(w_bar = mean(share, na.rm = TRUE), .groups = "drop") %>%
  arrange(rank)

bind_rows(
  emp_cdc %>% mutate(source = "Empirical"),
  sim_cdc %>% mutate(source = "Simulated")
) %>%
  ggplot(aes(rank, w_bar, linetype = source)) +
  geom_line() +
  scale_x_log10() +
  scale_y_log10(labels = percent_format(accuracy = 0.01)) +
  labs(
    title = "CDC: Empirical vs Simulated",
    subtitle = paste0("mu=", CFG$sim_mu, ", sigma=", CFG$sim_sigma,
                      ", h=", Smoothing_h),
    x = "Rank (log scale)",
    y = "Mean share within top-K (log scale)",
    linetype = NULL
  ) +
  theme_bw()

# Scoreboard
score_tbl <- tibble(
  rmse_cdc = cdc_rmse(emp_cdc, sim_cdc),
  rmse_durable = durable_rmse(emp_targets, sim_out$growth, bucket_def),
  rmse_xi = xi_rmse(rho_by_k_med %>% filter(k <= sim_K_xi - 1L),
                    sim_out$xi %>% filter(k <= sim_K_xi - 1L))
)

score_tbl
```
