## Candidate model families (model zoo)

We test heavy-tailed innovations (Student-t, skew-t, GH), explicit jump mixtures (Merton, Kou), and factor + idiosyncratic structures with asymmetric jumps to capture directional tail behavior and common shocks [@merton1976option; @kou2002jump; @eberlein1995hyperbolic; @azzalini2003skewt]. Optional clustering extensions (self-exciting intensity or regime switching) are gated by diagnostics [@hawkes1971spectra; @hamilton1989new]. Skew-t and GH fits run only when the `sn`/`ghyp` packages are available.

## Estimation protocol

We estimate one-step standardized rank-slot increments first (bucketed by default, with optional smooth-by-rank for jump intensities), then evaluate simulation fit to CDC/durable/Xi and tail/skew targets. Jump-mixture parameters are estimated by EM with explicit constraints and convergence checks; out-of-sample validation uses a time split.

Model selection follows three layers: (i) distributional fit (AIC/BIC + tail/skew diagnostics), (ii) out-of-sample log-likelihood and target fit, and (iii) simulation-based predictive envelopes for core targets.

Key constraints enforced in estimation: jump probabilities bounded by `p_max`, Student-t `df > 2`, and mixture ordering (low-variance vs high-variance components).

## Model zoo comparison

```{r model-zoo-compare}
model_comp <- NULL
if (isTRUE(CFG$run_jump_model_zoo)) {
  model_comp <- run_model_comparison(
    cfg = CFG,
    endpoint_weekly = endpoint_weekly,
    rank_panel = rank_panel,
    sm_params = sm_params,
    w0_ext = w0_ext,
    entrant_sampler = entrant_sampler,
    bucket_def = bucket_def,
    K_cut = K_cut,
    K_max = K_max,
    K_xi = sim_K_xi,
    pca_K = pca_K
  )
  model_comp$model_table
}
```

## Model zoo visual comparison (side-by-side)

```{r model-zoo-qq-empirical, echo=FALSE, out.width="100%"}
if (!is.null(model_comp)) {
  plot_dir <- here::here("output", "plots")
  dir.create(plot_dir, showWarnings = FALSE, recursive = TRUE)

  model_meta <- model_comp$model_table %>%
    dplyr::mutate(label = ifelse(is.na(name) | name == "", model, name))
  model_ids <- model_meta$model
  model_labels <- model_meta$label

  moment_curves_zoo <- list(
    mean_vec = sm_params$mean_dlogw_s,
    sd_vec = sm_params$sd_dlogw_s,
    bucket_def = bucket_def
  )

  sim_samples <- purrr::map_dfr(seq_along(model_ids), function(i) {
    mid <- model_ids[i]
    label <- model_labels[i]
    res <- model_comp$results[[mid]]
    fit <- res$fit
    if (is.null(fit)) return(NULL)
    model_sim <- prepare_model_for_sim(fit, CFG, moment_curves_zoo)
    sim_inc <- tryCatch({
      sample_increments_for_tail(
        model = model_sim,
        moment_curves = moment_curves_zoo,
        cache = list(beta_k = fit$beta_k %||% pca_cache$beta_k, state = new.env(parent = emptyenv())),
        K_cut = K_cut,
        n_steps = 80L,
        n_paths = 15L,
        ranks_per_bucket = 120L
      )
    }, error = function(e) NULL)
    if (is.null(sim_inc)) return(NULL)
    standardize_increments(sim_inc, sm_params$mean_dlogw_s, sm_params$sd_dlogw_s, bucket_def = bucket_def) %>%
      dplyr::mutate(z = dlogw_std, model = label)
  })

  if (nrow(sim_samples) > 0) {
    qq_probs <- seq(0.01, 0.99, by = 0.01)
    emp_q <- rank_inc_std %>%
      dplyr::group_by(bucket) %>%
      dplyr::summarise(q = list(stats::quantile(z, probs = qq_probs, na.rm = TRUE)), .groups = "drop") %>%
      tidyr::unnest_longer(q, values_to = "emp_q") %>%
      dplyr::group_by(bucket) %>%
      dplyr::mutate(p = qq_probs) %>%
      dplyr::ungroup() %>%
      dplyr::select(bucket, p, emp_q)

    model_q <- sim_samples %>%
      dplyr::group_by(model, bucket) %>%
      dplyr::summarise(q = list(stats::quantile(z, probs = qq_probs, na.rm = TRUE)), .groups = "drop") %>%
      tidyr::unnest_longer(q, values_to = "model_q") %>%
      dplyr::group_by(model, bucket) %>%
      dplyr::mutate(p = qq_probs) %>%
      dplyr::ungroup() %>%
      dplyr::select(model, bucket, p, model_q)

    qq_df <- model_q %>%
      dplyr::left_join(emp_q, by = c("bucket", "p")) %>%
      dplyr::mutate(model = factor(model, levels = model_labels))

    qq_plot <- ggplot(qq_df, aes(x = emp_q, y = model_q)) +
      geom_point(size = 0.6, alpha = 0.6) +
      geom_abline(linetype = "dashed", color = "gray40") +
      facet_grid(model ~ bucket, scales = "free") +
      labs(
        title = "Model vs empirical QQ (standardized increments)",
        subtitle = "Each panel compares model-simulated quantiles to empirical quantiles",
        x = "Empirical quantiles",
        y = "Model quantiles"
      ) +
      theme_bw()

    height_in <- max(4, 1.1 * length(model_labels) + 1)
    qq_path <- file.path(plot_dir, "qq_model_vs_empirical.png")
    ggsave(qq_path, qq_plot, dpi = 300, width = 10, height = height_in, units = "in")
    knitr::include_graphics(qq_path)
  }
}
```

```{r model-zoo-tail-metrics-plot, echo=FALSE, out.width="100%"}
if (!is.null(model_comp)) {
  plot_dir <- here::here("output", "plots")
  dir.create(plot_dir, showWarnings = FALSE, recursive = TRUE)

  model_meta <- model_comp$model_table %>%
    dplyr::mutate(label = ifelse(is.na(name) | name == "", model, name))
  model_ids <- model_meta$model
  model_labels <- model_meta$label

  emp_long <- tail_rank %>%
    tidyr::pivot_longer(-bucket, names_to = "metric", values_to = "emp")

  sim_long <- purrr::map_dfr(seq_along(model_ids), function(i) {
    mid <- model_ids[i]
    label <- model_labels[i]
    res <- model_comp$results[[mid]]
    tail_sim <- res$tail_sim
    if (is.null(tail_sim)) return(NULL)
    tail_sim %>%
      tidyr::pivot_longer(-bucket, names_to = "metric", values_to = "sim") %>%
      dplyr::mutate(model = label)
  })

  if (nrow(sim_long) > 0) {
    tail_plot_df <- sim_long %>%
      dplyr::left_join(emp_long, by = c("bucket", "metric")) %>%
      dplyr::mutate(model = factor(model, levels = model_labels))

    tail_plot <- ggplot(tail_plot_df, aes(x = emp, y = sim, color = model)) +
      geom_abline(linetype = "dashed", color = "gray40") +
      geom_point(size = 1.1, alpha = 0.7) +
      facet_grid(metric ~ bucket, scales = "free") +
      labs(
        title = "Tail and skew metrics: simulated vs empirical",
        x = "Empirical",
        y = "Simulated"
      ) +
      theme_bw() +
      guides(color = guide_legend(ncol = 2))

    metric_count <- length(unique(emp_long$metric))
    height_in <- max(4, 0.7 * metric_count + 2)
    tail_path <- file.path(plot_dir, "tail_metrics_model_compare.png")
    ggsave(tail_path, tail_plot, dpi = 300, width = 10, height = height_in, units = "in")
    knitr::include_graphics(tail_path)
  }
}
```

```{r model-zoo-params}
if (!is.null(model_comp)) {
  param_tables <- purrr::imap(model_comp$results, function(res, name) {
    fit <- res$fit
    if (is.null(fit)) return(NULL)
    if (fit$type == "factor_kou_smooth_twostage") {
      return(list(
        factor = fit$factor_fit,
        idio = list(theta = fit$idio_fit$theta, param_curves = fit$param_curves),
        diagnostics = list(stage1 = fit$stage1, stage2 = fit$stage2, stage3 = fit$stage3)
      ))
    }
    if (fit$type %in% c("factor_kou", "factor_kou_tv", "factor_t", "factor_regime")) {
      return(list(factor = fit$factor_fit %||% fit$factor_params, idio = fit$idio_fit$params))
    }
    fit$params
  })
  param_tables
}
```

## Preferred model and interpretation

The default lead specification is the factor + asymmetric Kou jump model (C2), unless the comparison table clearly favors a simpler alternative on both distributional and simulation targets.

```{r model-zoo-best}
if (!is.null(model_comp)) {
  best_id <- model_comp$model_table$model[1]
  best_fit <- model_comp$results[[best_id]]$fit
  list(best_model = best_id, type = best_fit$type)
}
```

```{r model-zoo-tail-compare}
if (!is.null(model_comp)) {
  best_id <- model_comp$model_table$model[1]
  tail_best <- model_comp$results[[best_id]]$tail_sim
  tail_rank %>%
    pivot_longer(-bucket, names_to = "metric", values_to = "emp") %>%
    left_join(
      tail_best %>% pivot_longer(-bucket, names_to = "metric", values_to = "sim"),
      by = c("bucket", "metric")
    ) %>%
    arrange(bucket, metric)
}
```

## Robustness: block bootstrap + sensitivity

```{r model-zoo-bootstrap}
if (!is.null(model_comp) && CFG$model_zoo_bootstrap_B > 0) {
  best_id <- model_comp$model_table$model[1]
  fit_fn <- model_comp$results[[best_id]]$fit_fn
  boot_params <- bootstrap_model_params(
    fit_fn,
    model_comp$z_rank,
    B = CFG$model_zoo_bootstrap_B,
    block_length = CFG$model_zoo_bootstrap_block_length
  )
  boot_params
}
```

Sensitivity checks for alternative bucket cutoffs and smoothing windows can be run by re-rendering the report with different `CFG$bucket_breaks` and `CFG$smoothing_h` values (not shown by default).

Bootstrap resamples are block-based with length `CFG$model_zoo_bootstrap_block_length` to preserve temporal dependence.
