## Sigma calibration loop (v10) (optional but included)

This reproduces the v10 approach: scan `sigma` and pick the value that minimizes durable mismatch, with a small CDC penalty.

```{r sigma-setup, include=FALSE}
if (!exists("endpoint_bundle")) {
  endpoint_bundle <- load_endpoint_weekly(cfg = CFG)
}
endpoint_weekly <- endpoint_bundle$endpoint_weekly
K_cut <- endpoint_bundle$K_cut
K_max <- endpoint_bundle$K_max
max_rank_seen <- endpoint_bundle$max_rank_seen

bucket_def <- CFG$bucket_def
Smoothing_h <- CFG$smoothing_h

if (!exists("rank_panel")) {
  rank_panel <- load_rank_panel(endpoint_weekly, cfg = CFG)
}
if (!exists("gauss_params_raw")) {
  gauss_params_raw <- load_gauss_params_raw(rank_panel, K_max, cfg = CFG)
}
if (!exists("sm_params")) {
  sm_params <- load_gauss_params_smoothed(gauss_params_raw, K_max, Smoothing_h, cfg = CFG)
}

if (!exists("emp_cdc")) {
  emp_cdc <- load_emp_cdc(endpoint_weekly, K_cut, cfg = CFG)
}
if (!exists("emp_targets")) {
  emp_targets <- load_emp_targets(endpoint_weekly, K_cut, CFG$horizons_durable, bucket_def, cfg = CFG)
}

entrant_pool <- build_entrant_pool(endpoint_weekly, K_cut, CFG$K_tail_buffer, max_rank_seen)
entrant_sampler <- make_entrant_sampler(entrant_pool)
w0_ext <- build_w0_ext(endpoint_weekly, K_max, entrant_sampler)

mean_vec_s <- sm_params$mean_dlogw_s
sd_vec_s <- sm_params$sd_dlogw_s
```

```{r sigma-calibration}
sigma_grid <- seq(0.15, 0.40, by = 0.01)

sigma_results <- purrr::map_dfr(
  sigma_grid,
  calibrate_sigma,
  w0_ext = w0_ext,
  mean_vec_s = mean_vec_s,
  sd_vec_s = sd_vec_s
) %>%
  arrange(score)

best_sigma <- sigma_results %>% slice(1)
best_sigma

ggplot(sigma_results, aes(x = sigma, y = score)) +
  geom_line() +
  geom_vline(xintercept = best_sigma$sigma, linetype = "dashed") +
  labs(
    title = "Sigma calibration (v10-style)",
    x = "sigma",
    y = "score = durable RMSE + 0.25 * CDC RMSE"
  ) +
  theme_bw()
```

## Sigma calibration by horizon (v10-style) (optional)

This is the v10 extension: pick a separate `sigma` for each horizon by minimizing the horizon-specific
durable-change mismatch.

```{r sigma-calibration-by-horizon}
if (isTRUE(CFG$run_grid_search)) {
  sigma_h_tbl <- purrr::map_dfr(
    sigma_grid,
    sigma_by_horizon,
    w0_ext = w0_ext,
    mean_vec_s = mean_vec_s,
    sd_vec_s = sd_vec_s
  )

  best_sigma_by_horizon <- sigma_h_tbl %>%
    group_by(horizon) %>%
    arrange(rmse_h, .by_group = TRUE) %>%
    slice(1) %>%
    ungroup()

  best_sigma_by_horizon

  ggplot(sigma_h_tbl, aes(x = sigma, y = rmse_h)) +
    geom_line() +
    facet_wrap(~ horizon, scales = "free_y") +
    labs(
      title = "Sigma calibration by horizon",
      x = "sigma",
      y = "RMSE within horizon"
    ) +
    theme_bw()
}
```

## Grid search over (mu, sigma, h) (v9-style) (optional)

This reproduces v9’s idea: jointly tune a global drift `mu`, global volatility scale `sigma`,
and smoothing half-width `h`, using a weighted score combining CDC and durable-change mismatch.

```{r grid-search-mu-sigma-h}
if (isTRUE(CFG$run_grid_search)) {

  mu_grid <- c(-0.01, -0.005, 0.0, 0.005)
  sigma_grid2 <- c(0.1, 0.3, 0.5, 0.7, 1.0)
  h_grid <- c(0L, 2L, 5L)

  lambda_cdc <- 0.7

  param_grid <- tidyr::expand_grid(mu = mu_grid, sigma = sigma_grid2, h = h_grid)
  grid_results <- vector("list", nrow(param_grid))

  for (i in seq_len(nrow(param_grid))) {

    mu_i <- param_grid$mu[i]
    sigma_i <- param_grid$sigma[i]
    h_i <- as.integer(param_grid$h[i])

    sm_i <- build_gaussian_params_smoothed(gauss_params_raw, K_max, h = h_i)

    sim_i <- simulate_rank_paths(
      w0 = w0_ext,
      K_cut = K_cut,
      K_max = K_max,
      T = CFG$sim_T_weeks,
      n_paths = CFG$sim_n_paths,
      mu = mu_i,
      sigma = sigma_i,
      entry_frac = CFG$sim_entry_frac,
      mean_vec = sm_i$mean_dlogw_s,
      sd_vec = sm_i$sd_dlogw_s,
      horizons = CFG$horizons_durable,
      K_xi = NULL,
      entrant_sampler = entrant_sampler
    )

    sim_cdc_i <- sim_i$snapshots %>%
      group_by(rank) %>%
      summarise(w_bar = mean(share, na.rm = TRUE), .groups = "drop")

    rmse_cdc_i <- cdc_rmse(emp_cdc, sim_cdc_i)
    rmse_dur_i <- durable_rmse(emp_targets, sim_i$growth, bucket_def)

    score_i <- lambda_cdc * rmse_cdc_i + (1 - lambda_cdc) * rmse_dur_i

    grid_results[[i]] <- tibble(
      mu = mu_i, sigma = sigma_i, h = h_i,
      rmse_cdc = rmse_cdc_i,
      rmse_durable = rmse_dur_i,
      score = score_i
    )
  }

  grid_tbl <- bind_rows(grid_results) %>% arrange(score)
  grid_tbl %>% slice(1:10)

  ggplot(grid_tbl, aes(x = score)) +
    geom_histogram(bins = 30) +
    labs(
      title = "Grid-search score distribution",
      x = "score",
      y = "count"
    ) +
    theme_bw()
}
```
